### 2023-02-02 scripts/0_whole_game.qmd
removed extra files

---
title: "whole_game"
format: html
editor: visual
---

## Setup

```{r setup}
library(tidyverse)
library(terra)
library(sf)
library(bcmaps)
library(sgsR)
library(exactextractr)

#### setup ----
terraOptions(
  memfrac = 0.75,
  tempdir = "E:\\temp",
  todisk = T,
  progress = 100
)
options(scipen = 999)

source(here::here("scripts", "get_keys.R"))
```

Needed things

```{r loads}
vlce <- rast("G:/mosaiced/VLCE2.0/LC_Class_HMM_2015_v20_v20.dat")
bcb <- bcmaps::bc_bound() %>%
  vect()

bcb_rast <- bcb %>%
  rasterize(y = vlce) %>%
  crop(bcb)

vlce <- crop(vlce, bcb_rast) %>%
  mask(bcb_rast)
```

Run the focal cell analysis. Only runs if not done yet via if statement

```{r run-focal}
focal_loc <- "G:/mosaiced/focal"

if (length(list.files(focal_loc)) != 11) {
  print("running focal analysis on vlce and 10 structure variables")
  rmarkdown::render(here::here("scripts", "1_bc_focal_processing.qmd"))
} else {
  print("focal analysis already done")
}
```

Run the sampling across each BEC zone

```{r run-sample}
sample_loc <- here::here("data", "samples")

list.files(sample_loc, pattern = ".shp$")

rmarkdown::render(here::here("scripts", "2_sampling.qmd"))
```

```{r}
print("hello")
```