### 2023-02-02 scripts/1_bc_focal_processing.qmd
changed to whole game approach rather than various random scripts. uses python for focal processing. implemented log %>% scale for all variables. prior to log %>% scale, DHI vars are min-max scaled from 0-1.

---
title: "focals"
format: html
editor: visual
---

Make extended tilings

```{r make-grid}
grid_ncol <- 2
grid_nrow <- 2

# make tiles to grab samples from
grid <-
  rast(ncol = grid_ncol,
       nrow = grid_ncol,
       ext = ext(bcb_rast))

values(grid) <- 1:(grid_ncol*grid_nrow)

tiles_focal_loc <- here::here("scratch", "tiles_focal")

dir.create(tiles_focal_loc, showWarnings = F)

focal_files <- list.files(tiles_focal_loc,
                        pattern = "^b.*\\.dat$",
                        full.names = T)

max_tile <- str_extract(focal_files, "[0-9]{1,2}.dat$") %>%
  str_extract("[0-9]{1,2}") %>%
  as.numeric() %>%
  max()

if (max_tile != (grid_ncol * grid_nrow)) {
  file.remove(list.files(focal_files, full.names = T))

  focal_files <- terra::makeTiles(
    x = bcb_rast,
    y = grid,
    here::here(tiles_focal_loc, "tile_.dat"),
    na.rm = TRUE,
    overwrite = TRUE,
    filetype = "ENVI",
    datatype = "INT1U"
  )
}

extended_focal <- focal_files %>%
  map(rast) %>%
  map(extend, y = 1)
```

```{r forest-rcl}
forest_rcl <- keys$vlce %>%
  mutate(class_name = ifelse(forest == "Forest", class_val, NA)) %>%
  select(class_val, class_name) %>%
  as.matrix(ncol = 2)

forests <- classify(vlce, forest_rcl)
```

Function to allow mapping on the inverted variable order

```{r function}
rev_crop <- function(x, y) {
  crop(y, x)
}
```

Extended tiles (vlce)

```{r tile-extend}
forests_focal_tile_loc <- here::here(tiles_focal_loc, "forests", "input")
dir.create(forests_focal_tile_loc, showWarnings = F, recursive = T)

forest_save_names <- here::here(forests_focal_tile_loc,
                                paste0("focal-forests-", 1:length(extended_focal), ".dat"))

forests_extend <- extended_focal %>%
  map(rev_crop, y = forests) %>%
  map2(.x = ., .y = forest_save_names, .f = writeRaster,
       filetype = "ENVI", overwrite = T)
```

Make clean structure directories

```{r dir-creation}
structure_focal_tile_loc <- here::here(tiles_focal_loc, "structure", "input")
dir.create(structure_focal_tile_loc, showWarnings = F, recursive = T)

struct_loc <- here::here("G://", "mosaiced", "structure")

struct_vars <- list.dirs(struct_loc, full.names = F) %>%
  stringi::stri_remove_empty()

structure_subdirs <- here::here(structure_focal_tile_loc, struct_vars)

map(structure_subdirs, dir.create, showWarnings = F)
```

Make extended tiles for each structure variable

```{r structure-extend}
for(var in struct_vars) {
  print(var)

  r <- here::here(struct_loc, var, paste0(var, "_2015.dat")) %>%
  rast() %>%
  crop(bcb_rast) %>%
  mask(bcb_rast)

struct_save_names <- here::here(structure_focal_tile_loc,
                                var,
                                paste0("focal-", var, "-", 1:length(extended_focal), ".dat"))

struct_extend <- extended_focal %>%
  map(rev_crop, y = r) %>%
  map2(.x = ., .y = struct_save_names, .f = writeRaster,
       filetype = "ENVI", overwrite = T
  )
}
```

```{r reticulate}
# this does the focal cell processing for all of BC based on the tiles 
# created earlier in this script
library(reticulate)
use_python("C:/Users/evanmuis/Miniconda3/envs/geospatial/python.exe")

source_python(here::here("scripts", "1a_focal_tiles.py"))

rm(stack)
rm(nptile)
rm(open_tile)
rm(valid_locs)
```

```{r mosaic}
# mosaics the focal tiles together

folders <- "E:/Sync/Masters/analysis_sem/scratch/tiles_focal" %>%
  list.dirs() %>%
  str_subset("output")

for_fold <- str_subset(folders, "forests")

struct_fold <- folders %>%
  str_subset("structure") %>%
  str_subset("output$", negate = T)

iter_folds = c(for_fold, struct_fold)

focal_save <- here::here("G:/", "mosaiced", "focal")

# only do the 3 currently used

# iter_folds <- c(iter_folds[1], iter_folds[8:9])

mosaic_rcl <- matrix(c(0, 1, NA, 1), ncol = 2)

for (folder in iter_folds) {
  print(folder)
  var = folder %>%
    list.files() %>%
    .[1] %>% # first file
    str_split("-") %>%
    .[[1]] %>% # unlist
    .[2] # second item (variable)

  # lol what the fuck is this evan nobody can read it
  # take first file w/o full names (since naming conventions the same)
  # make it not a list
  # take the second item in that vector (should always be the variable)

  save_name = here::here(focal_save, paste0("focal-", var, ".tif"))
  print(save_name)

  mosaiced <- list.files(folder, full.names = T, pattern = ".dat$") %>%
    map(rast) %>%
    map(trim, padding = 0, value = 0) %>%
    sprc() %>%
    mosaic(fun = "min") %>%
    extend(bcb_rast)

  mosaiced_rc <- classify(mosaiced, mosaic_rcl)

  #mosaiced <- mosaiced == 1

  # if (var != "forests") {
  #   save_name = here::here(focal_save, "structure", var, paste0("focal-", var, ".tif"))
  # }

  plot(mosaiced_rc, col = "green")

  dir.create(dirname(save_name), showWarnings = F)

  writeRaster(mosaiced_rc, save_name, overwrite = T)

}
```