### 2023-02-02 scripts/3_sem_visualization.qmd

---
title: "sem"
format: html
editor: visual
---

```{r}
library(lavaan)

theme_set(theme_light())
```

extract datasets

```{r}
samples <- here::here("data", "samples", "bec") %>%
  list.files(full.names = T, pattern = ".shp$") %>%
  map(read_sf) %>%
  bind_rows()

samples %>%
  st_drop_geometry() %>%
  count(zone) %>%
  ggplot(aes(
    x = n,
    y = fct_reorder(zone, n),
    fill = zone
  )) +
  geom_col() +
  ggsci::scale_fill_igv() +
  theme_bw() +
  labs(x = "n",
       y = NULL,
       fill = NULL)
```

```{r}
all_rasts_loc <- here::here("data", "rasts", "all_rasts.dat")

if (!file.exists(all_rasts_loc)) {
  struct_locs <- here::here("G://", "mosaiced", "structure")
  struct_rasts <- list.files(
    struct_locs,
    recursive = T,
    pattern = ".dat$",
    full.names = T
  ) %>%
    map(rast) %>%
    rast()

  struct_names <- str_split(sources(struct_rasts), pattern = "/") %>%
    lapply("[[", 4) %>% # get fourth index from names
    unlist()

  names(struct_rasts) <- struct_names

  struct_rasts <- struct_rasts %>%
    crop(bcb_rast, mask = T)

  dhi_rasts <-
    list.files(here::here("G://", "mosaiced", "DHI"),
               pattern = "2010s.tif$",
               full.names = T) %>%
    map(rast) %>%
    rast()

  dhi_names <- str_split(sources(dhi_rasts), pattern = "/") %>%
    lapply("[[", 4) %>% # get fourth index from names
    unlist() %>%
    str_split(pattern = "[[:punct:]]") %>%
    lapply("[[", 2) %>% # get second index from split names
    unlist()

  names(dhi_rasts) <- dhi_names

  dhi_rasts <- dhi_rasts %>%
    crop(bcb_rast, mask = T)

  dhi_mm <- minmax(dhi_rasts)

  dhi_rasts_mm <- (dhi_rasts - dhi_mm[1,]) / (dhi_mm[2,] - dhi_mm[1,])

  all_rasts <- c(vlce, struct_rasts, dhi_rasts_mm)

  writeRaster(all_rasts, all_rasts_loc, overwrite = T, filetype = "envi",
              datatype = "FLT4S")
}

all_rasts <- rast(all_rasts_loc)
```

```{r}
extracted_samples <- extract(all_rasts, samples, bind = T)
```

)

```{r}
invalid_zones <- c("CMA", "IMA", "BAFA", "BG")
# alpine (CMA, IMA, BAFA) & non-forest dominant (BG)

extracted_df <- extracted_samples %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  filter(!(loreys_height < elev_mean | elev_p95 < elev_mean)) %>%
  pivot_longer(basal_area:total_biomass, names_to = "variable") %>%
  left_join(keys$continuous) %>%
  mutate(value = value / divide_by) %>%
  select(zone, category, ends_with("DHI"), variable, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  unchop(everything()) %>% # somehow does the fix i want
  filter(elev_cv <= 1) %>% # remove erroneous values
  drop_na() # remove anything w/ NA

# groups by zone and all other structure variables 
# takes the mean of each DHI dataset
# effectively removing duplicates, but takes the mean instead of the first value
extracted_df_b <- extracted_df %>%
  group_by(across(c(zone, category, basal_area:total_biomass))) %>%
  summarize(across(ends_with("DHI"), mean)) %>%
  ungroup()
```

```{r}
# downsampling procedure - allows for comparisons between models if sample size is the same (AIC and BIC are based on sample)

downsample_df <- extracted_df_b %>%
  count(zone) %>%
  filter(!(zone %in% invalid_zones))

downsample_no <- downsample_df %>%
  pull(n) %>%
  min()
```

```{r}
set.seed(69420)

clean_df <- extracted_df_b %>%
  mutate(
    overstory_per = percentage_first_returns_above_mean,
    understory_per = percentage_first_returns_above_2m - percentage_first_returns_above_mean,
    total_cover = percentage_first_returns_above_2m,
    overstory_range = loreys_height - elev_mean,
    understory_range = elev_mean - 2
  ) %>%
  relocate(zone,
           category,
           ends_with("DHI")) %>%
  filter(!(zone %in% invalid_zones)) %>%
  group_by(zone) %>%
  slice_sample(n = downsample_no) %>%
  select(!starts_with("perc"))
```

```{r}
mylog <- function(data) {
  if (any(data <= 0)) {
    return(log1p(data))
  } else {
    return(log(data))
  }
}

normal <- function(data) {
  data %>%
    mylog() %>%
    scale() %>% 
    as.numeric()
}

log_normal_df <- clean_df %>%
  group_by(zone) %>%
  mutate(across(CumDHI:understory_range, .fns = ~ normal(.x)))

log_normal_df %>%
  summarize(across(CumDHI:understory_range, .fns = ~ shapiro.test(.x)$p.value))
```

```{r}
log_splits <- log_normal_df %>%
  group_by(zone) %>%
  group_split()
```

sample each bec zone down to the lowest number of samples

```{r}
set.seed(69420)

zone_dfs <- extracted_df %>%
  filter(!(zone %in% invalid_zones)) %>%
  group_by(zone) %>%
  group_split() %>%
  map(slice_sample, n = downsample_no)
```

```{r}
library(Hmisc)
library(corrplot)
corr <- log_normal_df %>% 
  ungroup() %>%
  select(CumDHI:total_biomass) %>%
  as.matrix() %>%
  rcorr()

corrplot(corr$r,
         type = "upper",
         tl.col = "black",
         tl.srt = 45,
         p.mat = corr$P, 
         sig.level = 0.01)
```

```{r}
sample_names <- names(extracted_samples)

dhi_vars <-
  sample_names[(length(sample_names) - 2):length(sample_names)]

v_models <-
  paste0(
    dhi_vars,
    " ~ loreys_height + total_biomass + elev_stddev
elev_stddev ~ loreys_height
total_biomass ~ loreys_height"
  )

h_models <-
  paste0(
    dhi_vars,
    " ~ basal_area + percentage_first_returns_above_2m + percentage_first_returns_above_mean
percentage_first_returns_above_mean ~ basal_area
percentage_first_returns_above_2m ~ basal_area"
  )

models <- c(v_models, h_models)
```

```{r some-weird-map-lol}
bec_zones <- bec() %>%
  mutate(area = st_area(.)) %>%
  group_by(ZONE) %>%
  dplyr::summarize(area = sum(area)) %>%
  st_intersection(bc_bound())

bec_counts_sf <- bec_zones %>%
  left_join(bec_counts, by = c("ZONE" = "zone")) %>%
  select(ZONE, n) %>%
  mutate(ZONE = ifelse(!(ZONE %in% invalid_zones), ZONE, "Not Included"))

ggplot(bec_counts_sf) +
  geom_sf(aes(fill = ZONE), colour = "#00000000") +
  ggsci::scale_fill_igv() +
  scale_alpha_manual(values = c(0.5, 1)) +
  theme_void()
```

```{r zonal-models}
sem_extract <- function(split_df, model) {
  zone <- split_df %>%
    pull(zone) %>%
    unique()

  var <- str_split(model, pattern = " ")[[1]][1]
  fit <- sem(model = model, data = split_df)

  fits <-
    fitmeasures(fit, c("chisq", "df", "pvalue", "cfi", "rmsea", "aic", "r2", "srmr"))
  pars <- parameterestimates(fit) %>%
    tibble() %>%
    mutate(zone = zone, var = var, model = model)

  fits <- tibble(fit = names(fits), value = as.numeric(fits)) %>%
    pivot_wider(names_from = fit, values_from = value) %>%
    mutate(zone = zone,
           var = var, model = model)

  interpreted <- fit %>%
    effectsize::interpret() %>%
    mutate(zone = zone, var = var, model = model)

  return(list(
    # fitmeasures = fits,
    parameters = pars,
    interpreted = interpreted,
    fits = fits
  ))
}

crossed <- crossing(zone_dfs, models)

dfs <- crossed %>% pull(zone_dfs)
mods <- crossed %>% pull(models)

outs <- map2(.x = dfs, .y = mods, .f = sem_extract)

params <- map_df(outs, function(x) {
    x$parameters
  })

# fitmeasures <- map_df(outs, function(x) {
#   x$fitmeasures
# })

fitmeasures <- map_df(outs, function(x) {
  x$interpreted
})

fits <- map_df(outs, function(x) {
  x$fits
})
```

```{r}
vlce_samples <- here::here("data", "samples", "vlce") %>%
  list.files(pattern = ".shp", full.names = T) %>%
  map(read_sf) %>%
  bind_rows()

extracted_vlce <- extract(all_rasts, vlce_samples)
```

```{r}
extracted_vlce %>%
  st_drop_geometry() %>%
  count(category)
```

```{r}
fitmeasures %>% 
  group_by(zone) %>%
  dplyr::summarize(across(cfi:aic, mean))

fitmeasures %>%
  mutate(v_h = ifelse(model %in% h_models, "h", "v")) %>%
  group_by(model) %>%
  dplyr::summarize(across(c(cfi:aic, chisq), mean))
```

```{r}
params %>%
  filter(op == "~") %>%
  group_by(model, lhs, rhs) %>%
  dplyr::summarize(per_sig = (pvalue > 0.05) / n()) %>%
  view()
```

```{r}
test_mod <- "CumDHI ~ loreys_height + percentage_first_returns_above_2m + percentage_first_returns_above_mean
percentage_first_returns_above_2m ~ total_biomass
percentage_first_returns_above_mean ~ percentage_first_returns_above_2m + total_biomass
total_biomass ~ loreys_height"

data <- dfs[[19]] %>%
  mutate(across(basal_area:total_biomass, scale),
         across(basal_area:total_biomass, as.numeric))

(egg <- sem(test_mod, data = data))

varTable(egg)
fitmeasures(egg)
parameterestimates(egg)

graph_sem(egg)
```

```{r}
vlce_samples <- here::here("data", "samples", "vlce") %>%
  list.files(full.names = T, pattern = ".shp$") %>%
  map(read_sf) %>%
  bind_rows()

vlce_samples
```