### 2023-04-12 scripts/manuscript/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  review,
  3p,
  twocolumn]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\usepackage{float}
\makeatletter
\let\oldlt\longtable
\let\endoldlt\endlongtable
\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
\def\longtable@i[#1]{\begin{figure}[H]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldlt[#1]
}
\def\longtable@ii{\begin{figure}[H]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldlt
}
\def\endlongtable{\endoldlt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother
\journal{Remote Sensing of Environment}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Disentangling linkages between satellite derived forest structure and productivity across a range of forest types and ecosystems.},
  pdfauthor={Evan R. Muise; Margaret E. Andrew; Nicholas C. Coops; Txomin Hermosilla; A. Cole Burton; Stephen S. Ban},
  pdfkeywords={structural equation modelling, remote
sensing, landsat, forest structure, forest productivity, dynamic habitat
indices},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Disentangling linkages between satellite derived forest structure
and productivity across a range of forest types and ecosystems.}
\author[1]{Evan R. Muise%
\corref{cor1}%
}
 \ead{evan.muise@student.ubc.ca} 
\author[2]{Margaret E. Andrew%
%
}
 \ead{M.Andrew@murdoch.edu.au} 
\author[1]{Nicholas C. Coops%
%
}
 \ead{nicholas.coops@ubc.ca} 
\author[3]{Txomin Hermosilla%
%
}
 \ead{txomin.hermosillagomez@NRCan-RNCan.gc.ca} 
\author[1]{A. Cole Burton%
%
}
 \ead{cole.burton@ubc.ca} 
\author[4]{Stephen S. Ban%
%
}
 \ead{Stephen.Ban@gov.bc.ca}

\affiliation[1]{organization={University of British Columbia, Forest
Resources Management},addressline={2424 Main Mall},city={Vancouver, BC,
Canada},postcode={V6T 1Z4},postcodesep={}}
\affiliation[2]{organization={Murdoch University, Environmental and
Conservation Sciences and Harry Butler Institute},city={Murdoch, WA,
Australia},postcode={6150},postcodesep={}}
\affiliation[3]{organization={Natural Resources Canada, Canada Forest
Service (Pacific Forestry Centre)},addressline={506 Burnside Rd
W},city={Victoria, BC, Canada},postcode={V8Z 1M5},postcodesep={}}
\affiliation[4]{organization={Ministry of Environment and Climate Change
Strategy, BC Parks},addressline={525 Superior Street},city={Victoria,
BC, Canada},postcode={V8V 1T7},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
{[}insert abstract here - nicholas ignore the formatting on the authors
for now, i can output to pdf and have it formatted nicely for
submission{]}
\end{abstract}

\begin{keyword}
    structural equation modelling \sep remote
sensing \sep landsat \sep forest structure \sep forest
productivity \sep 
    dynamic habitat indices
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\begin{verbatim}
This is lavaan 0.6-14
lavaan is FREE software! Please report any bugs.
\end{verbatim}

\begin{verbatim}
terra 1.7.3
\end{verbatim}

\begin{verbatim}
Linking to GEOS 3.9.3, GDAL 3.5.2, PROJ 8.2.1; sf_use_s2() is TRUE
\end{verbatim}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Biodiversity is the summation of variation in biological life, across
genes, species, communities, and ecosystems. Currently, biodiversity is
in decline, facing extinction rates above the background extinction rate
\citep{thomas2004, urban2015}, and homogenization of communities at
various scales \citep{mcgill2015}. In response, the global biodiversity
community is making efforts to assess and halt the degradation of
biodiversity on earth. The Group for Earth Observation Biodiversity
Observation Network has developed the Essential Biodiversity Variables
\citep[EBVs,][]{pereira2013}, designed as an analog to the Essential
Climate Variables framework \citep{bojinski2014}. EBVs are designed to
be global in scope, relevant to biodiversity information, feasible to
utilize, and complementary to one another \citep{skidmore2021}. There
are six EBV classes, each of which correspond to a different facet of
biodiversity including species populations, species traits, community
composition, ecosystem structure, ecosystem function, and genetic
composition \citep{pereira2013}.

Satellite remote sensing has proven to be capable of measuring five of
the six classes, with the exception being genetic composition, which
requires in-situ observation and sampling \citep{skidmore2021}. Species
populations - and in turn community composition - can be assessed with
very-high-resolution imagery to identify tree species at the tree-crown
scale, however it is difficult and computationally expensive to extend
these analyses to broader extents \citep{fassnacht2016, graves2016}.
Species traits such as phenology have been on a single-tree scale using
PlanetScope imagery and drone-based measurements \citep{wu2021}. Coarser
measurements, such as those taken by the MODIS \citep{zhang2003},
Landsat \citep{fisher2006}, or Sentinel
\citep{helfenstein2022, darvishzadeh2019} series of satellites can
monitor these functional processes at larger extents, corresponding to
ecosystem function, but the coarse spatial resolution removes the
ability to relate these traits to individual trees. Remote sensing data
has been shown to be effective at monitoring ecosystem structure at
regional to global extents through the use of optical imagery
\citep{cohen2004} and lidar datasets
\citep{lefsky2002, lang2021, neuenschwander2019}.

Notably, while remote sensing can provide information on the
species-level EBVs, this information is typically acquired with an
\emph{ad-hoc} approach requiring high spatial and spectral resolution
data \citep{skidmore2021}. This data can often only be collected at
small extents, rather than the global or regional scales required for
biodiversity trend assessment \citep{valdez2023}. The two
landscape-level EBVs (ecosystem structure and function) are well suited
to be examined at global or regional scales using moderate-resolution
satellite imagery, such as those provided by the Landsat or Sentinel
series of satellites \citep{skidmore2021}.

Simple metrics to complex for structure.

Forest structural diversity has been linked to biodiversity at various
scales \citep{guo2017, bergen2009, gao2014}. Increased structural
complexity is hypothesized to create additional niches, leading to
increased species diversity \citep{bergen2009}, which has been
frequently demonstrated using avian species diversity metrics
\citep{macarthur1961}. For example: \citet{herniman2020} used spectral
and lidar derived forest structure data to model avian habitat
suitability; \citet{clawges2008} found that lidar derived forest
structural attributes are capable of identifying habitat types
associated with avian species in pine/aspen forests; \citet{goetz2007}
used canopy structural diversity to predict bird species richness,
finding that canopy vertical distribution was the strongest predictor of
species richness. Forest structural metrics have also been used to study
other clades and habitats \citep{davies2014, nelson2005}. Many metrics
derived from lidar remote sensing have been used as local indicators of
biodiversity, including simple metrics such as canopy cover and canopy
height as well as derived metrics including vertical profiles,
aboveground biomass \citep{lefsky1999}. Other second order derived
metrics such as canopy texture, height class distribution, edges, and
patch metrics have also been used to examine habitat and biodiversity at
landscape scales \citep{bergen2009}.

Simple productivity (NDVI, indices) to complex aggregates (DHIs)

The dynamic habitat indices (DHIs) are indicators of productivity
calculated by summarizing vegetation indices over the course of one (or
multiple) years \citep{radeloff2019}. These indices have been related to
multiple facets of biodiversity at a range of scales, including species
occurrence and abundance \citep{razenkova2020}, alpha
\citep{radeloff2019} and beta diversity \citep{andrew2012}. Hypotheses
behind the biodiversity productivity relationships have been
established, including the species-energy hypothesis, the environmental
stress hypothesis, and the environmental stability hypothesis
\citep{coops2019}. The cumulative DHI calculates the total amount of
energy available in a given pixel over the course of a year. Cumulative
DHI is strongly linked to the available energy hypothesis, which
suggests that with greater available energy species richness will
increase \citep{wright1983}. The minimum DHI, which calculates the
lowest productivity over the course of a year can be matched to the
environmental stress hypothesis, which proposes that higher levels of
minimum available energy will lead to higher species richness
\citep{currie2004}. Finally, the variation DHI, which calculates the
coefficient of variance in a vegetation index through the course of a
year, corresponds to the environmental stability hypothesis which states
that lower energy variation throughout a year will lead to increased
species richness \citep{williams2008}.

Biodiversity has been shown to be linked to both forest structure
\citep{guo2017, gao2014}, and productivity \citep{radeloff2019} at a
range of scales. Research has highlighted that the relationship between
productivity and biodiversity is reciprocal \citep{worm2003}, making
productivity a suitable biodiversity indicator. Within a forest
environment, there are many possible ecosystem structure and function
metrics that can be assessed. Structural variables range in complexity
from simple (canopy cover; canopy height), to more complex (elevation
coefficient of variation; elevation standard deviation) to modelled
(aboveground biomass; basal area), all of which can be assessed using
lidar data \citep{coops2021}. Data on ecosystem functionality ranges
from phenology metrics (date of leaf on/leaf off; length of growing
season) to functional traits (chlorophyll content; leaf area index) to
productivity estimates (GPP; total available energy)
\citep{pettorelli2018}. Advances in satellite remote sensing processing
have allowed 3d forest structure data to be imputed across wide spatial
scales \citep{matasci2018, coops2021} using data fusion approaches
involving collected lidar data and optical/radar data. Other advances in
image compositing have allowed yearly summaries of vegetation
productivity to be calculated at regional to global scales, summarizing
the yearly total, minimum, and variation in energy availability at
similar spatial resolutions \citep{radeloff2019}.

In forested ecosystems, stand structural attributes have also been shown
to influence productivity, with a range of responses dependent on
environmental conditions \citep{ali2019}. Linkages between forest
structure and productivity (namely, vegetation indices) have been
examined for nearly 20 years
\citetext{\citealp[\citet{knyazikhin1998}]{huete2002}; \citealp{myneni1994}}.
While there is significant theoretical and empirical evidence for their
relationship at single time points (within a single image)
\citep{myneni1994}, various relationship directions and shapes have been
found between forest structure and productivity metrics \citep{ali2019}.
These relationships, their shapes, and their strengths have been
attributed to multiple possible hypotheses and have been shown to vary
based on environmental conditions \citep{ali2019}. The relationship
between forest structural diversity metrics and annual productivity
summaries has yet to be fully examined, including the DHIs.

Understanding the complementarity between potential EBVs is an integral
component of their creation \citep{skidmore2021}. Structural equation
modelling and path analyses have been commonly used in ecology to assess
the causal effects behind various hypotheses \citep{fan2016, grace2010},
including the relationship between forest structure, functioning, and
biodiversity \citep{ali2019}. In this paper we seek to untangle the
relationship between two EBVs: forest structure diversity metrics and
yearly summaries of forest productivity. To accomplish this, we assess
this relationship using path analysis to assess the direct and indirect
(as mediated by more complex forest structural diversity metrics)
effects of commonly collected forest structural metrics (canopy height
and canopy cover) on yearly productivity summaries. Further, we use
exploratory structural equation modelling (ESEM) to identify the latent
variables driving the productivity metrics. We run this analysis both
globally, and stratified by the forested ecosystems of British Columbia,
Canada, and compare models using AIC scores and global fit measures
across the ecosystems. The results from this study will assess the
linkages and complementarity between two biodiversity indicators.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{study-area}{%
\subsection{Study Area}\label{study-area}}

British Columbia is the westernmost province of Canada, and is home to a
variety of terrestrial ecosystems \citep{pojar1987}. Approximately 64\%
of the province is forested \citep{bcministryofforests2003}. There is a
large amount of ecosystem variation in the province, with large climate
and topographic gradients. The Biogeoclimatic Ecosystem Classification
(BEC) system identifies 16 zones based on the dominant tree species and
the ecosystems general climate. These zones can be further split into
subzones, variants, and phases based on microclimate, precipitation, and
topography \citep{pojar1987}.

\{insert figure showing ecosystems of interest\} need to remake figure
showing ecosystems, highlighting that bunchgrass isn't included

\{table showing elevations and climate ranges for each included bec
zone\}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{ntems}{%
\subsubsection{NTEMS}\label{ntems}}

\hypertarget{dynamic-habitat-indices}{%
\subsubsection{Dynamic Habitat Indices}\label{dynamic-habitat-indices}}

The Dynamic Habitat Indices are a set of satellite remote sensing
derived productivity variables that summarize the cumulative amount of
available energy, the minimum available energy, and the variation in
available energy throughout a given year \citep{radeloff2019}. The DHIs
have previously been produced at a global extent using MODIS imagery
\citep{radeloff2019}, however, recent studies have examined how these
indices can be constructed at a finer spatial resolution using
multi-annual Landsat imagery.

\hypertarget{sampling}{%
\subsection{Sampling}\label{sampling}}

We conducted model based sampling across the fifteen forest dominated
ecosystems found within British Columbia {[}BC figure reference, I have
not made this figure yet{]}. Samples were randomly selected within BEC
zones of interest alongside multiple criteria. Each sampled pixel had to
have a forested land cover class (coniferous, deciduous, mixed-wood, or
wetland-treed), and be surrounded by the same land cover class.
Additionally, each pixel had to have a coefficient of variation less
than 0.5 in the Lorey's height and canopy cover forest structural
metrics. A maximum of 3000 samples were sampled in each BEC zone.

Due to variations in ecosystem size, differing numbers of samples were
generated for each ecosystem. To account for this during model
comparison, each ecosystem and land cover sample set was downsampled to
the minimum number of samples for a single BEC zone (in this case, 170).
To meet the normality assumptions of path analysis, all variables were
natural-log transformed and standardized, as per \citet{grace2016}.
Variables containing zeros were natural-log plus one transformed.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

\hypertarget{path-analysis}{%
\subsubsection{Path Analysis}\label{path-analysis}}

To determine the relationships between simple lidar derived forest
structural attributes, complex/derived forest structural attributes, and
forest productivity, we used path analysis to analyze two causal models
(Figure~\ref{fig-pathdag}). To determine primary drivers of the three
DHIs, we will summarize the predictors across ecosystems by counting the
strongest predictor in each BEC zone. This will determine if the primary
driver for each facet of the DHI is simple or complex, and allow us to
assess ecosystem differences.

\begin{figure}

{\centering \includegraphics[width=1.6in,height=\textheight]{../../outputs/path_dags.png}

}

\caption{\label{fig-pathdag}Proposed path diagrams}

\end{figure}

\hypertarget{exploratory-structural-equation-modelling}{%
\subsubsection{Exploratory Structural Equation
Modelling}\label{exploratory-structural-equation-modelling}}

Secondly, we will use exploratory structural equation modelling
\citep{marsh2020, asparouhov2009} to identify latent forest structural
variables within the data, and use these latent variables to predict the
dynamic habitat indices. Exploratory structural equation modelling is a
combination of exploratory factor analysis (EFA) and structural equation
modelling, which relaxes the strict requirement of zero cross-loadings
in confirmatory factor analysis and allows for less strict measurement
models to be used. ESEM is used when it is known that there is a latent
structure to the data, but the specific indicator variables have not yet
been determined.

One advantage of ESEM is that it can create varying numbers of latent
variables. For our analysis, we first ran exploratory factor analysis on
all forest structural attributes both globally and by BEC zone with
between 1-4 potential latent variables. We chose the most parsimonious
EFA model with the lowest AIC scores for each ecosystem, leading to
varying numbers of latent variables. We then determined the anchoring
indicator for each latent variable. The anchors were calculated by
choosing the indicator variable with the largest difference between the
maximum value in a given loading compared to to said indicators loadings
in all other latent variables. Each anchor variable was then assigned to
a named group (Canopy Cover, Height and Biomass, and Structural
Complexity).

Following the EFA, we used structural equation modelling, with latent
variables loadings determined by the EFA. We also filtered the loadings
to be greater than or equal to 0.5, allowing us to examine the number
and composition of latent variables found in each ecosystem. Following
this, we predicted the DHIs in a single SEM with covariances between the
DHIs. If a forest structural attribute did not end up in a latent
variable, it was included in the structural equation model as a DHI
predictor, without a latent variable as a mediator.

\hypertarget{software}{%
\subsection{Software}\label{software}}

The processing for this paper was done in R version 4.2.2
\citep{R-base}. We used \emph{sgsR} \citep{goodbody2023} for sampling,
\emph{terra} \citep{R-terra} for raster analyses, \emph{sf} for vector
analyses \citep{R-sf}, and \emph{lavaan} \citep{lavaan2012} for path
analysis and structural equation modelling.

These focal analyses were preprocessed in R and Python, and randomly
sampled across a masked suitability raster for each BEC zone using the
\emph{sgsR} R package version 1.3.4 \citep{goodbody2023}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{path-analysis-1}{%
\subsection{Path Analysis}\label{path-analysis-1}}

\begin{figure}

{\centering \includegraphics[width=14.74in,height=\textheight]{../../outputs/path_bars.png}

}

\caption{\label{fig-pathbar-prop}Proportion of strongest signficant
forest structural predictors of yearly productivity metrics in models
based on canopy cover vs canopy height.}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=14.74in,height=\textheight]{../../outputs/strongest_predictor_bar.png}

}

\caption{\label{fig-pathbar-pred}Bar plot of strongest predictors in
path analysis for each BEC zone and predicted DHI variable.}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=14.74in,height=\textheight]{../../outputs/path_map_edit.png}

}

\caption{\label{fig-pathmap}Map of strongest structural attribute
predictor strength on yearly productivity metrics by BEC zone for the
two proposed models. If there are no significant predictors, shown in
dark grey}

\end{figure}

\hypertarget{exploratory-structural-equation-modelling-1}{%
\subsection{Exploratory Structural Equation
Modelling}\label{exploratory-structural-equation-modelling-1}}

I also want to run exploratory SEM with and without the DHI's to see
what grouping it goes in. MinDHI has been breaking because some zones
only have 0 as mindhi (makes sense, they are in the alpine).

\begin{figure}

{\centering \includegraphics[width=7in,height=\textheight]{../../outputs/latent_groups.png}

}

\caption{\label{fig-esem-latent}Groups of latent variables with loadings
greater than or equal to 0.5 for each BEC zone. Loadings were determined
using exploratory factor analysis with up to four latent variables, and
selecting the number of possible latent variables with the lowest AIC.
All forest structure variables were potentially included as indicators
in each latent variable.}

\end{figure}

Exploratory factor analysis shows three groups of latent variables
(Figure~\ref{fig-esem-latent}). We assigned the groups to latent
variables indicating canopy cover / biomass metrics, canopy height /
biomass metrics, and structural complexity metrics. Indicators to latent
variable groups by examining anchors for each latent variable. The
anchors were calculated by choosing the indicator variable with the
largest difference between the maximum value in a given loading compared
to to said indicators loadings in all other latent variables.

Both the Ponderosa Pine (PP) and Sub-boreal Pine -- Spruce zones have
one latent variable. Coastal Western Hemlock (CWH), Engelmann Spurce --
Subalpine Fir (ESSF), and Montane Spruce (MS) are the only zones to have
structural complexity based latent variables. CWH is the only zone to
include elevation standard deviation in the structural complexity latent
variable.

\begin{figure}

{\centering \includegraphics[width=14.74in,height=\textheight]{../../outputs/latent_dhis.png}

}

\caption{\label{fig-esem-latent-dhi}Groups of latent variables with
loadings greater than or equal to 0.5 for each BEC zone, including the
DHIs. Loadings were determined using exploratory factor analysis with up
to five latent variables, and selecting the number of possible latent
variables with the lowest AIC. All forest structure variables were
potentially included as indicators in each latent variable.}

\end{figure}

Exploratory factor analysis shows three groups of latent variables. We
assigned these groups to attributes based on the indicator variables in
each latent variable, indicating canopy cover, height and biomass, and
structural complexity. When including the DHIs in this EFA, it is rare
that they are included with latent variables with large loadings. The
cumulative DHI is an indicator for the canopy cover latent varible in
the IDF (Interior Douglas-fir) and MS (Montane Spruce) BEC zones. The
DHIs are not included in any other latent variable, indicating that they
compose different information than the forest structural variables.

\begin{figure}

{\centering \includegraphics[width=14.74in,height=\textheight]{../../outputs/latent_boxplot.png}

}

\caption{\label{fig-latent-boxplots}Boxplots of latent variable loadings
across BEC zones as determined by exploratory factor analysis. Red
dashed lines show latent variable indicator cutoff before being used as
starting values for the SEM models.}

\end{figure}

Figure~\ref{fig-latent-boxplots} shows boxplots of latent variable
loadings from forest structure indicator variables. A maximum of three
valid latent variables were identified across the sixteen BEC zones. We
assigned indicators variables to latent variable groups by examining the
anchoring variables for each latent variable. The anchoring variables
were calculated by choosing the indicator variable with the largest
difference between the maximum value in a given loading compared to said
indicators loadings in all other latent variables. The only DHI variable
that was included in the latent variables after subsetting the latent
variables to only include strong loadings was the Cumulative DHI in the
canopy cover latent variable.

I need to actually make the DAGs and plot the parameters, but I haven't
gotten there yet. I'll plot the global one to show what it looks like,
then aggregate it in some way - probably similar to the path analysis. I
also may include the variable partitioning that Meg talked about in the
email if I can get it working.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\newpage

\renewcommand\refname{References}
  \bibliography{bibliography.bib,packages.bib}

\end{document}