### 2023-04-18 scripts/manuscript/manuscript.qmd

---
title: Disentangling linkages between satellite derived forest structure and productivity across a range of forest types and ecosystems.
author:
  - name: Evan R. Muise
    email: evan.muise@student.ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
    attributes:
        corresponding: true
  - name: Margaret E. Andrew
    email: M.Andrew@murdoch.edu.au
    affiliations: 
        - id: murd
          name: Murdoch University
          department: Environmental and Conservation Sciences and Harry Butler Institute
          city: Murdoch, WA, Australia
          postal-code: 6150
  - name: Nicholas C. Coops
    email: nicholas.coops@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Txomin Hermosilla
    email: txomin.hermosillagomez@NRCan-RNCan.gc.ca
    affiliations: 
        - id: cfs
          name: Natural Resources Canada
          department: Canada Forest Service (Pacific Forestry Centre)
          address: 506 Burnside Rd W
          city: Victoria, BC, Canada
          postal-code: V8Z 1M5
  - name: A. Cole Burton
    email: cole.burton@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Stephen S. Ban
    email: Stephen.Ban@gov.bc.ca
    affiliations: 
        - id: bcp
          name: Ministry of Environment and Climate Change Strategy
          department: BC Parks
          address: 525 Superior Street
          city: Victoria, BC, Canada
          postal-code: V8V 1T7
abstract: |
  [insert abstract here - nicholas ignore the formatting on the authors for now, i can output to pdf and have it formatted nicely for submission]
keywords: 
  - structural equation modelling
  - remote sensing
  - landsat
  - forest structure
  - forest productivity
  - dynamic habitat indices
date: last-modified
bibliography: [bibliography.bib, packages.bib]
csl: remote-sensing-of-environment.csl
format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    journal:
      name: Remote Sensing of Environment
      formatting: review
      model: 3p
      cite-style: authoryear
      layout: twocolumn
  docx: 
    reference-doc: my-styles.docx
    keep-md: true
  pdf: 
    pdf-engine: xelatex
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: false
fig-width: 300
---

```{r}
#| warning: false
#| echo: false
#| output: false

# include all libraries required by the R chunks here
library(vegan)
library(sgsR)
library(terra)
library(sf)

library(bcmaps)

# generate package citations
loaded_pkgs = (.packages())
base_pkgs = getOption("defaultPackages")
contrib_pkgs = loaded_pkgs[!(loaded_pkgs %in% base_pkgs)]
pkgs_file = here::here("scripts", "manuscript", "packages.bib")
knitr::write_bib(contrib_pkgs, pkgs_file)
```

# Introduction

Biodiversity is the summation of variation in biological life, across
genes, species, communities, and ecosystems. Currently, biodiversity is
in decline, facing extinction rates above the background extinction rate
[@thomas2004; @urban2015], and homogenization of communities at various
scales [@mcgill2015]. In response, the global biodiversity community is
making efforts to assess and halt the degradation of biodiversity on
earth. The Group for Earth Observation Biodiversity Observation Network
has developed the Essential Biodiversity Variables [EBVs, @pereira2013],
designed as an analog to the Essential Climate Variables framework
[@bojinski2014]. EBVs are designed to be global in scope, relevant to
biodiversity information, feasible to utilize, and complementary to one
another [@skidmore2021]. There are six EBV classes, each of which
correspond to a different facet of biodiversity including species
populations, species traits, community composition, ecosystem structure,
ecosystem function, and genetic composition [@pereira2013].

Satellite remote sensing has proven to be capable of measuring five of
the six EBV classes, the exception being genetic composition, which
requires in-situ observation and sampling [@skidmore2021]. Species
populations - and in turn community composition - can be assessed with
very-high-resolution imagery to identify tree species at the tree-crown
scale, however it is difficult and computationally expensive to extend
these analyses to broader extents [@fassnacht2016; @graves2016] while
species traits such as vegetation phenology have been observed at the
single-tree scale using, for example, PlanetScope imagery and
drone-based measurements [@wu2021]. Coarser spatial measurements, such
as those taken from satellites by the MODIS [@zhang2003], Landsat
imaging systems[@fisher2006], or Sentinel [@helfenstein2022;
@darvishzadeh2019] can monitor these functional processes at broader
extents, corresponding to ecosystem function, but the coarse spatial
resolution removes the ability to relate these traits to individual
trees. Remote sensing data has been shown to be arguably the most
effective at monitoring ecosystem structure and function at regional to
global extents through the use of optical imagery [@cohen2004] and
active sensors such as lidar and radar datasets [@lefsky2002; @lang2021;
@neuenschwander2019].

Notably, while remote sensing can provide information on the
species-level EBVs, this information is typically acquired with an
*ad-hoc* approach requiring high spatial and spectral resolution data
[@skidmore2021]. This data can often only be collected at small extents,
rather than the global or regional scales required for biodiversity
trend assessment [@valdez2023]. The two landscape-level EBVs (ecosystem
structure and function) are well suited to be examined at global or
regional scales using moderate-resolution satellite imagery, such as
those provided by the Landsat or Sentinel series of satellites
[@skidmore2021].

Forest structural diversity has been linked to biodiversity at various
scales [@guo2017; @bergen2009; @gao2014]. Many metrics derived from
lidar remote sensing have been used as local indicators of biodiversity,
including simple metrics such as canopy cover and canopy height as well
as derived metrics including vertical profiles, aboveground biomass
[@lefsky1999]. Other second order derived metrics such as canopy
texture, height class distribution, edges, and patch metrics have also
been used to examine habitat and biodiversity at landscape scales
[@bergen2009]. Increased structural complexity has been hypothesized to
create additional niches, leading to increased species diversity
[@bergen2009], which has been frequently demonstrated using avian
species diversity metrics [@macarthur1961]. For example: @herniman2020
used spectral and lidar derived forest structure data to model avian
habitat suitability; @clawges2008 found that lidar derived forest
structural attributes are capable of identifying habitat types
associated with avian species in pine/aspen forests; @goetz2007 used
canopy structural diversity to predict bird species richness, finding
that canopy vertical distribution was the strongest predictor of species
richness. Forest structural metrics have also been used to study
biodiversity in other clades as well [@davies2014; @nelson2005].

With respect to ecosystem function, energy availability in an ecosystem
has shown to be a predictor of species richness and abundances at
various scales [@chase2002; @radeloff2019; @coops2019; @razenkova], and
is measurable using satellite remote sensing via the use of various
vegetation indices [@huete2002; @radeloff2019]. Vegetation indices,
which are indicative of photosynthetic activity, are commonly used as
proxies of gross primary productivity [@huang2019]. These vegetation
indices have also been used to assess patterns in biodiversity at single
time points [@bonn2004], and more recently, through yearly summaries of
productivity [@radeloff2019]. The relationship between energy
availability and biodiversity occurs via various hypotheses, such as the
available energy hypothesis, the environmental stress hypothesis, and
the environmental stability hypothesis. These three hypotheses have in
turn been linked to patterns of surface reflectance in remote sensing
data [@radeloff2019]. The Dynamic Habitat Indices (DHIs) have been shown
to be well suited to assess these hypotheses. The cumulative DHI
calculates the total amount of energy available in a given pixel over
the course of a year. Cumulative DHI is strongly linked to the available
energy hypothesis, which suggests that with greater available energy
species richness will increase [@wright1983]. The minimum DHI, which
calculates the lowest productivity over the course of a year can be
matched to the environmental stress hypothesis, which proposes that
higher levels of minimum available energy will lead to higher species
richness [@currie2004]. Finally, the variation DHI, which calculates the
coefficient of variance in a vegetation index through the course of a
year, corresponds to the environmental stability hypothesis which states
that lower energy variation throughout a year will lead to increased
species richness [@williams2008].

Within forested ecosystems, there are therefore many possible remotely
sensed derived ecosystem structure and function metrics that can be
assessed. Structural variables range in complexity from simple (canopy
cover; canopy height), to more complex (elevation coefficient of
variation; elevation standard deviation) to modelled (aboveground
biomass; basal area), all of which can be assessed using lidar data
[@coops2021]. Data on ecosystem functionality ranges from phenology
metrics (date of leaf on/leaf off; length of growing season) to
functional traits (chlorophyll content; leaf area index) to productivity
estimates (GPP; total available energy) [@pettorelli2018]. Advances in
satellite remote sensing processing have allowed 3D forest structure
data to be imputed across wide spatial scales [@matasci2018; @coops2021]
using data fusion approaches involving collected lidar data and
optical/radar data. Other advances in image compositing have allowed
yearly summaries of vegetation productivity to be calculated at regional
to global scales, summarizing the yearly total, minimum, and variation
in energy availability at similar spatial resolutions [@radeloff2019].

Linkages between forest structure and function have been examined within
a remote sensine context for over 20 years [@huete2002, @knyazikhin1998;
@myneni1994]. While there is significant theoretical and empirical
evidence for their relationship at single time points (within a single
image) [@myneni1994], various relationship directions and shapes have
been found between forest structure and function metrics [@ali2019].
These relationships, their shapes, and their strengths have been
attributed to multiple possible hypotheses and have been shown to vary
based on environmental conditions [@ali2019]. The relationship in
particular between forest structural diversity metrics - which are now
more accurately and comprehensively derived from lidar data - and
temporal variation in functional metrics, specifically the metrics of
ecosystem productivity via the DHI framework, have yet to be fully
examined.

Understanding the complementarity between potential EBVs is an integral
component of their creation [@skidmore2021]. Structural equation
modelling (SEM) and path analyses have been commonly used in ecology to
assess the causal effects behind various hypotheses [@fan2016;
@grace2010], including the relationship between forest structure,
functioning, and biodiversity [@ali2019]. For example, SEM has commonly
been used to assess the role of structural diversity of carbon storage
as compared to species diversity [@ali2016; @zhang2015].

The overall goal of this paper is to characterize patterns of forest
ecosystem structure and function and their interrelationships across a
wide range of ecosystems encompassing significant environmental
gradients. To do so, we synthesize data from moderate-scale
remote-sensing derived metrics of structure, represented as both simple
ALS metrics of canopy height and cover, as well as modeled forest
structure variables including volume, and above ground biomass, with a
remote sensing derived index on ecosystem function. Our first question
is to examine how ecosystem structure and function vary across a large
environmental gradient and then compare the simple and modeled
representations of forest structure to different levels of ecosystem
function. This question is important as it provides insights to the EBV
community around redundancy in remote sensing metrics when describing
the structure and function of ecosystems and proposes a method to
examine potential overlap when generating remote sensing EBV's.

Our second question examines the independent and shared relationships of
ecosystem structure height and cover, with modeled forest structure, on
function. This provides insight into the choice of remote sensing
attributes to use when developing EBV's. Remote sensing datasets can
comprise relatively unprocessed, observations, in this case ALS measures
of height and cover which are retrieval from the raw 3D point cloud, vs
modeled attributes, such as biomass and volume, which involve the use
the statistical relationships to transform the observations into more
refined data products. Assessing which of these two (or combination of
the two) approaches has more or less redundancy with estimates of
function provides insights into the choice of data used to build EBV's.
Lastly we examine how the primary and modeled structure attributes vary
in their variance partitioning within key biomes across this large
environmental range, examining to what extend ecosystem type impacts
these interrelationships and thus the applicability of these results
globally.

# Methods

## Study Area

British Columbia is the westernmost province of Canada, and is home to a
variety of terrestrial ecosystems [@pojar1987]. Approximately 64% of the
province is forested [@bcministryofforests2003]. There is a large amount
of ecosystem variation in the province, with large climate and
topographic gradients. The Biogeoclimatic Ecosystem Classification (BEC)
system identifies 16 zones based on the dominant tree species and the
ecosystems general climate. These zones can be further split into
subzones, variants, and phases based on microclimate, precipitation, and
topography [@pojar1987]. To examine trends across the large
environmental gradients, we group the BEC zones into five broad biomes,
specifically, the south interior, northern, montane, alpine, and coastal
groups similar to @hamann2006.

{insert figure showing ecosystems of interest} need to remake figure
showing ecosystems, highlighting that bunchgrass and alpine zones are
not included

{table showing elevations and climate ranges for each included bec zone}

## Data

### Forest Structure

We utilize a suite of forest structural diversity variables (canopy
height, canopy cover, Lorey's height, overstory cover, basal area,
aboveground biomass, gross stem volume, mean elevation, elevation
standard deviation, and structural complexity \[coefficient of variation
in elevation returns\]). This dataset is created at a 30 m spatial
resolution according to @matasci2018. In brief, the method utilizes a
set of lidar collections and field plots across Canada, and imputes the
remaining pixels using a random forest k-Nearest Neighbour approach.
Detailed information on the creation of this dataset can be found in
@matasci2018.

### Dynamic Habitat Indices

The Dynamic Habitat Indices are a set of satellite remote sensing
derived productivity variables that summarize the cumulative amount of
available energy, the minimum available energy, and the variation in
available energy throughout a given year [@radeloff2019]. The DHIs have
previously been produced at a global extent using MODIS imagery
[@radeloff2019], however, recent studies have examined how these indices
can be constructed at a finer spatial resolution using multi-annual
Landsat imagery.

The DHIs were calculated according to {RAZENKOVA LANDSAT DHI PAPER IN
PRESS} for all of terrestrial British Columbia. In brief, they utilized
Google Earth Engine [@gorelick2017] to obtain all valid Landsat pixels
for a given study area by filtering out low quality pixels, then
calculated the NDVI for each pixel. They use the median NDVI value for
each month across the ten year time span (2011-2020) to calculate the
sum, minimum, and coefficient of variation in NDVI values. More detailed
information can be found in {Razenkova In Press}.

## Sampling

We conducted model based sampling across the fifteen forest dominated
ecosystems found within British Columbia \[BC figure reference, I have
not made this figure yet\]. Samples were randomly selected within BEC
zones of interest alongside multiple criteria. Each sampled pixel had to
have a forested land cover class (coniferous, deciduous, mixed-wood, or
wetland-treed), and be surrounded by the same land cover class.
Additionally, each pixel had to have a coefficient of variation less
than 0.5 in the Lorey's height and canopy cover forest structural
metrics. A maximum of 3000 samples were sampled in each BEC zone. All
variables were natural-log transformed and standardized. Variables
containing zeros were natural-log plus one transformed.

## Analysis

### Partial Redundancy Analysis and Variation Partitioning

Partial redundancy analysis and variation partitioning was used to
relate the primary and modeled forest structure variables to ecosystem
structure, to assess the inter-relationships on forest ecosystem
function across the broad environment range. Redundancy analysis has
widely been used in community ecology where environmental variables of
interest are compared to species composition, and when the effect of
space for example is assessed as a third covariate, in an approach known
as partial redundancy analysis. While RDA is common in the ecological
literature, this represents one of the first times this technique has
been applied to assess redundancy in proposed essential biodiversity
variables. Redundancy analysis utilizes multiple regression in that it
models the effect of an explanatory matrix on a response matrix; with
the key difference being the use of a response matrix rather than a
single response variable. The first step is a multiple regression where
each of the variables in the response and explanatory matrices are
regressed resulting in a matrix of fitted values. PCA is then used in
the second step to reduce dimensionality through eigenvectors and to
allow the strength of relationships between assessed, taking into
account the number of explanatory variables. Partial Redundancy Analysis
(PRA) allows the use of additional explanatory variables, or
co-variates. This is useful, as it allow the assessment of the effect of
one set of variables on another while taking into account variations due
to other third set that are not the focus of the study. The output is
similar to standard RDA except it is also possible to examine how much
variation is explained in the primary variables of interest, and the
co-variates alone, and then when combined together. This variance
partitioning is then shown typically through a Venn diagram, in which
the percentage of explained variance by each explanatory data set is
reported. We use ANOVAs to test which canonical axes are significant,
and eigenvalues to determine the proprotion of variance explained by
each axis.

## Software

The processing for this paper was done in R version 4.2.2 [@R-base]. We
used *sgsR* [@goodbody2023] for sampling, *terra* [@R-terra] for raster
analyses, *sf* for vector analyses [@R-sf], and *vegan* [@R-vegan] for
redundancy analysis and variation partitioning.

Focal analyses were conducted in R and Python, and the sampling was
conducted across a masked suitability raster for each BEC zone using the
*sgsR* R package version 1.3.4 [@goodbody2023].

All code associated with this analysis is available at
https://github.com/emuise/code-structProdSem.

# Results

# Discussion

# Conclusion

\newpage

# References {.unnumbered}