### 2023-06-01 scripts/manuscript/manuscript.qmd

---
title: Disentangling linkages between satellite derived forest structure and productivity essential biodiversity variables.
author:
  - name: Evan R. Muise
    email: evan.muise@student.ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
    attributes:
        corresponding: true
  - name: Margaret E. Andrew
    email: M.Andrew@murdoch.edu.au
    affiliations: 
        - id: murd
          name: Murdoch University
          department: Environmental and Conservation Sciences and Harry Butler Institute
          city: Murdoch, WA, Australia
          postal-code: 6150
  - name: Nicholas C. Coops
    email: nicholas.coops@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Txomin Hermosilla
    email: txomin.hermosillagomez@NRCan-RNCan.gc.ca
    affiliations: 
        - id: cfs
          name: Natural Resources Canada
          department: Canada Forest Service (Pacific Forestry Centre)
          address: 506 Burnside Rd W
          city: Victoria, BC, Canada
          postal-code: V8Z 1M5
  - name: A. Cole Burton
    email: cole.burton@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Stephen S. Ban
    email: Stephen.Ban@gov.bc.ca
    affiliations: 
        - id: bcp
          name: Ministry of Environment and Climate Change Strategy
          department: BC Parks
          address: 525 Superior Street
          city: Victoria, BC, Canada
          postal-code: V8V 1T7
abstract: In today's changing climate and in light of the biodiversity crises, it is integral to be able to monitor the change in biodiversity at large scales. Key to this is the development of variables capable of monitoring biodiversity trends. One framework developed to tackle this problem is the Essential Biodviersity Variables (EBV), which is designed to be analogous to the Essential Climate variables, and capable of monitoring biodiversity globally while also being complementary to one another, within EBV classes. The EBV framework contains six classes, some of which are extremely well suited to be monitoring from satellite remote sensing. Among these with the potential to be satellite derived, the ecosystem structure and function classes stand out. In this paper, we seek to assess the complementarity of forest structure, as imputed across the entirety of British Columbia, Canada, with the Dynamic Habitat Indices, a yearly summary of productivity indices. Both datasets have been previously linked with biodiversity metrics across a range of scales. Using redundancy analysis, we find that forest structural attributes and the DHIs are decoupled from one another, with the forest structure datasets explaining 14% of the variation in the DHIs. Further, we explore how the proportion of variance explained in the DHIs from primary structural attributres such as canopy height and canopy cover and modelled structural attributes such as aboveground biomass varies. We find there is generally large overlap between the two forest structure datasets. Overall, we find that forest structure as an ecosystem structure EBV and the DHIs as an ecosystem function EBV are well suited to be complementary EBVs across large environmental gradients.
keywords: 
  - remote sensing
  - landsat
  - forest structure
  - forest productivity
  - dynamic habitat indices
  - essential biodiversity variables
  - redundancy analysis
date: last-modified
bibliography: [bibliography.bib, packages.bib]
csl: remote-sensing-of-environment.csl
format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    journal:
      name: Remote Sensing of Environment
      formatting: review
      model: 3p
      cite-style: authoryear
      layout: twocolumn
  docx: 
    reference-doc: my-styles.docx
    keep-md: true
  pdf: 
    pdf-engine: xelatex
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: false
fig-width: 300
---

```{r}
#| warning: false
#| echo: false
#| output: false

# include all libraries required by the R chunks here
library(vegan)
library(sgsR)
library(terra)
library(sf)
library(gt)
library(tidyverse)

library(bcmaps)

# generate package citations
loaded_pkgs = (.packages())
base_pkgs = getOption("defaultPackages")
contrib_pkgs = loaded_pkgs[!(loaded_pkgs %in% base_pkgs)]
pkgs_file = here::here("scripts", "manuscript", "packages.bib")
knitr::write_bib(contrib_pkgs, pkgs_file)

bec_tab <- read_csv(here::here("data", "climate.csv")) %>%
  mutate(across(.cols = c(tmax, tmin), \(x) round(x, digits = 1)))
```

# Introduction

With biodiversity being in decline, and facing extinction rates above
the background extinction rate [@thomas2004; @urban2015], as well as the
homogenization of communities at various scales [@mcgill2015] it is
integral to be able to monitor how biodiversity is changing across the
globe. In response, the global biodiversity community is making efforts
to assess and halt the degradation of biodiversity. The Group for Earth
Observation Biodiversity Observation Network has developed the Essential
Biodiversity Variables [EBVs, @pereira2013], designed as an analog to
the Essential Climate Variables framework [@bojinski2014]. EBVs are
designed to be global in scope, relevant to biodiversity information,
feasible to use, and complementary to one another [@skidmore2021]. While
it can be incredibly difficult, time consuming, and expensive to collect
data on biodiversity across wide swaths of land and varying ecosystems,
EBVs, which can be correlated to sampled biodiversity information, allow
for the monitoring and assessment of protected area effectiveness and
ecosystem health at large spatial scales [@hansen2021]. There are six
EBV classes, each of which correspond to a different facet of
biodiversity including species populations, species traits, community
composition, ecosystem structure, ecosystem function, and genetic
composition [@pereira2013].

Satellite remote sensing has proven to be capable of measuring five of
the six EBV classes, the exception being genetic composition, which
requires in-situ observation and sampling [@skidmore2021]. Species
populations - and in turn community composition - can be assessed with
very-high-resolution imagery to identify tree species at the tree-crown
scale, however it is difficult and computationally expensive to extend
these analyses to broader extents [@fassnacht2016; @graves2016] while
species traits such as vegetation phenology have been observed at the
single-tree scale using, for example, PlanetScope imagery and
drone-based measurements [@wu2021]. However, the spatially limited and
often *ad-hoc* data collection approaches associated with monitoring
individuals is not conducive to the global or regional scales required
for biodiversity trend assessment [@valdez2023].

The two landscape-level EBVs (ecosystem structure and function) are well
suited to be examined at large spatial scales using coarser spatial
measurements, such as those taken from satellites by the Moderate
Resolution Imaging Spectroradiometer [MODIS; @zhang2003], the Landsat
imaging systems [@fisher2006], or Sentinel-2 [@helfenstein2022;
@darvishzadeh2019] programs. These mid-resolution satellites can monitor
processes at broader extents but the coarse spatial resolution removes
the ability to relate these traits to individual organisms. As a result,
satellite remote sensing data has been shown to be arguably the most
effective at monitoring ecosystem based EBVs focused on structure and
function. These EBVs classes can be monitored at regional to global
extents through the use of optical imagery [@cohen2004], as well as
active sensors such as lidar (light detection and ranging) and radar
[@guo2017; @lefsky2002; @lang2021; @neuenschwander2019; @coops2016].

## EBV - Ecosystem Structure

Forest structural diversity has been linked to biodiversity at various
scales [@guo2017; @bergen2009; @gao2014]. Structural attributes range in
complexity from simple (canopy cover; canopy height), to more complex
(vertical and horizontal structural complexity) to modelled (aboveground
biomass; basal area), all of which can be assessed using lidar data
[@coops2021]. A suite of these lidar-derived attributes have been used
as local indicators of biodiversity, including simple metrics such as
canopy cover and canopy height as well as derived metrics including
vertical profiles, aboveground biomass [@lefsky1999; @guo2017;
@coops2016]. Other second order derived metrics such as canopy texture,
height class distribution, edges, and patch metrics have also been used
to examine habitat and biodiversity at landscape scales [@bergen2009].
Advances in satellite remote sensing processing have allowed 3D forest
structure data to be imputed across wide spatial scales [@matasci2018;
@coops2021] using data fusion approaches involving collected lidar data
and optical/radar data.

Increased forest structural complexity has been hypothesized to create
additional niches, leading to increased species diversity [@bergen2009],
which has been frequently demonstrated using avian species diversity
metrics [@macarthur1961]. For example: @herniman2020 used spectral and
lidar derived forest structure data to model avian habitat suitability;
@clawges2008 found that lidar derived forest structural attributes are
capable of identifying habitat types associated with avian species in
pine/aspen forests; @goetz2007 used canopy structural diversity to
predict bird species richness, finding that canopy vertical distribution
was the strongest predictor of species richness. Forest structural
metrics have also been used to study biodiversity in other clades as
well [@davies2014; @nelson2005].

## EBV - Ecosystem Function

With respect to ecosystem function, energy availability in an ecosystem
has shown to be a predictor of species richness and abundances at
various scales [@chase2002; @radeloff2019; @coops2019; @razenkova2023],
and is measurable using satellite remote sensing via the use of various
vegetation indices [@huete2002; @radeloff2019]. Vegetation indices,
which are indicative of photosynthetic activity, are commonly used as
proxies of gross primary productivity [@huang2019]. These vegetation
indices have also been used to assess patterns in biodiversity at single
time points [@bonn2004], and more recently, through yearly summaries of
productivity [@berry2007; @radeloff2019]. The relationship between
energy availability and biodiversity occurs via various hypothesized
mechanisms, such as the available energy hypothesis [@currie2004;
@wright1983], the environmental stress hypothesis [@currie2004], and the
environmental stability hypothesis [@williams2008]. These three
hypotheses have in turn been linked to patterns of annual surface
reflectance in remote sensing data [@berry2007; @radeloff2019].

@berry2007 first explored this idea by proposing the linkage of
intra-annual summaries of MODIS derived GPP to dispersive bird species.
This idea was further refined into the Dynamic Habitat Indices [DHIs;
@coops2008], which have now been shown to be well suited to assess the
three aforementioned hypotheses at global scales [@radeloff2019]. The
cumulative DHI calculates the total amount of energy available in a
given pixel over the course of a year. Cumulative DHI is strongly linked
to the available energy hypothesis, which suggests that with greater
available energy species richness will increase [@wright1983]. The
minimum DHI, which calculates the lowest productivity over the course of
a year can be matched to the environmental stress hypothesis, which
proposes that higher levels of minimum available energy will lead to
higher species richness [@currie2004]. Finally, the variation DHI, which
calculates the coefficient of variance in a vegetation index through the
course of a year, corresponds to the environmental stability hypothesis
which states that lower energy variation throughout a year will lead to
increased species richness [@williams2008].

## Biodiversity Monitoring with EBVs

Biodiversity monitoring programs often require a range of information in
order to accurately assess changes in ecosystem integrity
[@lindenmayer2010]. Choosing datasets that are most closely related to
the phenomenon of interest in a given application allows for direct
connections to monitoring results and management actions [@pressey2021].
With the advent of large-extent monitoring methods like satellite remote
sensing, and a proliferation of potential EBVs datasets, it becomes
important to assess the interrelationships between these datasets, and
assess their complementary of the information to reduce duplication of
efforts [@pereira2013; @skidmore2021]. When strong relationships are
present between EBV classes, it becomes possible to assess the
ecological relationships between potential EBVs. On the other hand, when
datasets do not appear related, they may be well suited to be used in
monitoring programs together, as complementary EBVs.

Linkages between forest ecosystem structure and function have been
examined within a remote sensing context for over 20 years [@huete2002,
@knyazikhin1998; @myneni1994]. While there is significant theoretical
and empirical evidence for their relationship at single time points
(within a single image) [@myneni1994], various relationship directions
and shapes have been found between forest structure and function metrics
[@ali2019]. Hypothesized mechanisms such as niche complementary have
shown that aboveground biomass increases with stand structure
[@zhang2012], while asymmetric competition for light can reduce forest
productivity with increased structural complexity [@bourdier2016]. The
relationship in particular between forest structural diversity metrics -
which are now more accurately and comprehensively derived from lidar
data - and temporal variation in functional metrics, specifically the
metrics of ecosystem productivity via the DHI framework, have yet to be
fully examined.

The overall goal of this paper is to assess patterns of forest ecosystem
structure and function and their complementarity across a wide range of
ecosystems encompassing significant environmental gradients. To do so,
we synthesize data from moderate-scale remote-sensing derived metrics of
forest structure, represented as both simple ALS-extracted metrics of
canopy height, cover and vertical complexity, as well as modelled forest
structure attributes including volume, and aboveground biomass, with a
well-established remote sensing derived index on ecosystem function. Our
first question is to examine how ecosystem structure and function
complement one another across a large environmental gradient and then
compare the simple and modelled representations of forest structure to
different levels of ecosystem function. This question is important as it
provides insights to the EBV community around complementarity of remote
sensing metrics when describing the structure and function of ecosystems
and proposes a method to examine potential overlap when generating
remote sensing EBVs.

Our second question examines the independent and shared relationships of
ecosystem structure height and cover, with modelled forest structure, on
ecosystem function. This provides insight into the choice of remote
sensing attributes to use when developing EBVs within a single EBV
class. Remote sensing datasets can comprise relatively unprocessed
observations, in this case ALS measures of height and cover which are
derived from the raw 3D point cloud vs modelled attributes, such as
biomass and volume, which involve the use the statistical relationships
with field data to transform the observations into more refined data
products. Assessing which of these two (or combination of the two)
approaches has stronger or weaker correlations with estimates of
function provides insights into the choice of data used to build EBVs.
Lastly, we examine how the primary and modelled structure attributes
partition the variance of the DHIs within key biomes and forest types
across a large environmental range, examining to what extent ecosystem
and forest types impacts these relationships and thus providing insight
into the applicability of these results globally.

# Methods

## Study Area {#sec-study-area}

British Columbia is the westernmost province of Canada, and is home to a
variety of terrestrial ecosystems. Approximately 64% of the province is
forested, with large environmental and topographic gradients
[@pojar1987; @bcministryofforests2003]. The Biogeoclimatic Ecosystem
Classification (BEC) system identifies 16 zones based on the dominant
tree species and the ecosystems general climate (@fig-study-map).

```{r}
#| label: fig-study-map
#| fig-cap: Location of biogeoclimatic ecosystem classification (BEC) zones in British Columbia.

knitr::include_graphics(here::here("outputs", "bec_map.png"))
```

These zones can be further split into subzones, variants, and phases
based on microclimate, precipitation, and topography [@pojar1987]. To
examine trends across the large environmental gradients, we group the
BEC zones into five broad biomes, specifically, the southern interior,
northern interior, montane, alpine, and coastal groups similar to
@hamann2006. We also report each BEC zone's average climate data,
according to @wang2020 (@tbl-bec-group).

```{r}
#| label: tbl-bec-group
#| tbl-cap: BEC Zones, their aggregate groups, and their average climate values for precipitation, maximum temperature, and minimum temperature. Climate data from Wang et al. 2016.

# bec_tab %>%
#   gt() %>%
#   cols_label(
#     group = "BEC Group",
#     zone = "BEC Zone",
#     zone_nm = "Full Name",
#     ppt = "Precipitation (mm/yr)",
#     tmax = "Max Temperature (°C)",
#     tmin = "Min Temperature (°C)"
#   ) %>%
#   as_word()

knitr::kable(
  bec_tab,
  col.names = c(
    "BEC Group",
    "BEC Zone",
    "Full Name",
    "Precipitation (mm/yr)",
    "Max Temperature (°C)",
    "Min Temperature (°C)"
  )
)
```

## Data

### Forest Structure

We used a suite of forest structural attributes (canopy height, canopy
cover, overstory cover, basal area, aboveground biomass, gross stem
volume, mean elevation, elevation standard deviation, and structural
complexity \[coefficient of variation in elevation returns\]). This
dataset was created for the year 2015 at a 30 m spatial resolution
according to @matasci2018. In brief, the method used a set of lidar
collections and field plots across Canada, and imputed the remaining
pixels using a random forest k-Nearest Neighbour approach on
Landsat-derived surface reflectance and auxiliary data such as
topography. Detailed information on the creation of this dataset can be
found in @matasci2018.

### Dynamic Habitat Indices

We use an established set of indices of annual productivity shown to be
related to global biodvierstiy trends: the Dynamic Habitat Indices
[@radeloff2019]. The DHIs are a set of satellite remote sensing derived
productivity variables that summarize the cumulative amount of available
energy, the minimum available energy, and the variation in available
energy throughout a given year [@berry2007; @radeloff2019]. The DHIs
have previously been produced at a global extent using MODIS imagery ,
and have been used to assess alpha [@radeloff2019] and beta[@andrew2012]
diversity, species abundances [@razenkova2023], and construct novel
ecoregionalizations [@coops2009]; @andrew2013\] . Recent studies have
began to examine how these indices can be constructed at a finer spatial
resolution by using multi-annual Landsat imagery to generate a single
synthetic year of monthly observations [@razenkova2022].

The DHIs were calculated according to [@razenkova2022; Razenkova et al.,
In Press] for all of terrestrial British Columbia. In brief, Google
Earth Engine [@gorelick2017] was used to obtain all valid Landsat pixels
for a given study area, filtering out pixels containing shadows, slouds,
and cloud shadows within each image [@zhu2012], then calculated the NDVI
for each pixel in each image. They then calculated the median NDVI value
for each month across the ten year time span (2011-2020) to generate a
synthetic year of monthly data. The sum, minimum, and coefficient of
variation across this synthetic year of NDVI values is then calculated.
More detailed information can be found in Razenkova et al. (In Press).

## Sampling

We conducted model-based sampling across the sixteen ecosystems found
within British Columbia (@fig-study-map). Samples were randomly selected
within each BEC zone, in undisturbed pixels. Each sampled pixel had to
have a forested land cover class (coniferous, deciduous, mixed-wood, or
wetland-treed), and be surrounded by the same land cover class. The land
cover mask was generated following [@hermosilla2022] using a
best-available-pixel composite, and an inverse-distance weighted random
forest approach across Canada. Additionally, each pixel had to have a
coefficient of variation less than 0.5 in surrounding pixels in the two
simplest forest structural attributes, canopy height and canopy cover. A
maximum of 3000 samples were sampled in each BEC zone with a 1 km
minimum sampling distance to reduce the effects of spatial
autocorrelation. All variables were natural-log transformed and
standardized. Variables containing zeros were natural-log plus one
transformed. Sampling was conducted in R [@R-base] version 4.2.2 using
the **sgsR** package [@R-sgsR]. Focal analyses for the land cover
classes and coefficient of variation were calculated in Python version
3.9.

# Sampling 2

We implement two sampling schemes -- one for ecosystems, and one for
forest types -- across the terrestrial ecosystems of British Columbia
(@fig-study-map). Both have samples randomly selected. For both sampling
schemes, each individual sample was taken from a forested pixel which
was surrounded by the same land cover class. The land cover class map
was generated for the year 2015 following [@hermosilla2022] using a
best-available-pixel composite, and an inverse-distance weighted random
forest approach across Canada. Additionally, each pixel had to have a
coefficient of variation less than 0.5 in surrounding pixels in the two
simplest forest structural attributes, canopy height and canopy cover. A
minimum sampling distance of 1 km was implemented to reduce the effects
of spatial autocorrelation, and samples that had been disturbed in the
last 30 years were discarded by using a disturbance mask generated for
the forested ecosystems of Canada by @hermosilla2016.

The ecosystem sampling scheme was implemented by randomly sampling
pixels matching the above criteria within each BEC zone of British
Columbia, up to a maximum of 3000 pixels, regardless of land cover
class. The forest type sampling scheme was similar in that each BEC zone
was sampled, however, it implemented stratified sampling on each forest
type, up to a maximum of 500 pixels per BEC zone and forest type.

Each sampling scheme was natural-log transformed and standardized to
Z-scores. Variables containing zeros were natural-log plus one
transformed. Sampling was conducted in R [@R-base] version 4.2.2 using
the **sgsR** package [@R-sgsR]. Focal analyses for the land cover
classes and coefficient of variations of canopy height and cover were
calculated in Python version 3.9.

## Analysis

### Redundancy Analysis and Variation Partitioning

Redundancy analysis (RDA) and variation partitioning were used to relate
the primary and modelled forest structure attributes to ecosystem
function across a broad environment range. Redundancy analysis functions
similarly to a multiple linear regression, except it is capable of
predicting multiple response variables. It accomplishes this by first
running a multiple linear regression of each predictor variable on each
response variable, then running a principle component analysis on the
residuals from each multiple linear regression. This reduces the
dimensionality of the output, and allows the relationship strength to be
assessed by calculating the loadings of both predictor and response
variables on the RDA axes. Partial redundancy analysis function
similarly, except also considers co-variates [@legendre2012]. Redundancy
analysis has widely been used in community ecology where environmental
variables of interest are compared to species composition
[@blanchet2014; @kleyer2012] and has similarities to partial least
squares regression -- namely, their multivariate approach, usage of
dimensionality reduction, and linearity assumptions -- which is commonly
used in remote sensing literature [@roelofsen2014; @burnett2021].

Following the RDA, we employ ANOVAs to determine which axes are
significant, and calculate the proportion of variance attributable to
each axis using the eigenvalues generated from the RDA. We calculate
axis loadings for both predictor and response variables by calculating
the correlation between the variables and the RDA axes. Axis loadings
represent the relationship between a given variable and the RDA axis. We
only consider and display significant axes. To visualize the RDA for
both predictor and response variables, we display the results as path
diagrams, with loadings from each predictor to the RDA axis to the
response variables. The variance explained by each axis is also
displayed in the RDA box (@fig-rda-var). All RDA calculations were done
in R [@R-base] version 4.2.2 using the **vegan** package [@R-vegan].

Variation partitioning is an extension of partial RDA which can assess
the overlap between the explanatory power of two datasets by utilizing
multiple partial RDAs and exchanging which datasets are considered the
predictor, and which is considered the co-variate [@legendre2012].
Variation partitioning is traditionally displayed using a Venn diagram,
in which the percentage of variance explained by each dataset is in a
circle, and the overlap between circles represents the overlap in
variance explained. All variation partition analyses were done in R
[@R-base] version 4.2.2 using the **vegan** package [@R-vegan].

RDA and variation partitioning analyses were conducted for all samples
in the ecosystem sampling scheme, as well as individually across each
ecosystem and forest type. The results were aggregated to BEC zone
groups (see @sec-study-area).

All code associated with the processing and analysis is available at
https://github.com/emuise/code-structProdSem.

# Results

```{r}
#| label: fig-rda-var
#| fig-cap: A) Axis loadings from redundancy analysis of primary and modelled forest structural attributes on the dynamic habitat indices. B) Results from variation partitioning of primary and modelled forest structural attributes on the DHIs. Both visualized analyses are across all collected samples. See supplementary information for results from each BEC zone and forest type.

knitr::include_graphics(here::here("outputs", "all_data_figure.drawio.png"))
```

To examine the relationship between ecosystem structure and function
across a large environmental gradient, we present the results of a
redundancy analysis of forest structural attributes on the dynamic
habitat indices across the entire sampled dataset in @fig-rda-var A.
While there are three RDA axes associated with the full dataset, the
third axis explains 0.05% of the variance in the DHIs, and as a result
it is not shown. The first axis is strongly represents all the DHIs
(loadings \> 0.85 for all DHIs), and has the strongest loadings from
canopy cover, basal area, aboveground biomass, and gross stem volume.
The other input attributes (canopy height, structural complexity) have
smaller loadings. The second axis primarily represents the seasonality
(Minimum and Variation DHI) of the DHIs, with a very small loading on
the Cumulative DHI, and has the largest loadings in canopy cover and
complexity, with the rest of the forest structural attributes having
negligible loadings (@fig-rda-var A).

The results from the variation partitioning analysis (@fig-rda-var B)
show that the majority of the variance explained by the input datasets
is shared across both primary and modelled attributes. In total, only
13.4% of the variation in the DHIs is explained by the structural
information, with 9.4% of this being from overlap between the datasets.
The primary and modelled attributes explain 2.7 and 1.3% of the
variation on their own, respectively (@fig-rda-var B).

```{r}
#| label: fig-radar
#| fig-cap: Radar plots of average loadings strength by group. A and B show input and response loadings, respectively. C) Boxplots of BEC zone, forest types, and all data loadings for predictor and response variables.

knitr::include_graphics(here::here("outputs", "radar_boxplot.png"))
```

Further, we wanted to examine the strength of the relationship between
the DHIs and primary structural attributes vs modelled structural
attributes across the forest types and ecosystems of the province.
@fig-radar A and B show the axis loadings for structural attributes and
the DHIs, respectively. Across the BEC zone groups, the loadings are
generally similar in the structural attributes. Southern ecosystems have
low or non-significant structural complexity loadings, while the forest
types have structural complexity as their largest loading. The loadings
for the DHIs in the first axis are generally larger than the loadings
for the structural attributes, with the cumulative DHI often being the
strongest loading overall. In alpine BEC zones, the minimum DHI loading
is smaller than the variation and cumulative DHIs. In contrast to the
first RDA axis, the secondary RDA axis is primarily driven by variation
in the minimum DHI, with medium loadings in the variation DHI, and small
loadings in the cumulative DHI (@fig-radar A & B). @fig-radar C shows
boxplots of individual BEC zone loadings for structural attributes and
the DHIs. Across the BEC zones, canopy cover is generally the strongest
predictor loading of the first RDA axis, and structural complexity is
generally the weakest predictor loading. The exception is in the forest
types, where structural complexity is often the largest driver of the
first axis. When significant, the second RDA axis always has small
structural attribute loadings.

```{r}
#| label: fig-fracts-bar
#| fig-cap: Stacked bar plot proportion of DHI variation explained by extracted, modelled, and the overlap between primary and modelled structural attributes.

knitr::include_graphics(here::here("outputs", "fracts_bar_plot.png"))
```

@fig-fracts-bar shows stacked bar plots of the proportion of variation
explained in the DHIs by primary, modelled, and the overlap between
primary and modelled attributes. The forest types sampling scheme
generally results in higher amounts of variance explained, which is
especially prevalent in broadleaf and mixed wood forests. The overlap
between modelled and primary attributes is generally higher than each
dataset on their own, with the exception being mixed wood and broadleaf
forests and the Coastal Western Hemlock and Engellmann Spruce - -
Subalpine Fir BEC zone zones. The Interior Mountain-heather Alpine and
Montane Spruce ecosystems have larger amounts of DHI variance explained
than the rest of their ecosystem groups. Notably, the variation
explained by the primary and modelled attributes is commonly presented
through a single RDA axis, which generally corresponds to overall
productivity through the year (@fig-radar). Generally, the DHIs are
decoupled from the structural attributes, as shown by the overall
variance explained being under 30%, regardless of whether the attribute
is directly extracted from the point cloud or is a modelled variable.

```{r}
#| label: fig-fcc
#| fig-cap: False colour maps of axis loadings for the first RDA axis (top) and second RDA axis (bottom). All colour values are normalized to the maximum loading of each variable. A and D show axis loadings for canopy height (r), canopy cover (g) and structural complexity (b). B and E show axis loadings for basal area(r), total biomass (g) and gross stem volume (b). C and F show axis loadings for the cumulative DHI (r), variation DHI (g) and minimum DHI (b).

knitr::include_graphics(here::here("outputs", "rgb_norm_edited.png"))
```

@fig-fcc shows false colour composites of the primary (A & C) and
modelled (B & D) loadings across the ecosystems of British Columbia. In
the first RDA axis there is spatial variation in the primary attributes,
with interior zones being primarily driven by canopy cover (green), the
coastal regions having strong structural complexity and canopy height
loadings, and the boreal in the northwest having the strongest loading
in canopy height (@fig-fcc A). The modelled attributes generally show
gray-scale colour, indicating that basal area, total biomass, and gross
stem volume explain similar variation in the DHIs (@fig-fcc B). Interior
Cedar -- Hemlock, Boreal White and Black Spruce, Interior Douglas-fir,
Mountain Hemlock, Montane Spruce and Sub-Boreal Spruce have a secondary
axis, shown in @fig-radar C & D. All zones with a secondary axis are
located away from the Pacific coast. Canopy cover has the strongest
secondary axis loading in the boreal white and black spruce, while the
secondary axis in the remaining four zones have the highest loadings in
canopy height (@fig-fcc C). Again, the modeled attributes are in
grayscale, indicating similar loadings across the three modelled
attributes (@fig-fcc E).

# Discussion

Complementarity plays a crucial role in the context of Essential
Biodiversity Variables (EBVs) and holds significant importance for
monitoring and understanding biodiversity [@pereira2013; @skidmore2021].
At the ecosystem scale, the interconnectedness and interdependance of
ecosystem structure and function is important to recognize and
disentangle. In forested environments, forest structure and productivity
have been shown to be linked in multiple studies spanning three decades
[@ali2019; @myneni1994], however, a linkage between intra-annual
production and forest structure has yet to be shown. In this study, we
use statistical analyses from community ecology - namely redundancy
analysis and variation partitioning - to assess the complementarity of
forest structure and yearly productivity summaries. While RDA is common
in the ecological literature, this analysis represents one of the first
times this technique has been applied to assess the complimentary of
proposed satellite-derived EBVs. We find that the datasets do not
strongly overlap, with forest structure explaining 13.4% of the
variation in the DHIs in samples taken across the entirety of British
Columbia in the ecosystem sampling scheme (@fig-rda-var B). This
indicates that they are suitable to be used in tandem with one another
when used as ecosystem EBVs.

Across most of British Columbia's ecosystem, we identified a single RDA
axis associated with the DHIs, encompassing the variation in yearly
productivity (@fig-rda-var A). Within this first axis, the strongest
loadings were canopy cover and the modelled attributes. In Alpine
ecosystems the relationship between the first axis and the Minimum DHI
was lowest, likely due to the low amount of variation in the Minimum DHI
in these ecosystems. When a second axis was significant, it consistently
had strong loadings on the minimum and variation DHIs, which indicates a
complex internal productivity relationship in certain ecosystems. This
secondary axis has smaller axis loadings associated with the primary and
modelled attributes, with the strongest loadings being canopy cover and
structural complexity across the entire dataset (@fig-rda-var A).

Our second question sought to explore whether modelled attributes (basal
area, gross stem volume, aboveground biomass) add additional explanatory
information when predicting the DHIs, as compared to primary forest
structural attributes. We generally found that canopy cover had the
largest axis loadings in the first RDA axis (@fig-radar). Modelled
attributes such as basal area, aboveground biomass, and gross stem
volume shared similar loading magnitudes across the range of studied
ecosystems, indicating they do not add additional value when utilized
with one another (@fig-radar C; @fig-fcc B). The loadings between the
modelled attributes and canopy cover are often similar, and as such, we
recommend utilizing the attributes derived directly from the point cloud
in the case of ALS data, or selecting a single modelled attribute.

During the analysis we found a high amount of total variance explained
in deciduous forests (mixed-wood and broadleaf) when compared to the
other two forest types and most BEC zones (@fig-fracts-bar). This could
potentially be due to the temporal linkages between canopy cover in
deciduous forests and productivity. In evergreen coniferous forests
where canopy cover is present year-round, there is likely a temporal
mismatch between forest structural attributes and the DHIs. Further, the
strongest loadings in these two forest types was vertical structural
complexity, rather than canopy cover. This indicates that the DHIs are
more closely related to forest structure in deciduous forests, however
even there they are still decoupled from one another.

Finally, we explored the amount of variation explained by primary vs
modelled attributes, as well as their overlap (@fig-fracts-bar). We
generally found that the overlap between the primary and modelled
attributes explained most of the variation, with some exceptions,
indicating that using either set of forest structural attributes may be
suitable when monitoring biodiversity. The exceptions are especially
prominent in the broadleaf and mixed wood forest classes, and the
Coastal Western Hemlock ecosystem, where there is no variation explained
by the overlap between the two datasets. The majority of the variance
explained in the broadleaf and mixed wood forests was in the forest
structural attributes derived directly from the point cloud. This was
expected, as the modelled attributes are calculated based on the derived
attributes as well as site index and ancillary information [@coops2021].

Recent advances in creating synthetic yearly observations have allowed
the DHIs to be generated at a finer scale (30 m; Razenkova et al., In
Press), rather than the previously used 1 km DHIs derived from MODIS
\[@radeloff2019; @razenkova2022\]. This represents a significant
advancement when assessing the utility of EBVs, as the 30 m scale is
well suited to examine a range of ecological applications, including
forest structure and productivity, and more closely approximates the
scale at which species perceive habitat [@cohen2004; @kennedy2014]. In
addition to more closely matching the scale of ecological phenomenon,
this allows these datasets to be matched and analyzed with other
datasets generated from the Landsat archive.

In conclusion, we used redundancy analysis and variation partitioning to
assess the complementarity of two potential EBV datasets - forest
structure and the DHIs. We also separated the forest structure datasets
into primary and modelled attributes in order to assess the need to
develop more complex structural attributes, or if data derived directly
from lidar datasets was suitable. We found that the structural
attributes are not strongly related to the DHIs, indicating that they
are suitable to be used together as ecosystem scale EBVs when monitoring
forest environments. We also found that variation explained by the
overlap between primary and modelled attributes was often higher than
the variation explained by either individually.

\newpage

# References {.unnumbered}