### 2023-06-05 scripts/manuscript/manuscript.qmd
full draft to tx & mea

---
title: Disentangling linkages between satellite derived forest structure and productivity essential biodiversity variables.
author:
  - name: Evan R. Muise
    email: evan.muise@student.ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
    attributes:
        corresponding: true
  - name: Margaret E. Andrew
    email: M.Andrew@murdoch.edu.au
    affiliations: 
        - id: murd
          name: Murdoch University
          department: Environmental and Conservation Sciences and Harry Butler Institute
          city: Murdoch, WA, Australia
          postal-code: 6150
  - name: Nicholas C. Coops
    email: nicholas.coops@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Txomin Hermosilla
    email: txomin.hermosillagomez@NRCan-RNCan.gc.ca
    affiliations: 
        - id: cfs
          name: Natural Resources Canada
          department: Canada Forest Service (Pacific Forestry Centre)
          address: 506 Burnside Rd W
          city: Victoria, BC, Canada
          postal-code: V8Z 1M5
  - name: A. Cole Burton
    email: cole.burton@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Stephen S. Ban
    email: Stephen.Ban@gov.bc.ca
    affiliations: 
        - id: bcp
          name: Ministry of Environment and Climate Change Strategy
          department: BC Parks
          address: 525 Superior Street
          city: Victoria, BC, Canada
          postal-code: V8V 1T7
abstract: Under a changing climate and in light of the biodiversity crises, monitoring change in biodiversity at broad scales has become ever more imperative. Key to regular biodiversity monitoring over time and space is the definition and selection of variables capable of accurately monitoring biodiversity trends. One framework is the Essential Biodiversity Variables (EBV), which are designed to be analogous to the Essential Climate variables, and capable of monitoring biodiversity globally while also being complementary to one another, within EBV classes. Five of six EBV classes have been shown to be well suited to be monitored using satellite remote sensing, with particular focus warranted on the ecosystem structure and function classes. In this paper, we seek to address the complementarity of variables describing forest structure, as imputed across the entirety of British Columbia, Canada, with variables representing ecosystem functioning, namely the Dynamic Habitat Indices, a yearly summary of productivity indices. Both sets of variables have previously been linked with biodiversity metrics across a range of scales. Using redundancy analysis, we find that forest structural attributes and the DHIs are essentially decoupled from each other, with the forest structure variables explaining 14% of the variation in the DHIs. Further, we explore how the proportion of variance explained in the DHIs varies with either primary structure variables, such as canopy height and cover, vs more complex modelled forest structural attributes, such as aboveground biomass and volume. Results suggest there is large overlap between the two types of forest structure variables suggesting little added benefit of derived attributes beyond height and cover, in this context. Overall, we find that forest structure as an ecosystem structure EBV and the DHIs as an ecosystem function EBV represent markedly different processes as observed by remote sensing and they are highly complementary across large environmental gradients, suggesting that in an earth-observation based monitoring program both are needed.
keywords: 
  - remote sensing
  - landsat
  - forest structure
  - forest productivity
  - dynamic habitat indices
  - essential biodiversity variables
  - redundancy analysis
date: last-modified
bibliography: [bibliography.bib, packages.bib]
csl: remote-sensing-of-environment.csl
format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    journal:
      name: Remote Sensing of Environment
      formatting: review
      model: 3p
      cite-style: authoryear
      layout: twocolumn
  docx: 
    reference-doc: my-styles.docx
    keep-md: true
  pdf: 
    pdf-engine: xelatex
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: false
fig-width: 300
---

```{r}
#| warning: false
#| echo: false
#| output: false

# include all libraries required by the R chunks here
library(vegan)
library(sgsR)
library(terra)
library(sf)
library(gt)
library(tidyverse)

library(bcmaps)

# generate package citations
loaded_pkgs = (.packages())
base_pkgs = getOption("defaultPackages")
contrib_pkgs = loaded_pkgs[!(loaded_pkgs %in% base_pkgs)]
pkgs_file = here::here("scripts", "manuscript", "packages.bib")
knitr::write_bib(contrib_pkgs, pkgs_file)

bec_tab <- read_csv(here::here("data", "climate.csv")) %>%
  mutate(across(.cols = c(tmax, tmin), \(x) round(x, digits = 1)))
```

# Introduction

With biodiversity being in decline, and facing extinction rates above
the background extinction rate [@thomas2004; @urban2015], as well as the
homogenization of communities at various scales [@mcgill2015] it is
integral to be able to monitor how biodiversity is changing across the
globe. In response, the global biodiversity community is making efforts
to assess and halt the degradation of biodiversity. The Group for Earth
Observation Biodiversity Observation Network has developed the Essential
Biodiversity Variables [EBVs, @pereira2013], designed as an analog to
the Essential Climate Variables framework [@bojinski2014]. EBVs are
designed to be global in scope, relevant to biodiversity information,
feasible to use, and complementary to one another [@skidmore2021]. While
it can be difficult, time consuming, and expensive to collect data on
biodiversity across wide swaths of land and varying ecosystems, EBVs,
which can be correlated to sampled biodiversity information, allow for
the monitoring and assessment of protected area effectiveness and
ecosystem health at large spatial scales [@hansen2021]. There are six
EBV classes, each of which correspond to a different facet of
biodiversity, including species populations, species traits, community
composition, ecosystem structure, ecosystem function, and genetic
composition [@pereira2013].

Satellite remote sensing has proven to be capable of measuring five of
the six EBV classes, the exception being genetic composition, which
requires in-situ observation and sampling [@skidmore2021]. Species
populations - and in turn community composition - can be assessed with
very-high-resolution imagery to identify tree species at the tree-crown
scale, however it is difficult and computationally expensive to extend
these analyses to broader extents [@fassnacht2016; @graves2016] while
species traits such as vegetation phenology have been observed at the
single-tree scale using, for example, PlanetScope imagery and
drone-based measurements [@wu2021]. However, the spatially limited and
often *ad-hoc* data collection approaches associated with monitoring
individuals is not conducive to the global or regional scales required
for biodiversity trend assessment [@valdez2023].

The two landscape-level EBVs (ecosystem structure and function) are well
suited to be examined at large spatial scales using coarser spatial
measurements, such as those taken from satellites by the Moderate
Resolution Imaging Spectroradiometer [MODIS, @zhang2003], the Landsat
imaging systems [@fisher2006], or Sentinel-2 [@helfenstein2022;
@darvishzadeh2019] programs. These mid-resolution satellites can monitor
processes at broader extents but the coarse spatial resolution removes
the ability to relate these traits to individual organisms. As a result,
satellite remote sensing data has been shown to be arguably the most
effective at monitoring ecosystem based EBVs focused on structure and
function. These EBVs classes can be monitored at regional to global
extents through the use of optical imagery [@cohen2004], as well as
active sensors such as lidar (light detection and ranging) and radar
[@guo2017; @lefsky2002; @lang2021; @neuenschwander2019; @coops2016].

## EBV - Ecosystem Structure

Forest structural diversity has been linked to biodiversity at various
scales [@guo2017; @bergen2009; @gao2014]. Structural attributes range in
complexity from simple (canopy cover; canopy height), to more complex
(vertical and horizontal structural complexity) to modelled (aboveground
biomass; basal area), all of which can be assessed using lidar data
[@coops2021]. A suite of these lidar-derived attributes have been used
as local indicators of biodiversity, including simple metrics such as
canopy cover and canopy height as well as derived metrics including
vertical profiles, aboveground biomass, or lidar derived indices
[@lefsky1999; @guo2017; @coops2016]. Other second order derived metrics
such as canopy texture, height class distribution, edges, and patch
metrics have also been used to examine habitat and biodiversity at
landscape scales [@bergen2009]. Advances in satellite remote sensing
processing have now allowed 3D forest structure data to be imputed
across wide spatial scales [@matasci2018; @coops2021] using data fusion
approaches involving collected lidar data and optical/radar data.

Increased forest structural complexity has been hypothesized to create
additional niches, leading to increased species diversity [@bergen2009],
which has been frequently demonstrated using avian species diversity
metrics [@macarthur1961]. For example: @herniman2020 used spectral and
lidar derived forest structure data to model avian habitat suitability;
@clawges2008 found that lidar derived forest structural attributes are
capable of identifying habitat types associated with avian species in
pine/aspen forests; @goetz2007 used canopy structural diversity to
predict bird species richness, finding that canopy vertical distribution
was the strongest predictor of species richness. Forest structural
metrics have also been used to study biodiversity in other clades as
well [@davies2014; @nelson2005].

## EBV - Ecosystem Function

With respect to ecosystem function, energy availability in an ecosystem
has shown to be a predictor of species richness and abundances at
various scales [@chase2002; @radeloff2019; @coops2019; @razenkova2023],
and is measurable using satellite remote sensing via the use of various
vegetation indices [@huete2002; @radeloff2019]. Vegetation indices,
which are indicative of photosynthetic activity, are commonly used as
proxies of gross primary productivity [@huang2019]. These vegetation
indices have also been used to assess patterns in biodiversity at single
time points [@bonn2004], and more recently, through yearly summaries of
productivity [@berry2007; @radeloff2019]. The relationship between
energy availability and biodiversity occurs via various hypothesized
mechanisms, such as the available energy hypothesis [@currie2004;
@wright1983], the environmental stress hypothesis [@currie2004], and the
environmental stability hypothesis [@williams2008]. These three
hypotheses have in turn been linked to patterns of annual surface
reflectance in remote sensing data [@berry2007; @radeloff2019].

@berry2007 first explored this idea by proposing the linkage of
intra-annual summaries of MODIS derived GPP to dispersive bird species.
This idea was further refined into the Dynamic Habitat Indices [DHIs,
@coops2008], which have now been shown to be well suited to assess the
three aforementioned hypotheses at global scales [@radeloff2019]. The
cumulative DHI calculates the total amount of energy available in a
given pixel over the course of a year. Cumulative DHI is strongly linked
to the available energy hypothesis, which suggests that with greater
available energy species richness will increase [@wright1983]. The
minimum DHI, which calculates the lowest productivity over the course of
a year can be matched to the environmental stress hypothesis, which
proposes that higher levels of minimum available energy will lead to
higher species richness [@currie2004]. Finally, the variation DHI, which
calculates the coefficient of variance in a vegetation index through the
course of a year, corresponds to the environmental stability hypothesis
which states that lower energy variation throughout a year will lead to
increased species richness [@williams2008].

## Biodiversity Monitoring with EBVs

Biodiversity monitoring programs often require a range of information in
order to accurately assess changes in ecological integrity
[@lindenmayer2010]. Choosing datasets that are most closely related to
the phenomenon of interest in a given application allows for direct
connections to monitoring results and management actions [@pressey2021].
With the advent of large-extent monitoring methods like satellite remote
sensing, and a proliferation of potential EBVs datasets, it becomes
important to assess the interrelationships between these datasets, and
assess their complementary of the information to reduce the duplication
of efforts [@pereira2013; @skidmore2021]. When strong relationships are
present between EBVs, it is possible to assess the ecological
relationships between potential EBVs. On the other hand, when datasets
do not appear related, they may be well suited to be used in monitoring
programs together, as complementary EBVs.

Linkages between forest ecosystem structure and function have been
examined within a remote sensing context for over 20 years [@huete2002,
@knyazikhin1998; @myneni1994]. While there is significant theoretical
and empirical evidence for their relationship at single time points
(within a single image) [@myneni1994], various relationship directions
and shapes have been found between forest structure and function metrics
[@ali2019]. Hypothesized mechanisms such as niche complementary have
shown that aboveground biomass increases with stand structure
[@zhang2012], while asymmetric competition for light can reduce forest
productivity with increased structural complexity [@bourdier2016]. The
relationship in particular between forest structural diversity metrics -
which are now more accurately and comprehensively derived from lidar
data - and temporal variation in functional metrics, specifically the
metrics of ecosystem productivity via the DHI framework, have yet to be
fully examined.

The overall goal of this paper is to assess patterns of forest ecosystem
structure and function, as observed using remote sensing technologies,
and their complementarity across a wide range of ecosystems encompassing
significant environmental gradients. To do so, we synthesize data from
moderate-scale remote-sensing derived metrics of forest structure,
represented as both simple ALS-extracted metrics of canopy height, cover
and vertical complexity, as well as modelled forest structure attributes
including volume, and aboveground biomass, with a well-established
remote sensing derived index on ecosystem function. First, we examine
how ecosystem structure and function complement one another across a
large environmental gradient and then compare the simple and modelled
representations of forest structure to different levels of ecosystem
function. This question is important as it provides insights to the EBV
community around complementarity of remote sensing metrics when
describing the structure and function of ecosystems and proposes a
method to examine potential overlap when generating remote sensing EBVs.

Our second question examines the independent and shared relationships of
ecosystem structure height and cover, with modelled forest structure, on
ecosystem function. This provides insight into the choice of remote
sensing attributes to use when developing EBVs within a single EBV
class. Remote sensing datasets can comprise relatively unprocessed
observations, in this case ALS measures of height and cover which are
derived from the raw 3D point cloud vs modelled attributes, such as
biomass and volume, which involve the use the statistical relationships
with field data to transform the observations into more refined data
products. Assessing which of these two (or combination of the two)
approaches has stronger or weaker relationships with estimates of
ecosystem function provides insights into the choice of data used to
build EBVs. Lastly, we examine how the primary and modelled structure
variables partition the variance of the DHIs within key biomes and
forest types across a large environmental range, examining to what
extent ecosystem and forest types impacts these relationships and thus
providing insight into the applicability of these results globally.

# Methods

## Study Area {#sec-study-area}

British Columbia is the westernmost province of Canada, and is home to a
variety of terrestrial ecosystems. Approximately 64% of the province is
forested, with large environmental and topographic gradients
[@pojar1987; @bcministryofforests2003]. The Biogeoclimatic Ecosystem
Classification (BEC) system identifies 16 zones based on the dominant
tree species and the ecosystems general climate (@fig-study-map).

```{r}
#| label: fig-study-map
#| fig-cap: Location of biogeoclimatic ecosystem classification (BEC) zones in British Columbia.

knitr::include_graphics(here::here("outputs", "bec_map.png"))
```

These zones can be further split into subzones, variants, and phases
based on microclimate, precipitation, and topography [@pojar1987]. To
examine trends across the large environmental gradients, we group the
BEC zones into five broad biomes, specifically, the southern interior,
northern interior, montane, alpine, and coastal groups similar to
@hamann2006. We also report each BEC zone's average climate data,
according to @wang2020 (@tbl-bec-group).

```{r}
#| label: tbl-bec-group
#| tbl-cap: BEC Zones, their aggregate groups, and their average climate values for precipitation, maximum temperature, and minimum temperature. Climate data from Wang et al. 2016.

# bec_tab %>%
#   gt() %>%
#   cols_label(
#     group = "BEC Group",
#     zone = "BEC Zone",
#     zone_nm = "Full Name",
#     ppt = "Precipitation (mm/yr)",
#     tmax = "Max Temperature (°C)",
#     tmin = "Min Temperature (°C)"
#   ) %>%
#   as_word()

knitr::kable(
  bec_tab,
  col.names = c(
    "BEC Group",
    "BEC Zone",
    "Full Name",
    "Precipitation (mm/yr)",
    "Max Temperature (°C)",
    "Min Temperature (°C)"
  )
)
```

## Data

### Forest Structure

We used a suite of forest structure variables (canopy height, canopy
cover, structural complexity \[coefficient of variation of height
returns\], basal area, aboveground biomass, and gross stem volume). This
dataset was created for the year 2015 at a 30 m spatial resolution
according to @matasci2018. In brief, the method used a set of lidar
collections and field plots across Canada, and imputed the remaining
pixels using a random forest k-Nearest Neighbour approach on
Landsat-derived surface reflectance and auxiliary data such as
topography. Detailed information on the creation of this dataset can be
found in @matasci2018.

### Dynamic Habitat Indices

We use an established set of indices of annual productivity shown to be
related to global biodiversity trends: the Dynamic Habitat Indices
[@radeloff2019]. The DHIs are a set of satellite remote sensing derived
productivity variables that summarize the cumulative amount of available
energy, the minimum available energy, and the variation in available
energy throughout a given year [@berry2007; @radeloff2019]. The DHIs
have previously been produced at a global extent using MODIS imagery,
and have previously been used to assess alpha [@radeloff2019] and beta
[@andrew2012] diversity, species abundances [@razenkova2023], and
construct novel ecoregionalizations [@coops2009; @andrew2013] . Recent
studies have began to examine how these indices can be constructed at a
finer spatial resolution by using multi-annual Landsat imagery to
generate a single synthetic year of monthly observations
[@razenkova2022].

The DHIs were calculated according to [@razenkova2022, Razenkova et al.,
In Press] for all of terrestrial British Columbia. In brief, Google
Earth Engine [@gorelick2017] was used to obtain all valid Landsat pixels
for a given study area, filtering out pixels containing shadows, clouds,
and cloud shadows within each image using the fmask algorithm
[@zhu2012], then calculated the NDVI for each pixel in each image. They
then calculated the median NDVI value for each month across a ten year
time span (2011-2020) to generate a synthetic year of monthly data. The
sum, minimum, and coefficient of variation across this synthetic year of
NDVI values was then calculated. More detailed information can be found
in Razenkova et al. (In Press).

A summary of remote sensing derived datasets can be found in Table 2.

# Sampling

We implement two sampling schemes -- one for ecosystems, and one for
forest types -- across the terrestrial ecosystems of British Columbia
(@fig-study-map). For both sampling schemes, each individual sample was
taken from a forested pixel which was surrounded by the same land cover
class. The land cover class map was generated for the year 2015
following [@hermosilla2022] using a best-available-pixel composite, and
an inverse-distance weighted random forest classification approach
across Canada. Additionally, each pixel had to have a coefficient of
variation less than 0.5 in surrounding pixels in the two simplest forest
structure variables, canopy height and canopy cover. A minimum sampling
distance of 1 km was implemented to reduce the effects of spatial
autocorrelation, and samples that had been disturbed in the last 30
years were discarded by using a disturbance mask generated for the
forested ecosystems of Canada by @hermosilla2016.

The ecosystem sampling scheme was implemented by randomly sampling
pixels matching the above criteria within each BEC zone of British
Columbia, up to a maximum of 3000 pixels, regardless of forest class.
The forest type sampling scheme was similar in that each BEC zone was
sampled, however, it implemented stratified sampling on each forest
class, up to a maximum of 500 pixels per BEC zone and forest class.

Each sampling scheme was natural-log transformed and standardized to
Z-scores. Variables containing zeros were natural-log plus one
transformed. Sampling was conducted in R [@R-base] version 4.2.2 using
the **sgsR** package [@R-sgsR]. Focal analyses for the land cover
classes and coefficient of variations of canopy height and cover were
calculated in Python version 3.9.

## Analysis

### Redundancy Analysis and Variation Partitioning

Redundancy analysis (RDA) and variation partitioning were used to relate
the primary and modelled forest structure variables to ecosystem
function across a broad environment range. Redundancy analysis functions
similarly to a multiple linear regression, except it is capable of
predicting multiple response variables. It accomplishes this by first
running a multiple linear regression of each predictor variable on each
response variable, then running a principle component analysis on the
residuals from each multiple linear regression. This reduces the
dimensionality of the output, and allows the relationship strength to be
assessed by calculating the loadings of both predictor and response
variables on the RDA axes. Partial redundancy analysis functions
similarly, except also considers co-variates [@legendre2012]. Redundancy
analysis has widely been used in community ecology where environmental
variables of interest are compared to species composition
[@blanchet2014; @kleyer2012] and has similarities to partial least
squares regression -- namely, their multivariate approach, usage of
dimensionality reduction, and linearity assumptions -- which is commonly
used in remote sensing literature [@roelofsen2014; @burnett2021].

Following the RDA, we employ ANOVAs to determine which axes are
significant, and calculate the proportion of variance attributable to
each axis using the eigenvalues generated from the RDA. We calculate
axis loadings for both forest structure and DHI variables by calculating
the correlation between each variable and the RDA axes. Forest strcuture
variable axis loadings represent the strength of relationship between a
given variable and the RDA axis, while DHI loadings indicate what is
being represented by the RDA axes. We only consider and display
significant axes. To visualize the RDA for both predictor and response
variables, we display the results as path diagrams, with loadings from
each predictor to the RDA axis to the response variables. The variance
explained by each axis is also displayed in the RDA box (@fig-rda-var).

Variation partitioning is an extension of partial RDA which can assess
the overlap between the explanatory power of two datasets by utilizing
multiple partial RDAs and exchanging which datasets are considered the
predictor, and which is considered the co-variate [@legendre2012].
Variation partitioning is traditionally displayed using a Venn diagram,
in which the percentage of variance explained by each dataset is in a
circle, and the overlap between circles represents the overlap in
variance explained.

RDA and variation partitioning analyses were conducted for all samples
in the ecosystem sampling scheme, as well as individually across each
ecosystem and forest type. The results were aggregated to BEC zone
groups (see @sec-study-area). All RDA and variation partitioning
calculations were done in R [@R-base] version 4.2.2 using the **vegan**
package [@R-vegan].

All code associated with the processing and analysis is available at
https://github.com/emuise/code-structProdSem.

# Results

```{r}
#| label: fig-rda-var
#| fig-cap: A) Axis loadings from redundancy analysis of primary and modelled forest structure variables on the dynamic habitat indices. B) Results from variation partitioning of primary and modelled forest structure variables on the DHIs. Both visualized analyses are across all collected samples. See supplementary information for results from each BEC zone and forest type.

knitr::include_graphics(here::here("outputs", "all_data_figure.drawio.png"))
```

To examine the relationship between ecosystem structure and function
across a large environmental gradient, we present the results of a
redundancy analysis of forest structure variables on the dynamic habitat
indices across the entire sampled dataset in @fig-rda-var A. While there
are three RDA axes associated with the full dataset, the third axis
explains 0.05% of the variance in the DHIs, and as a result is not
shown. The first axis strongly represents all the DHIs (loadings \> 0.85
for all DHIs), with the highest loadings from canopy cover, basal area,
aboveground biomass, and gross stem volume. The other input variables
(canopy height, structural complexity) have smaller loadings. The second
axis primarily represents the seasonality (Minimum and Variation) of the
DHIs, with no correlation to Cumulative DHI, and has the highest
loadings from canopy cover and complexity, with the remainder of the
forest structure variables having negligible influences(@fig-rda-var A).

The results from the variation partitioning analysis (@fig-rda-var B)
show that the majority of the variance explained by the input datasets
is shared across both primary and modelled forest structure variables.
Of the 13.4% of the variation in the DHIs being explained by the
structural information, 9.4% is common across both sets of structure
variables. The primary and modelled variables only explain 2.7% and 1.3%
of the variation uniquely (@fig-rda-var B).

```{r}
#| label: fig-radar
#| fig-cap: Radar plots of average loadings strength by group. A and B show input and response loadings, respectively. C) Boxplots of BEC zone, forest types, and all data loadings for predictor and response variables.

knitr::include_graphics(here::here("outputs", "radar_boxplot.png"))
```

Further, we wanted to examine the strength of the relationship between
the DHIs and primary structure variables vs modelled structure variables
across the forest types and ecosystems of the province. @fig-radar A and
B show the axis loadings for structure variables and the DHIs,
respectively. Across the BEC zone groups, the loadings are generally
similar in the structure variables. Variations in the DHIs in the
southern ecosystems, which are typically warmer and drier, have low or
non-significant variance explained by the structure variables.
Conversely, across the forest types (coniferous, broadleaf, mixed wood,
and wetland-treed) structural complexity is often the highest loading.
The loadings for the DHIs in the first axis are generally larger than
the loadings for the structure variables, with the cumulative DHI often
being the strongest loading overall. In alpine BEC zones, the minimum
DHI loading is smaller than the variation and cumulative DHIs. In
contrast to the first RDA axis, the secondary RDA axis is primarily
driven by variation in the minimum DHI, with medium loadings in the
variation DHI, and small loadings in the cumulative DHI (@fig-radar A &
B). @fig-radar C shows boxplots of individual BEC zone loadings for
structure variables and the DHIs. Across the province, canopy cover
generally has the highest loadings, with structural complexity generally
being the weakest.

```{r}
#| label: fig-fracts-bar
#| fig-cap: Stacked bar plot proportion of DHI variation explained by extracted, modelled, and the overlap between primary and modelled structure variables.

knitr::include_graphics(here::here("outputs", "fracts_bar_plot.png"))
```

@fig-fracts-bar shows stacked bar plots of the proportion of variation
explained in the DHIs by primary, modelled, and the overlap between
primary and modelled variables. Forest types sampled across the
environmental gradient generally show higher amounts of variance
explained, which is especially prevalent in broadleaf and mixed wood
forests. The overlap between modelled and primary structure variables is
generally than the primary and modelled attributes individually, with
the exception being mixed wood and broadleaf forests and colder, wetter
forests. Higher elevation and colder forest stands have larger amounts
of DHI variance explained by the forest structure variables than other
environments in the province. Notably, the variation explained by the
primary and modelled structure variables is commonly presented through a
single RDA axis, which generally corresponds to overall productivity
through the year (@fig-radar). Overall, the DHIs are decoupled from the
structure variables, as shown by the overall variance explained being
under 30%, regardless of whether the attribute is directly extracted
from the point cloud or is a modelled variable.

```{r}
#| label: fig-fcc
#| fig-cap: False colour maps of axis loadings for the first RDA axis (top) and second RDA axis (bottom). All colour values are normalized to the maximum loading of each variable. A and D show axis loadings for canopy height (r), canopy cover (g) and structural complexity (b). B and E show axis loadings for basal area(r), total biomass (g) and gross stem volume (b). C and F show axis loadings for the cumulative DHI (r), variation DHI (g) and minimum DHI (b).

knitr::include_graphics(here::here("outputs", "rgb_norm_edit.png"))
```

@fig-fcc shows false colour composites of the primary (A & C) and
modelled (B & D) loadings across the ecosystems of British Columbia. In
the first RDA axis there is spatial variation in the primary structure
variables, with interior zones being primarily driven by canopy cover
(green), the coastal regions having strong structural complexity and
canopy height loadings, and the boreal in the northwest having the
strongest loading in canopy height (@fig-fcc A). The modelled structure
variables generally show gray-scale colour, indicating that basal area,
total biomass, and gross stem volume explain similar variation in the
DHIs (@fig-fcc B). All ecosystems with a secondary axis are located away
from the Pacific coast in cooler, drier environments. Canopy cover has
the strongest secondary axis loading in the northern boreal stands,
while the secondary axis in the remaining four zones have the highest
loadings in canopy height (@fig-fcc C). Again, the modeled structure
variables are in grayscale, indicating similar loadings across the three
modelled structure variables (@fig-fcc E).

# Discussion

Complementarity plays a crucial role in the context of Essential
Biodiversity Variables (EBVs) and holds significant importance for
monitoring and understanding biodiversity [@pereira2013; @skidmore2021].
At the ecosystem scale, the interconnectedness and interdependence of
ecosystem structure and function is important to recognize and
disentangle. In forested environments, forest structure and productivity
have been shown to be linked in multiple studies spanning three decades
[@ali2019; @myneni1994], however, a linkage between intra-annual
production and forest structure has yet to be shown. In this study, we
use statistical analyses commonly used in community ecology - namely
redundancy analysis and variation partitioning - to assess the
complementarity of forest structure and yearly productivity summaries.
While RDA is common in the ecological literature, this analysis
represents one of the first times this technique has been applied to
assess the complimentary of proposed satellite-derived EBVs. We find
that the example EBVs do not strongly overlap, with forest structure
explaining 13.4% of the variation in the DHIs in samples taken across a
broad environmental gradient (@fig-rda-var B). This indicates that they
are suitable to be used in tandem with one another when used as
ecosystem EBVs across wide environmental gradients.

Across most of British Columbia's ecosystems, we identified a single RDA
axis associated with the DHIs, encompassing the variation in annual
productivity (@fig-rda-var A). Within this first axis, the strongest
loadings were canopy cover and the modelled structure variables. In
Alpine ecosystems the relationship between the first axis and the
Minimum DHI was lowest, likely due to the low amount of variation in the
Minimum DHI in these ecosystems. When a second axis was significant, it
consistently had strong loadings on the minimum and variation DHIs,
which indicates a complex internal productivity relationship in certain
ecosystems. This secondary axis has smaller axis loadings associated
with the primary and modelled structure variables, with the strongest
loadings being canopy cover and structural complexity across the entire
dataset (@fig-rda-var A).

Our second question sought to explore whether modelled structure
variables (basal area, gross stem volume, aboveground biomass) add
additional explanatory information when predicting the DHIs, as compared
to primary forest structure variables. We generally found that canopy
cover had the largest axis loadings in the first RDA axis (@fig-radar).
Modelled structure variables such as basal area, aboveground biomass,
and gross stem volume shared similar loading magnitudes across the range
of studied ecosystems, indicating they do not add additional value when
utilized with one another (@fig-radar C; @fig-fcc B). The loadings
between the modelled structure variables and canopy cover are often
similar, and as such, in this context we recommend utilizing the
structure variables derived directly from the point cloud in the case of
ALS data, or selecting a single modelled attribute, such as aboveground
biomass alone [@duncanson2022].

During the analysis we found a higher amount of total variance explained
in deciduous forests (mixed-wood and broadleaf) when compared to the
other two forest types and most BEC zones (@fig-fracts-bar). This could
potentially be due to the temporal linkages between canopy cover in
deciduous forests and productivity. In evergreen coniferous forests
where canopy cover is present year-round, there is likely a temporal
mismatch between forest structure variables and the DHIs. Further, the
strongest loadings in these two forest types was vertical structural
complexity, rather than canopy cover. This indicates that the DHIs are
more closely related to forest structure in deciduous forests, however
even there significant decoupling remains.

Finally, we explored the amount of variation explained by primary vs
modelled structure variables, as well as their overlap
(@fig-fracts-bar). We generally found that the overlap between the
primary and modelled structure variables explained most of the
variation, with some exceptions, indicating that using either set of
forest structure variables may be suitable when monitoring biodiversity.
The exceptions are especially prominent in the broadleaf and mixed wood
forest classes, as well as cool, highly productive forests where there
is no variation explained by the overlap between the two datasets. The
majority of the variance explained in the broadleaf and mixed wood
forests was in the forest structure variables derived directly from the
point cloud. This was expected, as the modelled structure variables are
calculated based on the derived structure variables as well as site
index and ancillary information [@coops2021].

Recent advances in creating synthetic yearly observations have allowed
the DHIs to be generated at a finer scale (30 m; Razenkova et al., In
Press), rather than the previously used 1 km DHIs derived from MODIS
[@radeloff2019; @razenkova2022]. This represents a significant
advancement when assessing the utility of EBVs, as the 30 m scale is
well suited to examine a range of ecological applications, including
forest structure and productivity, and more closely approximates the
scale at which species perceive habitat [@cohen2004; @kennedy2014]. In
addition to more closely matching the scale of ecological phenomenon,
this allows these datasets to be matched and analyzed with other
datasets generated from the Landsat archive.

In conclusion, we used redundancy analysis and variation partitioning to
assess the complementarity of two potential EBV datasets - forest
structure and the DHIs. We also separated the forest structure datasets
into primary and modelled structure variables in order to assess the
need to develop more complex structure variables, or if data derived
directly from lidar datasets was suitable. We found that the structure
variables are not strongly related to the DHIs, indicating that they are
suitable to be used together as ecosystem scale EBVs when monitoring
forest environments. We also found that variation explained by the
overlap between primary and modelled structure variables was often
higher than the variation explained by either individually.

\newpage

# References {.unnumbered}