### 2023-07-24 scripts/manuscript/manuscript.qmd

---
title: Disentangling linkages between satellite-derived forest structure and productivity essential biodiversity variables.
author:
  - name: Evan R. Muise
    email: evan.muise@student.ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
    attributes:
        corresponding: true
  - name: Margaret E. Andrew
    email: M.Andrew@murdoch.edu.au
    affiliations: 
        - id: murd
          name: Murdoch University
          department: Environmental and Conservation Sciences and Harry Butler Institute
          city: Murdoch, WA, Australia
          postal-code: 6150
  - name: Nicholas C. Coops
    email: nicholas.coops@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Txomin Hermosilla
    email: txomin.hermosillagomez@NRCan-RNCan.gc.ca
    affiliations: 
        - id: cfs
          name: Natural Resources Canada
          department: Canada Forest Service (Pacific Forestry Centre)
          address: 506 Burnside Rd W
          city: Victoria, BC, Canada
          postal-code: V8Z 1M5
  - name: A. Cole Burton
    email: cole.burton@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Stephen S. Ban
    email: Stephen.Ban@gov.bc.ca
    affiliations: 
        - id: bcp
          name: Ministry of Environment and Climate Change Strategy
          department: BC Parks
          address: 525 Superior Street
          city: Victoria, BC, Canada
          postal-code: V8V 1T7
keywords: 
  - remote sensing
  - landsat
  - forest structure
  - forest productivity
  - dynamic habitat indices (DHIs)
  - essential biodiversity variables
  - redundancy analysis (RDA)
date: last-modified
bibliography: [bibliography.bib, packages.bib]
csl: remote-sensing-of-environment.csl
format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    journal:
      name: Nature Scientific Reports
      formatting: review
      model: 3p
      cite-style: authoryear
      layout: twocolumn
  docx: 
    reference-doc: my-styles.docx
    keep-md: true
  pdf: 
    pdf-engine: xelatex
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: false
fig-width: 300
number-sections: true
---

```{r}
#| warning: false
#| echo: false
#| output: false

# include all libraries required by the R chunks here
library(vegan)
library(sgsR)
library(terra)
library(sf)
library(gt)
library(tidyverse)

library(bcmaps)

# generate package citations
loaded_pkgs = (.packages())
base_pkgs = getOption("defaultPackages")
contrib_pkgs = loaded_pkgs[!(loaded_pkgs %in% base_pkgs)]
pkgs_file = here::here("scripts", "manuscript", "packages.bib")
knitr::write_bib(contrib_pkgs, pkgs_file)

bec_tab <- read_csv(here::here("data", "climate.csv")) %>%
  mutate(across(.cols = c(tmax, tmin), \(x) round(x, digits = 1)))
```

\newpage

**Abstract**

Key to regular biodiversity monitoring is the definition and selection
of variables capable of accurately monitoring biodiversity trends. One
framework is the Essential Biodiversity Variables (EBV), which are
standardized, comparable data that represents a minimum set of
biological information required for monitoring biodiversity change at
large spatial extents, with each class being complementary to one
another -- i.e., capturing unique biological information. The
interrelationships between forest ecosystem structure and function and
the complementarity between these two classes of EBVs across a range of
ecosystems encompassing significant environmental gradients are examined
using redundancy analysis (RDA) and variation partitioning. We found
that the primary and modelled forest structure variables are not
strongly related to the Dynamic Habitat Indices (DHIs), with the
structural information only explaining 16% of the variation in the DHIs,
indicating that structure and function EBVs are suitable to be used
together for monitoring forest ecosystems. We also found that forest
types and ecosystems dominated by coniferous trees had less variation
explained, potentially due to the temporal decoupling of canopy cover
and the DHIs. We suggest that biodiversity researchers focus on using
forest structural attributes derived directly from lidar data alongside
the DHIs as complementary EBVs for monitoring forest ecological
integrity.

\newpage

# Introduction

Monitoring the changing state of biodiversity globally is key to
understanding and mitigating the accelerated extinction risk for many
species [@thomas2004; @urban2015] and the homogenization of biotic
communities at various spatial and temporal scales
[@mcgillFifteenFormsBiodiversity2015]. The Group on Earth Observations
Biodiversity Observation Network has developed the Essential
Biodiversity Variables [EBVs, @pereira2013], which are standardized,
comparable data that represents a minimum set of biological information
required for monitoring biodiversity change at large spatial extents
[@skidmore2021]. The EBVs are designed as an analog to the Essential
Climate Variables framework [@bojinski2014] to be global in scope,
relevant to biodiversity information, feasible to implement and use, and
complementary to one another [@skidmore2021]. While it can be difficult,
time consuming, and expensive to collect data on biodiversity across
wide swaths of land, EBVs can be combined with sampled biodiversity
information to allow for the spatial extrapolation of more sparsely
collected in situ data [@cavender-bares2022]. The standardization and
global availability of the EBVs allows for the assessment and monitoring
of protected area effectiveness and ecosystem health at large spatial
scales [@hansen2021], which can then be integrated into policy
development [@geijzendorffer2016]. There are six EBV classes, each of
which provide information on a different facet of biodiversity: species
populations, species traits, community composition, ecosystem structure,
ecosystem function, and genetic composition [@pereira2013]. While these
EBV classes can be related to one another under a variety of ecological
and evolutionary mechanisms, they provide distinct information; multiple
in concert can provide a more complete picture on the status and trends
of biodiversity in a region.

The global coverage and moderate spatial resolutions of many optical
satellite remote sensing missions [@wulder2022], combined with open-data
policies [@wulder2012], make these datasets exceptionally well suited
for contributing to some EBV classes [@skidmore2021]. The six EBV
classes can be divided into two general approaches: species-focused and
ecosystem-focused [@fernández2020]. Genetic composition, species
populations, and species traits are species-focused EBVs, with community
composition being directly related to species populations. While
species-focused and community composition EBVs can be assessed from
measurements at the tree-crown scale using very-high spatial resolution
imagery [@fassnacht2016; @graves2016; @wu2021], the spatially
limited--and often *ad-hoc*--data collection associated with monitoring
individuals is not conducive to the global or regional scales often
required for biodiversity trend assessments [@valdez2023]. In contrast,
two ecosystem-based EBV classes (ecosystem structure and ecosystem
function) are well suited to be monitored across large extents using
moderate spatial resolution optical imagery, such as the Moderate
Resolution Imaging Spectroradiometer [MODIS, @zhang2003], Landsat
imaging systems [@wulder2022] or Sentinel-2 [@helfenstein2022;
@drusch2012] programs, especially in forested environments
[@pettorelli2018; @coops2016].

Ecosystem structure is the composition, abundance, and spatial
arrangement of different ecosystem components, and is indicative of
habitat quality and connectivity [@noss1990; @valbuena2020]. There are
both horizontal and vertical components of ecosystem structure. In
forest environments, forest structural diversity has been linked to
species richness at various spatial scales, ranging from individual
plots to landscapes [@bergen2009; @gao2014]. Structural attributes range
in complexity from simple (e.g., canopy cover, canopy height), to more
complex (e.g., foliage height diversity, leaf area index) to modelled
(e.g., aboveground biomass, basal area), which can be derived using
lidar (light detection and ranging) data [@coops2021]. A suite of these,
and other, structural attributes has been used as indicators of
biodiversity at local to landscape scales [@lefsky1999; @guo2017;
@coops2016; @bergen2009]. Increased forest structural complexity has
been hypothesized to create additional niches, leading to increased
species diversity [@bergen2009], which has been frequently demonstrated
using avian species diversity metrics [@macarthur1961], with lidar
[@clawges2008], a combination of lidar and satellite imagery
[@herniman2020], with some research indicating that canopy vertical
distribution was the strongest predictor of species richness
[@goetz2007]. Although lidar datasets have traditionally been local in
extent, advances in satellite remote sensing processing have allowed the
extension of three-dimensional forest structure attributes nationally in
a wall-to-wall fashion. The wall-to-wall data often uses data fusion
approaches involving lidar data and medium spatial resolution optical
imagery [@matasci2018; @matasci2018b; @coops2021].

Ecosystem functions are measures of ecosystem performance that are the
consequence of one or multiple ecosystem processes [@pettorelli2018;
@ecosyste2005]. Within the EBV framework, vegetation phenology and
metrics of primary productivity are commonly used as potential EBVs
[@skidmore2021]. Energy availability is the amount of accessible and
utilizable energy available for growth in an ecosystem, and is often
measured as net primary productivity [@bonn2004]. Energy availability
has shown to be a predictor of species richness and abundances at
various scales [@chase2002; @radeloff2019; @coops2019; @razenkova2023],
and is measurable using satellite remote sensing via various vegetation
indices [@huete2002; @radeloff2019]. Vegetation indices are indicative
of photosynthetic activity, and are commonly used as proxies of gross
primary productivity [@huang2019]. These vegetation indices enable the
assessment of biodiversity patterns at single time points [@bonn2004],
or through time when using time series data [@berry2007; @radeloff2019].

The relationship between energy availability and biodiversity can be
explained through various hypothesized mechanisms: including available
energy, environmental stress, and environmental stability. These three
hypotheses have been incorporated into the Dynamic Habitat Indices
(DHIs), which summarizes intra-annual dynamics in remotely-sensed
productivity metrics and have been evaluated as broad-scale biodiversity
indicators [@berry2007; @coops2008; @radeloff2019]. According to the
available energy hypothesis [@wright1983] an increase in available
energy results in higher species richness. In the context of the DHIs,
the total energy available within a specific area is quantified using
the cumulative DHI, which sums each productivity observation throughout
the course of a year. The environmental stress hypothesis [@currie2004]
suggests that higher levels of minimum available energy contribute to
greater species richness. The minimum DHI, derived from the lowest
productivity observed over the course of a year, indicates the level of
available energy during stressful periods. The environmental stability
hypothesis [@williams2008] proposes that lower energy variation
throughout the year leads to increased species richness. The variation
DHI captures the coefficient of variation in a vegetation index over the
course of a year, reflecting the stability of energy availability. The
DHIs have previously been produced at a global extent using MODIS
imagery and have been used to assess alpha [@radeloff2019] and beta
[@andrew2012] diversity, species abundances [@razenkova2023], and
construct novel ecoregionalizations [@coops2009; @andrew2013]. Recent
studies have increased the spatial resolution of these indices by
calculating the DHIs from multi-annual Landsat imagery
[@razenkovaMediumresolutionDynamicHabitatInPress].

Choosing datasets that are most closely related to the phenomenon of
interest in a given application allows for direct integration of
monitoring results with management actions [@pressey2021]. With the
advent of large-extent monitoring methods like satellite remote sensing
which have led to a proliferation of potential EBV datasets, it becomes
important to assess the interrelationships between these datasets and
the complementary of the information they provide to reduce redundancy
in future EBV development efforts [@pereira2013; @skidmore2021].
Identifying and exploring strong relationships between EBVs may help
elucidate the ecological relationships between these facets of
biodiversity, such as ecosystem structure and function, at local to
global scales. On the other hand, when variables are not related, they
may be well suited to be used in monitoring programs together, as
complementary EBVs.

Linkages between forest ecosystem structure and function have been
examined within a remote sensing context for decades [@huete2002,
@knyazikhin1998; @myneni1994]. While there is significant theoretical
and empirical evidence for their relationship [@myneni1994], the
direction and significance has not been consistent, with positive,
negative, and non-significant relationships being found [@ali2019].
Hypothesized mechanisms such as niche complementarity suggest that
productivity increases with stand structural complexity [@zhang2012],
while asymmetric competition for light can reduce forest productivity
with increased structural complexity [@bourdier2016]. Some aspects of
forest structure, such as canopy cover and leaf area index, also
directly influence vegetation indices, which are used to estimate
productivity, especially in very dense or very sparse forests
[@pettorelli2005]. Further, the relationship between forest structural
diversity metrics--which are now more accurately and comprehensively
derived from lidar data--and the temporal variation in functional
metrics, specifically the metrics of ecosystem productivity via the DHIs
framework, have yet to be fully examined.

The goal of this study is to examine relationships between forest
ecosystem structure and function, and the complementarity between these
two classes of EBVs across significant environmental gradients. We
synthesize data from moderate-scale remote sensing-derived metrics of
forest structure (simple lidar-extracted metrics of canopy height,
canopy cover, and structural complexity), as well as modelled forest
structure attributes (gross stem volume, basal area, and aboveground
biomass), with a well established remote sensing-derived index for
ecosystem function (DHIs). First, we analyze remote sensing metric
complementarity by examining the multivariate relationship between
ecosystem structure and function and propose a method to examine
potential overlap in remote sensing EBVs. Second, we propose a method
for the selection of remote sensing attributes for developing and
applying EBVs within a single EBV using the independent and shared
relationships of lidar-derived forest structural attributes and modelled
forest structure on ecosystem function. Third, we examine the global
applicability of these results using the relationship between simple and
modelled forest structural attributes with DHIs across a range of
environment gradients and forest types.

# Methods and Materials

## Study Area {#sec-study-area}

British Columbia (BC) is the westernmost province of Canada, and is home
to a variety of terrestrial ecosystems. Located between the Pacific
Ocean and the Rocky Mountains, approximately 64% of the province is
forested, with large environmental and topographic gradients
[@pojar1987; @bcministryofforests2003]. The majority of the forests of
BC are coniferous forests (87.4%), with the remaining forests being
broadleaf (8.9%), wetland-treed (2.7%) and mixed-wood (0.9%)
[@hermosilla2022]. There is a large topographic gradient in the
province, with elevations ranging from sea level to over 4000 m, with
much of the province being mountainous (elevations above 1000 m)
[@valentine1978soil]. Wildfires are a regular disturbance in the
province, with the historical fire regime being disrupted due to
management actions, leading to larger, higher intensity fires becoming
more common in recent years [@brookes2021]. In addition, forestry is a
large industry in the province, with approximately 136000 ha of the
province harvested accounting for \~54 million cubic metres of volume of
wood in 2021 [@canadiancouncilofforestministers2023].

BC is stratified into 16 zones based on the dominant tree species and
climate by the Biogeoclimatic Ecosystem Classification (BEC) system
(@fig-study-map). To examine trends across the large environmental
gradients, we group the BEC zones into five broad biomes: the Nouthern
interior, Sorthern interior, Montane, Alpine, and Coastal groups,
similar to @hamann2006. We also report the average climate data for BEC
zones from 1991-2020, according to @wang2016 (Table 1).

```{r}
#| label: fig-study-map
#| fig-cap: Location of Biogeoclimatic Ecosystem Classification (BEC) zones in British Columbia.

knitr::include_graphics(here::here("outputs", "bec_map.png"))
```

```{r}
#| warning: false
#| echo: false

bec_tab %>%
  gt() %>%
  cols_label(
    group = "BEC Group",
    zone = "BEC Zone",
    zone_nm = "Full Name",
    ppt = "Precipitation (mm/yr)",
    tmax = "Max Temperature (°C)",
    tmin = "Min Temperature (°C)"
  ) %>%
  tab_header(title = "Biogeoclimatic Ecosystem Classification (BEC) Zones, their aggregated biomes, and their average climate values for precipitation, maximum temperature, and minimum temperature. Climate data from Wang et al. 2016 averaged for 1991-2020.")
```

## Data

### National Terrestrial Ecosystem Monitoring System (NTEMS)

#### Best-Available-Pixel

The forest structure, forest disturbance, and land cover layers are
based on a 30-m best-available-pixel (BAP) composite generated by
@hermosilla2016 using the Composite2Change approach. The composites were
developed for each year from 1984-2019 by selecting the best available
imagery from the available growing season Landsat observations, removing
pixels with clouds, cloud shadows, and haze. The methodology uses the
scoring method from @white2014 to select the best pixel for a given year
from Landsat-5 Thematic Mapper, Landsat-7 Enhanced Thematic Mapper Plus,
and Landsat-8 Operational Land Imager imagery. To reduce noise in the
BAP between years, a spectral trend analysis on the Normalized Burn
Ratio was conducted on each pixel, which removes unscreened noise,
detects changes, and temporally interpolates data gaps. This resulted in
a gap-free surface-reflectance composite across the entirety of Canada
[@hermosilla2015]. We utilize data from 2015 for all NTEMS data, except
forest disturbances, where we use all available years of data in the
terrestrial area of BC, Canada.

#### Forest Disturbances

We used a forest disturbance layer generated by @hermosilla2015a to
restrict the analysis to stands older than 35 years as productivity and
structure metrics have been shown to be strongly decoupled in early
regenerative stands [@bolton2017], and also excluding recently
anthropogenically disturbed stands. Changes were detected using spectral
trend analysis and were attributed using an object-based analysis
approach to a disturbance agent (i.e., wildfire, harvest) using a Random
Forests classification model [@hermosilla2015a]. Changes were detected
with an overall accuracy of 89%, attributed with an accuracy of 92%, and
assigned to the correct year with an accuracy of 89% [@hermosilla2016].

#### Land Cover {#sec-lcc}

We used a land cover map generated by @hermosilla2022 to stratify our
samples alongside the BEC zones. The 30-m spatial resolution land cover
map for Canada's forested ecosystems was generated using the Virtual
Land Cover Engine framework [@hermosilla2022]. This framework uses a
Random Forests based classification approach, and integrates logical
land cover transitions using a Hidden Markov Model to generate 12 land
cover classes. Training samples were locally selected using an inverse
distance weighting approach [@hermosilla2022]. The treed land cover
classes are broadleaf, coniferous, mixed wood, and wetland-treed, which
we use as a stratification. The land cover classification reached an
overall accuracy of 77.9% ± 4% [@hermosilla2022].

#### Forest Structure

Forest structural attributes (canopy height, canopy cover, structural
complexity \[coefficient of variation of height returns\], basal area,
aboveground biomass, and gross stem volume) for 2015 were extracted for
BC from the 30-m spatial resolution dataset generated by Matasci et al.
(2018a, 2018b). In brief, the method extends structural attributes
derived from a set of lidar acquisitions and field plots across Canada.
These data were generated for the treed pixels of Canada using the BAP
Landsat surface reflectance composite and auxiliary data (i.e.,
topography, geography) using a k-Nearest Neighbour imputation
[@matasci2018; @matasci2018b]. Accuracy metrics for imputed forest
structural attributes range from an RMSE of 24.5% (structural
complexity) to 82.3% (gross stem volume) and *R^2^* ranging from 0.125
(structural complexity) to 0.712 (gross stem volume) [@matasci2018;
@matasci2018b].

### Dynamic Habitat Indices

The DHIs were calculated following
@razenkovaMediumresolutionDynamicHabitatInPress for the terrestrial area
of BC. In brief, we used Google Earth Engine [@gorelick2017] to obtain
valid Landsat pixels for the study area across a ten-year time span,
centered on 2015 (2011-2020), filtering out pixels containing shadows,
clouds, and cloud shadows using the QA band derived from the fmask
algorithm [@zhu2012], and then calculated the NDVI index. These
estimates were composited into a synthetic year of monthly data by
calculating the monthly median NDVI value. Finally, images of the sum,
minimum, and coefficient of variation across this synthetic year of NDVI
values were then calculated to represent the cumulative, minimum, and
variation DHIs, respectively. More detailed information can be found in
@razenkovaMediumresolutionDynamicHabitatInPress.

A summary of remote sensing derived datasets can be found in Table 2.

## Sampling

To obtain representative sample units of the EBV metrics across the
study area, we implemented stratified random sampling across the union
of BEC zones (Table 1) and forest types (i.e., coniferous, broadleaf,
mixed wood, and wetland treed; see @sec-study-area) of BC. A maximum of
500 sample units were selected from each stratum. Sample units were
restricted to forested pixels that were surrounded by the same forest
type to reduce the chance of sampling mixed pixels. To ensure that
sample units were selected from a homogeneous area, reducing uncertainty
associated with the input datasets [@shang2020], we required sample
units to have a coefficient of variation for canopy cover and canopy
height to be lower than 0.5 in a 3 x 3-pixel window. A minimum sampling
distance of 1-km was implemented to reduce the effects of spatial
autocorrelation, and sample units that had been disturbed in the last 30
years were discarded by using the disturbance mask generated for the
forested ecosystems of Canada as early regenerative forests have shown
to have strongly decoupled productivity and canopy cover [@bolton2017].

Sampling was conducted in R [@R-base] version 4.2.2 using the **sgsR**
package [@R-sgsR]. Neighbourhood analyses for the land cover classes and
coefficient of variations of canopy height and cover were calculated in
Python version 3.9.

## Analysis

Redundancy analysis (RDA) and variation partitioning were used to relate
the primary and modelled forest structure variables to ecosystem
function across a broad environment range. RDA has widely been used in
community ecology where environmental variables of interest are compared
to species composition [@blanchet2014; @kleyer2012]. RDA predicts
multiple response variables by first running a multiple linear
regression of the predictor variables on each response variable, and
then a principal component analysis on the fitted values from each
multiple linear regression. This reduces the dimensionality of the
fitted values by transforming them into a set of independent RDA axes,
and allows the strength of the multivariate relationship between
predictors and responses to be assessed by calculating eigenvalues of
the RDA. Relationships between variables are revealed by the loadings of
both predictor and response variables on the RDA axes. RDA axes are
labelled as RDA1, RDA2, etc. to indicate they are derived from
redundancy analysis, rather than other ordination techniques such as
principle component analysis or constrained correspondence analysis
[@legendre2012]. Variation partitioning is an extension of partial RDA
-- which allows predictor variables to be considered within pre-defined
groups of co-variates -- that can assess the overlap between the
explanatory power of two datasets by utilizing multiple partial RDAs and
exchanging which datasets are considered the predictor, and which are
considered the co-variates [@legendre2012].

We performed RDAs for our full sample, pooling BEC zones and forest
types, as well as individually by BEC zone and forest type, with results
aggregated to BEC zone groups (see @sec-study-area). All measured values
were natural-log transformed and standardized to Z-scores based on the
summary statistics of the full sample prior to analyses. Forest
structure variables were treated as predictor variables and were grouped
based on whether they were primary estimates from the lidar data or
modelled from lidar and ancillary information (Table 2) for variance
partitioning. The three DHIs were used as the response variables.
Analysis of variance tests were used to determine which RDA axes were
significant, with p-values below 0.1 using an F-test. The proportion of
variance attributable to each axis was calculated from the eigenvalues
generated by the RDA, and axis loadings were calculated as the Pearson
correlation between each forest structure or DHI variable and the RDA
axes. Forest structure variable axis loadings represent the strength of
relationship between a given variable and the RDA axis, while the DHIs
loadings indicate what is being represented by the RDA axes. To
visualize the RDA results for both predictor and response variables, we
display the results as path diagrams, with paths annotated by the
loadings and RDA axes annotated by the variance in the DHIs explained by
each axis (example shown in @fig-rda-var-template). We only consider and
display significant RDA axes.

Variation partitioning is displayed using a Venn diagram in which the
percentage of variance explained by each dataset is in a circle, and the
overlap between circles represents the overlap in variance explained. We
add paths to the Venn diagram to indicate how the structural attributes
contribute to the variation in the DHIs. Because the variation
partitioning results summarize the full multivariate solution, rather
than individual RDA axes, we present variation partitioning results
alongside the path diagrams illustrating the details of the RDA results
(@fig-rda-var-template).

```{r}
#| label: fig-rda-var-template
#| fig-cap: Template for visualization of RDA and variation partitioning results. Paths between RDA axes and predictor/response variables will be labelled with axis loadings, caluclated as the correlation between the variable and the RDA axis. If loadings are non-significant, NS will be shown on the path.

knitr::include_graphics(here::here("outputs", "template.drawio.png"))
```

RDA and variation partitioning calculations were done in R [@R-base]
version 4.2.2 using the **vegan** package [@R-vegan]. The code
associated with the processing and analysis is available at
https://github.com/emuise/code-structProdSem.

# Results

Overall in BC, the intra-annual patterns of productivity as represented
by the DHIs variables were weakly related to forest structure
(@fig-rda-var). Less than 16% of the overall variation in the DHIs was
explained by the forest structural attributes. Due to strong
correlations between the axes of the DHIs, the majority of the variation
was held within the first RDA axis, which had strong loadings for all
DHIs, with positive loadings for the cumulative and minimum DHIs, and
negative loadings for the variation DHI. As such, we identified the RDA1
axis as representing overall productivity. The overall productivity axis
had similar loadings (between 0.24-0.27) for all modelled predictors and
canopy cover. Canopy height had a smaller loading (0.1), while
structural complexity was negatively correlated with the overall
productivity axis. RDA2 explained much less of the variation in the
DHIs. RDA2 had a large positive loading on the minimum DHI (0.55), with
negative loadings on the cumulative (-0.19) and variation (-0.39) DHIs.
Due to the minimum and variation having the most influence on RDA2, we
identified it as a seasonality axis. Canopy cover and structural
complexity had the strongest (albeit negative; -0.17 and -0.08,
respectively) loadings on the seasonality axis (@fig-rda-var A).
Variation partitioning highlighted that the majority of the explained
variation in the DHIs was due to the primary attributes (canopy cover,
canopy height, and structural complexity; 9% variation explained). The
overlap between primary and modelled attributes was 3.5%, and the
modelled attributes explained 3.2% of the variation on their own
(@fig-rda-var B).

```{r}
#| label: fig-rda-var
#| fig-cap: A) Axis loadings from redundancy analysis (RDA) of primary and modelled forest structure variables on the dynamic habitat indices (DHIs). B) Results from variation partitioning of primary and modelled forest structure variables on the DHIs. Both visualized analyses are across the entire dataset. Non-significant paths are labelled as "NS". See supplementary information for results from each BEC zone and forest type.

knitr::include_graphics(here::here("outputs", "drawio", "British Columbia.drawio.png"))
```

Subdividing the sample by forest type led to similar or increased
amounts of variance explained when compared to the overall dataset
(@fig-fracts-bar). Subdividing by BEC zones generally led to smaller
amounts of variation explained than BC or stratifying by forest type,
which may be due to the length of the environmental gradients included
in the resulting subsamples (@fig-fracts-bar). The overlap between the
primary and modelled datasets often explained the majority of the
variation in the DHIs, except when stratifying by forest types, where
this is only true for the coniferous subsample. The broadleaf, mixed
wood, and wetland-treed forest types contained most of the DHIs
explained variation in the primary structural attributes. The BEC zones
followed a similar pattern, with most zones' DHI variation principally
explained by the overlap, with the exception being CWH, which has no
variation explained by the overlap between primary and modelled
datasets. The Coastal BEC group had the lowest amount of variation
explained, with Coastal Douglas-fir having \<1% of variation explained
by the structural data. Overall, the DHIs are decoupled from the forest
structural attributes, as shown by the overall variation explained being
under 30% regardless of data stratification (@fig-fracts-bar).

```{r}
#| label: fig-fracts-bar 
#| fig-cap: Stacked bar plot proportion of the variation in the Dynamic Habitat Indices (DHIs) variation explained by primary, modelled, and the overlap between primary and modelled structure variables across the stratified and overall datasets.  
knitr::include_graphics(here::here("outputs", "fracts_bar_plot.png"))
```

Stratifying the dataset by forest type and ecosystem led to varying
loadings between the RDA axes and predictor/response variables
(@fig-radar). The DHIs' loadings on RDA1 are large (similar to the
overall dataset), except in colder BEC groups (Montane, North, and
Alpine), where the minimum DHI loading half as large as the cumulative
or variation DHI loadings. As such, we do not identify the first axis as
being consistently associated with overall productivity. Non-coniferous
forest types (broadleaf, mixed wood and wetland-treed) typically had
large loadings between RDA1 and structural complexity. The mixed wood
forest and Coastal BEC group had no significant loadings from modelled
forest structural attributes for RDA1. Canopy cover generally had the
highest significant predictor loading on average in RDA1. Examining the
predictor loadings for RDA1 spatially across the province shows a
geographic pattern in the loading strengths of primary structural
attributes (@fig-fcc A). The interior (Northern and Southern BEC groups)
of the province generally showed high canopy cover loadings, while the
Boreal zones in the northwest had equal loadings in canopy cover and
canopy height. Coastal were driven similarly by canopy height and
structural complexity, with very low or non-significant loadings from
canopy cover (@fig-fcc A; @fig-radar). Modelled attributes for RDA1
generally show similar loadings across the province (@fig-fcc B).

```{r}
#| label: fig-radar
#| fig-cap: Parallel coordinate plots of average absolute significant loading strength by Biogeoclimatic Ecosystem Classification (BEC) group, forest type, and British Columbia. Note the varying y axes.

knitr::include_graphics(here::here("outputs", "parcoord_noflip_edit.png"))
```

Only seven of sixteen BEC zones had a second RDA axis, with none being
found in Coastal or Alpine BEC zones (@fig-radar; @fig-fcc) . All of the
Montane BEC zones had a secondary axis, while some in the Northern and
Southern interior had a secondary axis. RDA2 often showed the largest
absolute loadings in the minimum DHI, which was frequently negative
(@fig-radar; see supplementary material). The secondary axis had smaller
loadings in the variation and cumulative DHIs, with only the Southern
BEC group having a lower variation DHI loading than cumulative DHI. RDA2
has lower predictor loadings (maximum of \~0.25) when compared to RDA1
(maximum of \~0.5), with infrequently significant loadings for canopy
cover and structural complexity (@fig-radar). Canopy height had much
larger primary attribute loadings in the Southern group, while the
Northern group had slightly larger structural complexity loadings
(@fig-radar; @fig-fcc C) Similar to RDA1, the modelled structural
attribute loadings were generally similar to one another, although gross
stem volume was slightly higher in some southern BEC zones (@fig-radar;
@fig-fcc D).

```{r}
#| label: fig-fcc
#| fig-cap: False colour maps by Biogeoclimatic Ecosystem Classification (BEC) zone of axis loadings for the first redundancy analysis (RDA) axis (A and B) and second RDA axis (C and D). Colour values are normalized to the maximum loading of each variable. A and C show axis loadings for canopy height (CH; red), canopy cover (CC; green) and structural complexity (SC; blue). B and D show axis loadings for basal area (BA; red), total biomass (TB; green) and gross stem volume (GSV; blue). Black indicates no significant loadings, while greyscale colours indicate similar loadings across the three variables. Annotated colour wheels for each set of variables also shown.

knitr::include_graphics(here::here("outputs", "maps_colourwheel.png"))
```

# Discussion

Relationships between forest ecosystem structure and function are
increasingly being examined due to advances in data collection methods
[@ali2019; @radeloff2019; @ali2016]. Prior studies have shown both
theoretical [@knyazikhin1998] and empirical [@atkins2018] relationships
between forest ecosystem structure and function. We examined these
relationships across the province of BC, Canada, which has significant
environmental gradients [@valentine1978soil; @pojar1987] by using
wall-to-wall lidar-derived forest structural attributes [@matasci2018;
@matasci2018b] and intra-annual satellite-derived summaries of
vegetation productivity [@radeloff2019;
@razenkovaMediumresolutionDynamicHabitatInPress]. We found a weak
relationship between in forest ecosystem structure and function overall
across BC (15.7% of variation explained in the DHIs by forest structural
attributes), as well as within individual BEC zones and forest types
found in BC (@fig-fracts-bar).

Ecosystem structure and function are key EBV classes which are capable
of being measured using satellite remote sensing at global extents and
with large temporal depth [@pereira2013; @skidmore2021]. Both EBV
classes have been related to biodiversity metrics [@radeloff2019;
@bergen2009]. One of the key attributes of the EBV framework is the
complementarity of the proposed datasets, therefore it is necessary to
determine if potential EBVs comprise unique or duplicate information
when compared with one another [@skidmore2021]. This can be especially
important when both datasets being examined were primarily derived using
data from the same series of satellites, in our case, the Landsat series
[@matasci2018; @matasci2018b;
@razenkovaMediumresolutionDynamicHabitatInPress]. Our findings show weak
relationships between these two classes, indicating that these datasets
are very suitable to be used in a fashion complementary to one another,
especially when examining ecosystem integrity, even with the same data
provenance [@hansen2021].

While we assess the complementarity of ecosystem structure and function,
a single pair of EBVs easily measurable at similar scales, other EBV
class combinations should also be assessed for complementarity. This is
especially relevant in cases such as ours, where remote sensing datasets
can proliferate to a number of EBVs in different classes
[@skidmore2021]. Other EBVs, such as those found in the community
composition and species population classes, will be inherently linked
due to being based on the same data. In addition, @skidmore2021
highlighted that many satellite-derived biodiversity products can fall
into multiple EBV classes simultaneously -- e.g., leaf area index being
in the ecosystem structure, ecosystem function, and species trait EBV
classes -- depending on the spatial scale of analysis and goal of the
monitoring system. As such, it is relevant to consider which classes
each biodiversity product should belong to, for given extents and
temporal depths.

Further, we assessed the utility of primary forest structural attributes
when compared to modelled forest structural attributes to examine
duplicated information within a single EBV class. We found that standard
lidar-derived measurements, such as canopy height, canopy cover, and
structural complexity (as proposed by Valbuena et al., 2020) comprise
the majority of the explained variation in the DHIs, either
independently or in their overlap with the modelled attributes
(@fig-fracts-bar). This could be expected as modelled forest structural
attributes are generally directly based on the primary forest structural
attributes [@coops2021], however, the creation of modelled forest
structural attributes involves modelling expertise and processing time
overhead [@matasci2018; @matasci2018b; @duncanson2022]. One advantage of
these interpreted products is potential uptake by a broader set of
users, and their ability to be more closely linked to management actions
and monitoring results, as is the case with forest biomass for climate
change mitigation [@duncanson2022].

In the axis loadings we found a strong linkage between canopy cover,
structural complexity, and the DHIs, which was expected as these metrics
are more directly related to photosynthesis than canopy height
[@myneni1994]. We also found that axis loadings for modelled forest
structural attributes were more similar to canopy cover and structural
complexity than canopy height (@fig-rda-var). This was unexpected as
while the modelled forest structural attributes are generally correlated
due to their creation methods, these modelled attributes are primarily
generated using canopy height information [@duncanson2022], with
relationships often based on tree allometry rather than leaf volume or
placement [@chave2005; @lutz2018].

Forest types with deciduous trees (namely broadleaf and mixed wood) had
large, positive structural complexity loadings when compared to the
other forest structural attributes (@fig-radar), with mixed wood forests
having no modelled structural attributes associated with the DHIs. This
structural complexity-DHIs linkage may be due to understory cover
providing additional photosynthesis in these forest types
[@nilsson2005]. Coniferous forests DHIs, on the other hand, show a
weaker relationship with structural complexity than all other forest
structural attributes. Notably, the Alpine, Montane, and Northern BEC
groups show decoupling of the three DHIs (@fig-radar), with lower
loadings between the Minimum DHI and RDA1. This decoupling in the DHIs
is potentially due to the the near certainty of a pixel containing snow
cover over the course of a year leading to consistently low minimum DHI
values across the ecosystem and thus to weaker minimum DHI loadings
[@razenkova2020; @beck2006]. These ecosystems with DHI decoupling also
showed relatively strong relationships to canopy cover and weak
relationships with structural complexity, potentially due to the
dominance of coniferous forests in these ecosystems.

BC shows varying forest structural drivers of the DHIs depending on the
geographic location [@fig-fcc]. The interior of BC is drier and warmer,
and also has productivity being strongly driven by canopy cover, while
the Coastal zones are the only BEC zone with strong structural
complexity loadings. The climate of these regions could be influencing
the DHIs [@perez2016], leading to varying loadings strengths depending
on the distance to the coastline. Further, the extreme environmental and
topographic gradients within BC's ecosystems [@pojar1987;
@valentine1978soil] may be leading to highly variable data within the
BEC zones we used for this study, and a finer ecoregionalization may
reveal a different relationship between ecosystem structure and
function.

Across British Columbia, the DHIs have been shown to be strongly
correlated with one another (@fig-rda-var). While this does imply that
utilizing a single one of the three DHI axis may be suitable within this
geographic region, the the DHIs do provide unique information to one
another, which can be especially relevant when stratifying the study
area. High levels of minimum DHI, for example, may indicate winter
forage availability, which is difficult to capture in the variation or
cumulative DHIs [@razenkova2020]. The variation DHI has been shown to be
strongly related to seasonality, a key predictor of avian species
richness, especially in migratory species [@rowhani2008]. In Coastal
ecosystems, which are characterized by mild winters and high
precipitation, we found that forest structure and the DHIs are strongly
decoupled, with very little variance explained in the DHIs by the forest
structural attributes (@fig-fracts-bar). This indicates that the
complementarity of the DHIs and forest structural attributes should
remain, even when expanding the analysis outside of ecosystems with
little to no snow cover. Future research should continue to examine the
complementarity of various EBV pairs, across varying ecosystems and
geographic locations.

# Conclusion

We used redundancy analysis and variation partitioning to assess the
complementarity of two potential EBV datasets - forest structure and the
DHIs. We also separated the forest structure datasets into primary and
modelled structure variables in order to assess the need to develop more
complex structure variables, or if data derived directly from lidar
datasets are suitable. We found that the structure variables were not
strongly related to the DHIs, indicating that they are suitable to be
used together as complementary ecosystem-scale EBVs when monitoring
forest environments. It was also found that variation in the DHIs
explained by the overlap between primary and modelled structure
variables was often higher than the variation explained by either
individually, both in the BC-wide dataset and the stratified datasets.
These ecosystem structure and function variables are often attainable at
global scales using satellite remote sensing, and recent advances are
allowing them both to be generated at medium spatial resolutions. We
suggest that biodiversity researchers focus on using forest structural
attributes derived directly from the lidar data, and if needed, use a
single modelled forest structural attribute, such as total aboveground
biomass. In addition, we highlight that the usage of intra-annual
summarizations of productivity provide novel information for
biodiversity monitoring when used with forest structural attributes, and
utilizing them as complementary EBVs for a holistic monitoring system
for forest ecological integrity.

# Acknowledgements

This research was funded by NSERC support of Coops (RGPIN-2018-03851).
Remote sensing data products utilized in this research are free and open
and available for download at <https://ca.nfis.org/maps_eng.html>. We
thank Dr. Michael Wulder and Dr. Joanne White for development and early
access to these National Terrestrial Ecosystem Mapping System (NTEMS)
products. We thank Dr. Elena Razenkova for early access to the
Landsat-derived Dynamic Habitat Indices.

\newpage

# References {.unnumbered}