### 2023-11-08 scripts/manuscript/manuscript.qmd

---
title: Disentangling linkages between satellite-derived indicators of forest structure and productivity for ecosystem monitoring.
author:
  - name: Evan R. Muise
    email: evan.muise@student.ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
    attributes:
        corresponding: true
  - name: Margaret E. Andrew
    email: M.Andrew@murdoch.edu.au
    affiliations: 
        - id: murd
          name: Murdoch University
          department: Environmental and Conservation Sciences and Harry Butler Institute
          address: 90 South St
          city: Murdoch, WA, Australia
          postal-code: 6150
  - name: Nicholas C. Coops
    email: nicholas.coops@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
  - name: Txomin Hermosilla
    email: txomin.hermosillagomez@NRCan-RNCan.gc.ca
    affiliations: 
        - id: cfs
          name: Natural Resources Canada
          department: Canada Forest Service (Pacific Forestry Centre)
          address: 506 Burnside Rd W
          city: Victoria, BC, Canada
          postal-code: V8Z 1M5
  - name: A. Cole Burton
    email: cole.burton@ubc.ca
    affiliations: 
        - id: ubc
          name: University of British Columbia
          department: Forest Resources Management
          address: 2424 Main Mall
          city: Vancouver, BC, Canada
          postal-code: V6T 1Z4
  - name: Stephen S. Ban
    email: Stephen.Ban@gov.bc.ca
    affiliations: 
        - id: bcp
          name: Ministry of Environment and Climate Change Strategy
          department: BC Parks
          address: 525 Superior Street
          city: Victoria, BC, Canada
          postal-code: V8V 1T7
keywords: 
  - remote sensing
  - landsat
  - forest structure
  - forest productivity
  - dynamic habitat indices (DHIs)
  - essential biodiversity variables
  - redundancy analysis (RDA)
date: last-modified
bibliography: [bibliography.bib, packages.bib]
csl: scientific-reports.csl
format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    journal:
      name: Nature Scientific Reports
      formatting: review
      model: 3p
      cite-style: authoryear
      layout: twocolumn
  docx: 
    reference-doc: my-styles.docx
    keep-md: true
  pdf: 
    pdf-engine: xelatex
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: false
fig-width: 300
number-sections: true
nocite: |
  @matasci2018, @matasci2018b
---

```{r}
#| warning: false
#| echo: false
#| output: false

# include all libraries required by the R chunks here
library(vegan)
library(sgsR)
library(terra)
library(sf)
library(gt)
library(tidyverse)

library(bcmaps)

# generate package citations
loaded_pkgs = (.packages())
base_pkgs = getOption("defaultPackages")
contrib_pkgs = loaded_pkgs[!(loaded_pkgs %in% base_pkgs)]
pkgs_file = here::here("scripts", "manuscript", "packages.bib")
knitr::write_bib(contrib_pkgs, pkgs_file)

bec_tab <- read_csv(here::here("data", "climate.csv")) %>%
  mutate(across(.cols = c(tmax, tmin), \(x) round(x, digits = 1)))
```

\newpage

**Abstract**

The Essential Biodiversity Variables (EBV) framework has been proposed
as a monitoring system of standardized, comparable variables that
represents a minimum set of biological information to monitor
biodiversity change at large spatial extents. Six classes of EBVs
(genetic composition, species populations, species traits, community
composition, ecosystem structure and ecosystem function) are defined, a
number of which are ideally suited to observation and monitoring by
remote sensing systems. We used moderate-resolution remotely-sensed
indicators representing two ecosystem level EBV classes (ecosystem
structure and function) to assess their complementarity and redundancy
across a range of ecosystems encompassing significant environmental
gradients. Redundancy analyses found that remote sensing indicators of
forest structure were not strongly related to indicators of ecosystem
productivity (represented by the Dynamic Habitat Indices, DHIs), with
the structural information only explaining 16% of the variation in the
DHIs. Complex metrics of forest structure, such as above ground biomass,
did not contribute additional information over simpler height-based
attributes that can be directly estimated with light detection and
ranging (LIDAR) observations. With respect to ecosystem conditions, we
found that forest types and ecosystems dominated by coniferous trees had
less redundancy between the remote sensing indicators when compared to
broadleaf or mixed forest types. Likewise, higher productivity
environments exhibited the least redundancy between indicators, in
contrast to more environmentally stressed regions. We suggest that
biodiversity researchers continue to exploit multiple dimensions of
remote sensing data given the complementary information they provide on
structure and function focused EBVs, which makes them jointly suitable
for monitoring forest ecosystems.

\newpage

# Introduction

Monitoring the changing state of biodiversity globally is key to
understanding and mitigating the increased extinction risk for many
species [@thomas2004; @urban2015] and the homogenization of biotic
communities at various spatial and temporal scales
[@mcgillFifteenFormsBiodiversity2015]. The Group on Earth Observations
Biodiversity Observation Network has developed the Essential
Biodiversity Variables (EBVs) [@pereira2013], which are standardized,
comparable variables that represents a minimum set of biological
information required for monitoring biodiversity change at large spatial
extents [@skidmore2021]. The EBVs are designed as an analog to the
Essential Climate Variables framework [@bojinski2014] to be global in
scope, relevant to biodiversity information, feasible to implement and
use, and complementary to one another [@skidmore2021]. There are six
classes of EBVs collected at various scales. The ecosystem EBV classes
(ecosystem structure and ecosystem function) are well suited to be
monitored across large areas using remote sensing technologies
[@skidmore2021]. This has led to a proliferation of potential EBV
datasets that have been generated to describe these two ecosystem EBV
classes, especially in forested environments.

Ecosystem structure is the composition, abundance, and spatial
arrangement of different ecosystem components, and is indicative of
habitat quality and connectivity [@noss1990]. Forest ecosystem structure
is defined as the geographic distribution of trees or biomass, with
forest structural diversity encompassing the measurable morphological
components of the forests (e.g., canopy cover, canopy height, and
structural complexity) [@atkins2023; @valbuena2020]. Forest structural
diversity has been linked to species richness at various spatial scales,
ranging from individual plots to landscapes [@bergen2009; @gao2014]. A
variety of forest structural attributes can be extracted directly from
lidar (light detection and ranging) data, ranging in complexity from
simple (e.g., canopy cover, canopy height) to more complex (e.g.,
foliage height diversity, leaf area index), while others can be
indirectly modelled from lidar and auxiliary data (e.g., aboveground
biomass, basal area) [@coops2021]. A suite of these, and other,
structural attributes have been used as indicators of biodiversity at
local to landscape scales [@lefsky1999; @guo2017; @coops2016;
@bergen2009]. Increased forest structural complexity has been
hypothesized to create additional niches, leading to increased species
diversity [@bergen2009; @heidrich2020], which has been frequently
demonstrated using avian species diversity metrics [@macarthur1961] and
lidar [@clawges2008], or a combination of lidar and satellite imagery
[@herniman2020], with some research indicating that canopy vertical
distribution was the strongest predictor of species richness
[@goetz2007]. Although lidar datasets have traditionally been local in
extent, advances in satellite remote sensing processing have allowed the
extension of three-dimensional forest structure attributes nationally in
a wall-to-wall fashion. The computation of these attributes often
requires data fusion approaches involving lidar data and medium spatial
resolution optical imagery [@coops2021; @matasci2018; @matasci2018b].

Ecosystem functions are measures of ecosystem performance that are the
consequence of one or multiple ecosystem processes [@pettorelli2018;
@ecosyste2005]. Within the EBV framework, vegetation phenology --- the
timing and duration of leaf-on conditions --- and ecologically available
energy --- the amount of accessible and useable energy available for
growth in an ecosystem --- are useful predictors of species richness and
abundances at various scales [@chase2002; @radeloff2019; @coops2019;
@razenkova2023]. Ecologically available energy is the amount of energy
produced in an ecosystem, which is sensitive to limitations such as
recent disturbances and the water and nutrient supplies, and is
measurable via productivity [@bonn2004; @nguyen2022]. This differs from
ambient energy in that it goes beyond the climatic conditions of an
area, and concerns the amount of resources available for consumption in
an ecosystem [@hawkins2003]. Clear links have been developed between
remote sensing-derived spectral reflectance, particularly in the
near-infrared and the red parts of the electromagnetic spectrum
[@rouse1974], which relate to pigment concentrations and leaf structure,
respectively, as well as productivity [@yoder1994]. Consequently, the
normalized difference vegetation index (NDVI) and other vegetation
indices have been correlated with productivity measures [@huang2021;
@sims2006]. The relationship however, is typically non-linear,
principally due to saturation in the spectral response at moderate to
high leaf area index values [@myneni1994]. As a result, the strength of
the relationship between spectral indices and productivity depends on
the ecosystem type, land cover type, amount of canopy cover, and
productivity in the region [@shi2017]. Likewise, relationships between
productivity and biodiversity also contain uncertainty, recognizing that
biodiversity is governed by a variety of factors of which productivity
is but one [@skidmore2021]. Despite these uncertainties, remotely sensed
vegetation indices have enabled the assessment of biodiversity patterns
at single time points [@bonn2004; @nagendra2010; @phillips2008], or
through time when using time series data [@berry2007; @radeloff2019;
@coops2019]

While biodiversity is not solely driven by productivity, the
relationship between productivity and biodiversity can be explained
through various hypothesized mechanisms: including available energy,
environmental stress, and environmental stability. These three
hypotheses have been used to guide the development of a set of remote
sensing indicators, the Dynamic Habitat Indices (DHIs), which summarize
intra-annual dynamics in productivity metrics and have been evaluated as
broad-scale biodiversity indicators globally for a variety of taxa
[@berry2007; @coops2008; @radeloff2019]. According to the available
energy hypothesis [@wright1983; @bonn2004] an increase in ecologically
available energy results in higher species richness. In the context of
the DHIs, the total energy available within a specific area is
quantified using the cumulative DHI, which sums each productivity
observation throughout the course of a year. The environmental stress
hypothesis [@currie2004] suggests that higher levels of minimum
available energy contribute to greater species richness. The minimum
DHI, derived from the lowest productivity observed over the course of a
year, indicates the level of available energy during stressful periods.
The environmental stability hypothesis [@williams2008] proposes that
lower energy variation throughout the year leads to increased species
richness due to reduced resource bottlenecks. The variation DHI captures
the coefficient of variation in a vegetation index over the course of a
year, reflecting the prevalence of resource bottlenecks. The DHIs have
previously been produced at continental to global extents using Moderate
Resolution Imaging Spectroradiometer (MODIS) imagery, while recent
research has focused on producing the DHIs at finer spatial resolutions
using multi-annual time series of Landsat imagery
[@razenkovaMediumresolutionDynamicHabitatInPress].

The DHIs have been shown to be correlated with aspects of biodiversity
across a variety of spatial scales and locations, including Moose
occurrence and abundance in Ontario, Canada [@WOS:000343392200016] and
across Russia [@razenkova2020], beta diversity in butterflies in Canada
[@andrew2012], avian diversity in the coterminous U.S.
[@WOS:000659187700008], Ontario [@coops2009bird] and France
[@bonthoux2018], and global alpha diversity [@radeloff2019]. The DHIs
have also been used to construct novel ecoregionalizations [@andrew2013;
@coops2009]. DHIs, however, are not able to capture all components of
biodiversity due to the many environmental and landscape processes not
observable from satellite data [@coops2009bird].

The suite of remote sensing indicators used in biodiversity monitoring
should be carefully considered. Ideally, indicators should be selected
that are closely related to the phenomenon of interest to allow for
direct integration of monitoring results with management actions
[@pressey2021]. With the advent of large-extent monitoring methods like
satellite remote sensing which have led to a proliferation of potential
EBV datasets, it becomes important to assess the interrelationships
between these datasets and the complementary of the information they
provide to reduce redundancy in future EBV development efforts
[@pereira2013; @skidmore2021]. Identifying and exploring the strong
relationships that may exist between indicators of EBVs may help
elucidate the ecological relationships between these facets of
biodiversity, such as ecosystem structure and function, at local to
global scales. On the other hand, when remote sensing-derived EBVs are
not related, they may be well suited to be used in monitoring programs
together, as complementary sets of observations.

The relationship between ecosystem structure and function have been
examined within a remote sensing context for decades [@huete2002;
@knyazikhin1998; @myneni1994]. Hypothesized mechanisms such as niche
complementarity suggest that productivity increases with stand
structural complexity [@zhang2012], while asymmetric competition for
light can reduce forest productivity with increased structural
complexity [@bourdier2016]. Some aspects of forest structure, such as
canopy cover and leaf area index, also directly influence vegetation
indices, which are used to estimate productivity [@pettorelli2005].
Often, these large-extent estimates of structure and function metrics
are generated from similar data sources, such as the Landsat series of
satellites [@razenkovaMediumresolutionDynamicHabitatInPress ;
@radeloff2019; @matasci2018; @matasci2018b]. Due to this shared data
provenance, and the linkages between biodiversity and both structure and
function, it is important to assess how much unique information these
structure and function metrics represent when used as indicators of
biodiversity.

The goal of this study is to assess the complementarity of two remote
sensing datasets often proposed as indicators of ecosystem structure and
function EBVs. We interpret weak or absent relationships between metrics
to constitute complementary information. To do so, we synthesize
wall-to-wall forest structure data previously generated by modelling
lidar-derived forest structure metrics with medium spatial resolution
satellite imagery using a k-NN imputation method [@matasci2018;
@matasci2018b], alongside a well-established remote sensing-derived
index of ecosystem function (DHIs) [@coops2008; @berry2007;
@radeloff2019; @razenkovaMediumresolutionDynamicHabitatInPress]. First,
we analyze remote sensing indicator complementarity by examining the
multivariate relationship between ecosystem structure and function.
Second, we propose a method for selecting which remote sensing
attributes to develop and apply within a single EBV class by determining
the independent and shared contributions of lidar-derived forest
structural attributes and modelled forest structure to explaining
ecosystem function. Third, we examine the global applicability of these
results by evaluating the consistency of the observed forest
structure-function relationships across a broad range of environment
gradients and forest types.

# Methods and Materials

## Study Area {#sec-study-area}

British Columbia (BC) is the westernmost province of Canada, and is home
to a variety of terrestrial ecosystems. Located between the Pacific
Ocean and the Rocky Mountains, approximately 64% of the province is
forested, with large environmental and topographic gradients
[@pojar1987; @bcministryofforests2003]. The majority of the forests of
BC are coniferous forests (87.4%), with the remaining forests being
broadleaf (8.9%), wetland-treed (2.7%) and mixed-wood (0.9%)
[@hermosilla2022]. There is a large topographic gradient in the
province, with elevations ranging from sea level to over 4000 m, with
much of the province being mountainous (elevations above 1000 m)
[@valentine1978soil]. Wildfires are a regular disturbance in the
province, although larger, higher intensity fires have become more
common in recent years due to disruption of the historical fire regime
[@brookes2021]. In addition, forestry is a large industry in the
province, with approximately 136,000 ha of the province harvested
accounting for \~54 million cubic metres of volume of wood in 2021
[@canadiancouncilofforestministers2023].

BC is stratified into 16 zones based on the dominant tree species and
climate by the Biogeoclimatic Ecosystem Classification (BEC) system
(@fig-study-map). To examine trends across the large environmental
gradients, we group the BEC zones into five broad biomes: the Northern
interior, Southern interior, Montane, Alpine, and Coastal groups,
similar to Hamann et al. [@hamann2006]. We also report the average
climate data for BEC zones from 1991-2020, according to Wang et al.
[@wang2016] (Table 1).

```{r}
#| label: fig-study-map
#| fig-cap: Location of Biogeoclimatic Ecosystem Classification (BEC) zones in British Columbia.

knitr::include_graphics(here::here("outputs", "bec_map.png"))
```

```{r}
#| warning: false
#| echo: false

bec_tab %>%
  gt() %>%
  cols_label(
    group = "BEC Group",
    zone = "BEC Zone",
    zone_nm = "Full Name",
    ppt = "Average Monthly Precipitation (mm/yr)",
    tmax = "Average Monthly Max Temperature (°C)",
    tmin = "Average Monthly Min Temperature (°C)"
  ) %>%
  tab_header(title = "Biogeoclimatic Ecosystem Classification (BEC) Zones, their aggregated biomes, and their monthly average climate values for precipitation, maximum temperature, and minimum temperature. Climate data from Wang et al. 2016 averaged for 1991-2020.")
```

## Data

### National Terrestrial Ecosystem Monitoring System data products

#### Surface-reflectance image composites

The forest structure, forest disturbance, and land cover layers are
based on a 30-m best-available-pixel (BAP) composite generated by
Hermosilla et al. @hermosilla2016 using the Composite2Change approach.
The composites were developed for each year from 1984-2019 by selecting
the best available imagery from the available growing season Landsat
observations, removing pixels with clouds, cloud shadows, and haze. The
methodology uses the scoring method from White et al. @white2014 to
select the best pixel for a given year from Landsat-5 Thematic Mapper,
Landsat-7 Enhanced Thematic Mapper Plus, and Landsat-8 Operational Land
Imager imagery. A spectral trend analysis was conducted over these
initial BAP composites using the Normalized Burn Ratio (a vegetation
index) on each pixel, to remove unscreened noise, detect changes, and
temporally interpolate data gaps. This resulted in a gap-free
surface-reflectance composite across Canada [@hermosilla2015]. We use
data from 2015 for all National Terrestrial Ecosystem Monitoring System
products, except forest disturbances, which comprise the period
1985-2020.

#### Forest Disturbances

We used the forest disturbance layer generated by Hermosilla et al.
@hermosilla2016 to restrict the analysis to stands older than 35 years
as productivity and structure metrics have been shown to be strongly
decoupled in early regenerative stands [@bolton2017], and to exclude
recently anthropogenically disturbed stands. The disturbances detected
using spectral trend analysis were attributed to a disturbance agent
(i.e., wildfire, harvest) within an object-based analysis approach using
a Random Forests classification model [@hermosilla2015a]. Changes were
detected with an overall accuracy of 89%, attributed with an accuracy of
92%, and assigned to the correct year with an accuracy of 89%
[@hermosilla2016].

#### Land Cover {#sec-lcc}

We used the land cover map for 2015 generated by Hermosilla et al.
@hermosilla2022 to stratify our samples alongside the BEC zones. The
30-m spatial resolution land cover map for Canada's forested ecosystems
was generated using the Virtual Land Cover Engine framework
[@hermosilla2018]. This framework uses a Random Forests based
classification approach, and integrates logical land cover transitions
using a Hidden Markov Model to generate 12 land cover classes. The land
cover classification reached an overall accuracy of 77.9% ± 4%
[@hermosilla2022]. We restricted our sample to the four treed land cover
classes: broadleaf, coniferous, mixed wood, and wetland-treed.

#### Forest Structure

The forest structural attributes (canopy height, canopy cover,
structural complexity \[coefficient of variation of height returns\],
basal area, aboveground biomass, and gross stem volume) were derived
from the 30-m spatial resolution dataset generated by Matasci et al.
[@matasci2018; @matasci2018b]. In brief, the method extends structural
attributes derived from a set of lidar acquisitions and field plots
across Canada. These data were generated for the treed pixels of Canada
using the BAP Landsat surface reflectance composite and auxiliary data
(i.e., topography, geography) using a k-Nearest Neighbour imputation
[@matasci2018; @matasci2018b]. Accuracy metrics for imputed forest
structural attributes range from an RMSE of 24.5% (structural
complexity) to 82.3% (gross stem volume) and *R^2^* ranging from 0.125
(structural complexity) to 0.712 (gross stem volume) [@matasci2018;
@matasci2018b].

### Dynamic Habitat Indices

The DHIs were calculated following Razenkova et al.
@razenkovaMediumresolutionDynamicHabitatInPress for the terrestrial area
of BC. In brief, we used Google Earth Engine [@gorelick2017] to obtain
valid Landsat pixels for the study area across a ten-year time span,
centered on 2015 (2011-2020), filtering out pixels containing shadows,
clouds, and cloud shadows using the QA band derived from the fmask
algorithm [@zhu2012], and then calculated the NDVI index. These
estimates were composited into a synthetic year of monthly data by
calculating the monthly median NDVI value. Finally, images of the sum,
minimum, and coefficient of variation across this synthetic year of NDVI
values were then calculated to represent the cumulative, minimum, and
variation DHIs, respectively.

A summary of the datasets used in this study is presented in Table 2.

## Sampling

To obtain representative sample units of the EBV metrics across the
study area, we implemented stratified random sampling across the union
of BEC zones (Table 1) and forest types (i.e., coniferous, broadleaf,
mixed wood, and wetland treed; see @sec-study-area) of BC. Sample units
were restricted to forested pixels that were surrounded by the same
forest type to reduce the chance of sampling mixed pixels. We also
required that sample units were selected from a homogeneous area, to
reduce uncertainty associated with the input datasets [@shang2020];
homogeneous areas were defined as 3x3-pixel windows with a coefficient
of variation for canopy cover and canopy height below 0.5. A minimum
sampling distance of 1-km was implemented to reduce the effects of
spatial autocorrelation. Sample units that had been disturbed in the
last 35 years were discarded by using the disturbance mask generated for
the forested ecosystems of Canada [@hermosilla2016 ] . A maximum of 500
sample units were selected from each stratum; in the many strata where
this was not achieved, all sampling units meeting the above restrictions
were selected. Sampling was conducted in R version 4.2.2 \[@R-base\]
using the **sgsR** package [@R-sgsR]. Neighbourhood analyses for the
land cover classes and coefficient of variations of canopy height and
cover were calculated in Python version 3.9.

## Analysis

Redundancy analysis (RDA) and variation partitioning were used to relate
the primary and modelled forest structure variables to ecosystem
function across a broad environment range. RDA has widely been used in
community ecology where environmental variables of interest are compared
to species composition [@blanchet2014; @kleyer2012]. RDA predicts
multiple response variables by first running a multiple linear
regression of the predictor variables on each response variable, and
then a principal component analysis on the fitted values from each
multiple linear regression. This reduces the dimensionality of the
fitted values by transforming them into a set of independent RDA axes,
and allows the strength of the multivariate relationship between
predictors and responses to be assessed by calculating eigenvalues of
the RDA. Relationships between variables are revealed by the loadings of
both predictor and response variables on the RDA axes. RDA axes are
labelled as RDA1, RDA2, etc. to indicate they are derived from
redundancy analysis, rather than other ordination techniques such as
principal component analysis or constrained correspondence analysis
[@legendre2012]. Partial RDA extends RDA by allowing groups of predictor
variables to be considered as co-variates. Variation partitioning
leverages multiple partial RDAs to assess the overlap between the
explanatory power of the groups of predictor variables [@legendre2012].

We performed RDAs for our full sample, pooling BEC zones and forest
types, as well as individually by BEC zone and forest type, with results
aggregated to BEC zone groups (Table 1) . All measured values were
natural-log transformed and standardized to Z-scores based on the
summary statistics of the full sample prior to analyses. Following the
data transformation, we visually assessed the pairwise relationships
between the overall and stratified datasets for the assumptions of
linearity and homoskedasticity. Forest structure variables were treated
as predictor variables and were grouped based on whether they were
primary estimates from the lidar data or modelled from lidar and
ancillary information (Table 2) for variance partitioning. The three
DHIs were used as the response variables. Analysis of variance tests
were used to determine which RDA axes were significant, with p-values
below 0.1 using an F-test. The proportion of variance attributable to
each axis was calculated from the eigenvalues generated by the RDA, and
axis loadings were calculated as the Pearson correlation between each
forest structure or DHI variable and the RDA axes. To ensure similar
directionality (+ or -) in the loadings across each variable in every
strata, we transformed the data such that the cumulative DHI loading was
always positive. To do so, if the cumulative DHI loading was negative we
multiplied each loading in that strata and axis by -1. Forest structural
attribute axis loadings represent the strength of relationship between a
given variable and the RDA axis, while the DHI loadings indicate what is
being represented by the RDA axes. To visualize the RDA results, we plot
the significant loadings as annotated arrows between the
predictor/response variables and each RDA axis (shown as a box), similar
to a path diagram. We also display the amount of variance explained by
each RDA axis within their respect

Variation partitioning is displayed using a Venn diagram in which the
percentage of variance explained by each dataset is in a circle, and the
overlap between circles represents the overlap in variance explained. We
add paths to the Venn diagram to indicate how the structural attributes
contribute to the variation in the DHIs. Because the variation
partitioning results summarize the full multivariate solution, rather
than individual RDA axes, we present variation partitioning results
alongside the path diagrams illustrating the details of the RDA results.
RDA and variation partitioning calculations were done in R [@R-base]
version 4.2.2 using the **vegan** package [@R-vegan]. The code
associated with the processing and analysis is available at
https://github.com/emuise/code-structProdSem.

# Results

Overall in BC, the intra-annual patterns of productivity as represented
by the DHI variables were weakly related to forest structure
(@fig-rda-var). Less than 16% of the overall variation in the DHIs was
explained by the forest structural attributes. Due to strong
correlations between the axes of the DHIs, the majority of the variation
was held within the first RDA axis, which had strong loadings for all
DHIs, with positive loadings for the cumulative and minimum DHIs, and
negative loadings for the variation DHI. As such, we identified the RDA1
axis as representing overall productivity. The overall productivity axis
had similar loadings (between 0.24-0.27) for all modelled forest
structural attributes and canopy cover. Canopy height had a smaller
loading (0.1), while structural complexity was negatively correlated
with the overall productivity axis. RDA2 explained much less of the
variation in the DHIs. RDA2 had a large positive loading on the minimum
DHI (0.55), with negative loadings on the cumulative (-0.19) and
variation (-0.39) DHIs. Due to the minimum and variation having the most
influence on RDA2, we identified it as a seasonality axis. Canopy cover
and structural complexity had the strongest (albeit negative; -0.17 and
-0.08, respectively) loadings on the seasonality axis (@fig-rda-var A).
Variation partitioning highlighted that the majority of the explained
variation in the DHIs was due to the primary forest structural
attributes (canopy cover, canopy height, and structural complexity; 9%
variation explained). The overlap between primary and modelled
attributes was 3.5%, and the modelled attributes explained 3.2% of the
variation on their own (@fig-rda-var B).

```{r}
#| label: fig-rda-var
#| fig-cap: A) Axis loadings from redundancy analysis (RDA) of primary and modelled forest structure variables on the dynamic habitat indices (DHIs). B) Results from variation partitioning of primary and modelled forest structure variables on the DHIs. Both visualized analyses are across the entire dataset. See supplementary information for results from each Biogeoclimatic Ecosystem Classification zone and forest type.

knitr::include_graphics(here::here("outputs", "British Columbia.drawio.png"))
```

By forest type, results indicated similar or increased amounts of
variance explained when compared to the overall dataset
(@fig-fracts-bar). By BEC zones results generally indicated smaller
amounts of variation explained, which may be due to the relatively short
environmental gradients within the zones. The overlap between the
primary and modelled datasets was often responsible for the majority of
the variation explained in the DHIs, except when stratifying by forest
types, where this was only true for the coniferous forests. For
broadleaf, mixed wood, and wetland-treed forest types most of the
explained variation in the DHIs was contributed by the primary forest
structural attributes. A similar pattern was found by BEC zones, where
the variation in the DHIs was principally explained by the overlap
between primary and modelled forest structural attributes, except for
CWH, where no variation was explained by the overlap. The Coastal BEC
group had the lowest amount of variation explained, with Coastal
Douglas-fir having \<1% of variation explained by the structural data.
Overall, the DHIs were decoupled from the forest structural attributes,
as shown by the overall variation explained being under 30% regardless
of data stratification.

```{r}
#| label: fig-fracts-bar 
#| fig-cap: Stacked bar plot proportion of the variation in the Dynamic Habitat Indices (DHIs) variation explained by primary, modelled, and the overlap between primary and modelled structure variables across the stratified and overall datasets.  
knitr::include_graphics(here::here("outputs", "fracts_bar_plot.png"))
```

Stratification by forest type and ecosystem indicated varying loadings
between the RDA axes and predictor/response variables (@fig-radar). All
DHI variables had high loadings on RDA1 (similar to the overall
dataset), with positive cumulative and minimum DHI loadings, and
negative variation DHI loadings. Colder BEC groups (Montane, North, and
Alpine) had smaller minimum DHI loadings, which were half as large as
the other DHI loadings. As such, we identify the first axis as being
consistently associated with overall productivity. Canopy cover
generally had the highest significant predictor loading in RDA1, except
for in non-coniferous forest types (broadleaf, mixed wood and
wetland-treed). These forest types typically had large negative loadings
between RDA1 and structural complexity. The mixed wood forest and
Coastal BEC group had no significant loadings from modelled forest
structural attributes for RDA1. Examining the predictor loadings for
RDA1 spatially across the province shows a geographic pattern in the
loading strengths of primary forest structural attributes (@fig-fcc A).
The interior (Northern and Southern BEC groups) of the province
generally showed high canopy cover loadings, while the Boreal zones in
the northwest had equal loadings in canopy cover and canopy height.
Coastal zones were driven similarly by canopy height and structural
complexity, with very low or non-significant loadings from canopy cover
(@fig-fcc A; @fig-radar). Modelled forest structural attributes for RDA1
generally show similar loadings across the province (@fig-fcc B).

```{r}
#| label: fig-radar
#| fig-cap: Parallel coordinate plots of average absolute significant loading strength by Biogeoclimatic Ecosystem Classification (BEC) group, forest type, and British Columbia. Note the varying y axes.

knitr::include_graphics(here::here("outputs", "parcoord_noflip_edit.png"))
```

Only seven of sixteen BEC zones had a second RDA axis, with none being
found in Coastal or Alpine BEC zones (@fig-radar; @fig-fcc). All of the
Montane BEC zones had a secondary axis, while some in the Northern and
Southern interior had a secondary axis. RDA2 often showed the largest
absolute loadings in the minimum DHI, which was frequently, but not
always, negative (@fig-radar; see supplementary material). The montane
group, which had the largest absolute minimum DHI loading, also had the
largest absolute variation DHI loading. The secondary axis had smaller
loadings in the variation and cumulative DHIs, with only the Southern
BEC group having a lower variation DHI loading than cumulative DHI. RDA2
has lower absolute predictor loadings (maximum of \~0.25) when compared
to RDA1 (maximum of \~0.5), with infrequently significant loadings for
canopy cover and structural complexity (@fig-radar). Among the primary
forest structural attributes, RDA2 had the highest loadings with canopy
height in the Southern group, while the Northern group had slightly
larger structural complexity loadings (@fig-radar; @fig-fcc C). Similar
to RDA1, the modelled forest structural attribute loadings were
generally similar to one another (@fig-radar; @fig-fcc D).

```{r}
#| label: fig-fcc
#| fig-cap: False colour maps by Biogeoclimatic Ecosystem Classification (BEC) zone of axis loadings for the first redundancy analysis (RDA) axis (A and B) and second RDA axis (C and D). Colour values are normalized to the maximum loading of each variable. A and C show axis loadings for canopy height (CH; red), canopy cover (CC; green) and structural complexity (SC; blue). B and D show axis loadings for basal area (BA; red), total biomass (TB; green) and gross stem volume (GSV; blue). Black indicates no significant loadings, while greyscale colours indicate similar loadings across the three variables. Annotated colour wheels for each set of variables also shown.

knitr::include_graphics(here::here("outputs", "maps_colourwheel.png"))
```

# Discussion

Relationships between forest ecosystem structure and function are
increasingly being examined due to advances in data collection methods
[@ali2019; @radeloff2019; @ali2016]. Prior studies have shown both
theoretical [@knyazikhin1998] and empirical [@atkins2018] relationships
between forest ecosystem structure and function, however, the direction
and significance of these relationships have not been consistent
[@ali2019]. We examined these relationships across the province of BC,
Canada, which has significant environmental gradients
[@valentine1978soil; @pojar1987] by using wall-to-wall forest structural
attributes generated by imputing the lidar-derived attributes with
Landsat best-available-pixel composites [@matasci2018; @matasci2018b]
and Landsat-derived summaries of intra-annual vegetation productivity
[@radeloff2019; @razenkovaMediumresolutionDynamicHabitatInPress]. We
found a weak relationship between these remote sensing indicators of
forest ecosystem structure and function overall across BC (15.7% of
variation explained in the DHIs by forest structural attributes), as
well as within individual BEC zones and forest types found in BC
(@fig-fracts-bar).

Our findings indicate that remote sensing-derived forest ecosystem
structure and forest ecosystem function EBVs are well suited to be used
in a complementary fashion for ecosystem integrity monitoring, despite
both EBVs being derived from surface reflectance from the Landsat series
of satellites [@hansen2021]. While we assessed the complementarity of
ecosystem structure and function, a single pair of EBVs measurable at
the ecosystem scale, other EBV class combinations at finer scales should
also be assessed for complementarity. This is especially relevant in
cases such as ours, where remote sensing datasets such as Landsat can
generate a large number of EBVs in different classes [@skidmore2021]. In
addition, Skidmore et al. [@skidmore2021] highlighted that many
satellite-derived biodiversity products can fall into multiple EBV
classes simultaneously -- e.g., leaf area index being in the ecosystem
structure, ecosystem function, and species trait EBV classes --
depending on the spatial scale of analysis and goal of the monitoring
program. As such, it is relevant to consider which classes each
biodiversity product should belong to, for given spatial and temporal
extents. Other EBVs, such as those found in the community composition
and species population classes, may also be inherently linked due to
being based on the same data sources, even if not observed using remote
sensing platforms.

Further, we assessed the information-richness of primary forest
structural attributes when compared to modelled forest structural
attributes to support indicator selection within a single EBV class. We
found that standard lidar-derived measurements, such as canopy height,
canopy cover, and structural complexity (as proposed by Valbuena et al.
[@valbuena2020]) contribute to the majority of the explained variation
in the DHIs, either independently or in their overlap with the modelled
forest structural attributes (@fig-fracts-bar). This could be expected
as modelled forest structural attributes are generally directly based on
the primary forest structural attributes [@coops2021], however, the
creation of modelled forest structural attributes involves modelling
expertise and processing time overhead [@matasci2018; @matasci2018b;
@duncanson2022]. One advantage of these interpreted products is
potential uptake by a broader set of users, and their ability to be more
closely linked to management actions and monitoring results, as is the
case with forest biomass for climate change mitigation [@duncanson2022].
While either set of forest structural attributes may be most relevant
for a given monitoring task, we note that the modelled structural
attributes do not strongly overlap with structural complexity, and as
such, we generally recommend to use the primary forest structural
attributes.

In the axis loadings we found a strong linkage between canopy cover,
structural complexity, and the DHIs, which was expected as these metrics
are more directly related to photosynthesis than canopy height
[@myneni1994]. We also found that axis loadings for modelled forest
structural attributes were more similar to canopy cover and structural
complexity than canopy height (@fig-rda-var). This was unexpected as
while the modelled forest structural attributes are generally correlated
due to their creation methods, these modelled attributes are primarily
generated using canopy height information [@duncanson2022], with
relationships often based on tree allometry rather than leaf volume or
placement [@chave2005; @lutz2018].

Forest types with deciduous trees (namely broadleaf and mixed wood) had
large, negative structural complexity loadings when compared to the
other forest structural attributes (@fig-radar), with mixed wood forests
having no modelled structural attributes associated with the DHIs. This
structural complexity-DHIs linkage may be due to understory cover
providing additional photosynthesis in these forest types
[@nilsson2005]. In coniferous forests, on the other hand, the DHIs show
a weaker relationship with structural complexity than all other forest
structural attributes. Notably, the Alpine, Montane, and Northern BEC
groups show decoupling of the three DHIs (@fig-radar), with lower
loadings between the Minimum DHI and RDA1. This decoupling in the DHIs
is potentially due to the near certainty of a pixel containing snow
cover over the course of a year leading to consistently low minimum DHI
values across the ecosystem, and thus to weaker minimum DHI loadings
[@razenkova2020; @beck2006]. These ecosystems with DHI decoupling also
showed relatively strong relationships to canopy cover and weak
relationships with structural complexity. This weak relationship to
structural complexity may be due to the dominance of coniferous forests
in these ecosystems, which have been shown to have lower forest
structural complexity across the globe [@ehbrecht2021], with weaker
relationships between canopy height and structural complexity being
found [@atkins2022].

BC shows varying forest structural drivers of the DHIs depending on the
geographic location (@fig-fcc). The interior of BC is drier and warmer,
and also has productivity being strongly driven by canopy cover, while
the Coastal zones are the only BEC zone with strong structural
complexity loadings. The climate of these regions could be influencing
the DHIs [@perez2016], leading to varying loadings strengths depending
on the distance to the coast. For example, water availability has been
shown to be a driver of canopy height [@klein2015]. In these wetter
regions such as the Coastal group, structural variability is more likely
to be present in the vertical, rather than horizontal direction,
especially given the relationship between canopy height and structural
complexity [@atkins2022]. Further, the weakest relationships between
structure and function were found in the Coastal group of BEC zones
(@fig-fracts-bar), which may be due to the consistently high amounts of
canopy cover, and lower amounts of snow found in these zones leading to
potentially saturated NDVI values throughout the year. These consistent
forest structural metrics --- such as the high canopy cover values found
in the Coastal BEC zones --- may also lead to lower amounts of variation
explained by these metrics, leading to them being less informative in
the loadings, and other metrics, such as structural complexity having
higher amounts of variance explained (@fig-radar). These complex
linkages between climate, forest structure, and forest productivity
become further complicated by the extreme environmental and topographic
gradients within BC's ecosystems [@pojar1987; @valentine1978soil], which
may warrant study into the scale of these interactions, with a finer (or
non-climate based) ecoregionalization potentially revealing different
relationships between ecosystem structure and function.

Various relationship directions and strengths have been found between
forest structure and forest function [@ali2019]. Temperate and boreal
forests often have mixed or negative structure-function relationships,
while tropical forests often have positive relationships [@ali2019]. We
found that forest structure and function were not strongly linked to one
another, both across the full study area, and when stratified by
ecosystem and forest type (@fig-radar). Forest type models showed higher
amounts of variation explained, which may indicate the linkages between
forest structure and function are stronger at finer spatial scales,
potentially driven by finer scale landscape patterns such as terrain and
land cover, rather than climate [@thom2021]. Notably, the BEC zones and
forest types are delineated at different spatial scales; forest types
are mapped at a 30 m pixel size [@hermosilla2022; @hermosilla2018], and
can be found across the entire province, while BEC zones do not span the
entirety of the province, and are mapped at a much coarser scale
[@pojar1987].

Across British Columbia, the DHIs have been shown to be strongly
correlated with one another (@fig-rda-var). While this does imply that
using a single one of the three DHI axis may be suitable within this
geographic region, the DHIs do provide unique information to one
another, which can be especially relevant when stratifying the study
area. High levels of minimum DHI, for example, may indicate winter
forage availability, which is difficult to capture in the variation or
cumulative DHIs [@razenkova2020]. The variation DHI has been shown to be
strongly related to seasonality, a key predictor of avian species
richness, especially in migratory species [@rowhani2008]. In Coastal
ecosystems, which are characterized by mild winters and high
precipitation, we found that forest structure and the DHIs are strongly
decoupled, with very little variance explained in the DHIs by the forest
structural attributes (@fig-fracts-bar). This indicates that the
complementarity of the DHIs and forest structural attributes should
remain, even when expanding the analysis outside of ecosystems with
little to no snow cover. Future research should continue to examine the
complementarity of various EBV pairs, across varying ecosystems and
geographic locations.

A potential limitation of our study is due to the uncertainty inherent
to remotely sensed data. Wall-to-wall forest structural information was
produced by imputing lidar-derived metrics over satellite spectral
reflectance and geospatial data across the forested landscapes of
Canada. The accuracy of the forest structure dataset varied between
forest structural metrics [@matasci2018; @matasci2018b]. Likewise,
spectral vegetation indices, as discussed, are not direct measurements
of productivity, but rather are based on surface reflectance of
vegetation, which can be influenced by many factors such as soil, canopy
closure, viewing geometry, and atmospheric effects [@myneni1994;
@yoder1994]. These uncertainties may weaken the relationship between
these remote sensing indicators of forest structure and function.
Further, while forest structure and function have both been shown to be
drivers of biodiversity [@radeloff2019; @atkins2023], there are
ecological processes that cannot be assessed through remotely sensed
data [@coops2009exploring], including predation and fecundity, and thus
*in situ* data are needed to monitor these processes. Finally, our study
assumes that structure and function are indicators of biodiversity in
this system, and we draw on published demonstrations of the relevance of
each of them to species richness and composition, but we recommend
further evaluations of these indicators with biodiversity observations
to determine the relative contributions of each to estimating species
richness and turnover [@coops2018a].

# Conclusion

We used redundancy analysis and variation partitioning to assess the
complementarity of two potential EBV datasets - forest structure and the
DHIs. We also separated the forest structure indicator datasets into
primary and modelled structure metrics in order to assess the need to
develop more complex structure variables, or if the primary forest
structural attributes are suitable on their own. We found that the
structure metrics were not strongly related to the DHIs, indicating that
they are suitable to be used together as complementary ecosystem-scale
EBVs when monitoring forest environments. It was also found that
variation in the DHIs explained by the overlap between primary and
modelled structure variables was often higher than the variation
explained by either individually, both in the BC-wide dataset and the
stratified datasets. These ecosystem structure and function indicators
are often attainable at global scales using satellite remote sensing,
and recent advances are allowing them both to be generated at medium
spatial resolutions. We suggest that biodiversity researchers focus on
using forest structural attributes derived directly from the lidar data,
and if needed, use a single modelled forest structural attribute, such
as total aboveground biomass. In addition, we highlight that the usage
of intra-annual summarizations of productivity provide novel information
for biodiversity monitoring when used with forest structural attributes
as complementary EBVs for a holistic monitoring system for forest
ecological integrity.

# Acknowledgements

This research was funded by NSERC support of Coops (RGPIN-2018-03851).
Remote sensing data products used in this research are free and open and
available for download at <https://ca.nfis.org/maps_eng.html>. We thank
Dr. Michael A. Wulder and Dr. Joanne C. White for development and early
access to these National Terrestrial Ecosystem Mapping System (NTEMS)
products. We thank Dr. Elena Razenkova for early access to the
Landsat-derived Dynamic Habitat Indices.

# Author Contributions

Conceptualization: ERM, MEA, NCC, ACB and SSB. Methodology: ERM, MEA,
NCC and TH. Software: ERM and MEA. Formal Analysis: ERM, MEA, and NCC.
Investigation: ERM. Writing - Original Draft: ERM. Writing - Review and
Editing: ERM, MEA, NCC, TH, ACB and SSB. Visualization: ERM, MEA, TH,
NCC. Supervision: NCC. All authors approve the submitted version of the
manuscript.

# Competing Interests

MEA is an Editorial Board Member of *Scientific Reports*. The authors
declare no other competing interests.

# Data Availability

The 30 m spatial coverages of the NTEMS data are available online at
https://opendata.nfis.org/mapserver/nfis-change_eng.html. BEC zone
boundaries were obtained using the *bcmaps* R package. The Dynamic
Habitat Indices were calculated using freely available Landsat imagery
in Google Earth Engine. Details on their production can be found in
Razenkova et al. [@razenkovaMediumresolutionDynamicHabitatInPress]. All
code used for data manipulation and analysis can be found at
https://github.com/emuise/code-structProdSem.

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

**Legends**

Figure 1: Location of Biogeoclimatic Ecosystem Classification (BEC)
zones in British Columbia.

Figure 2: A) Axis loadings from redundancy analysis (RDA) of primary and
modelled forest structure variables on the dynamic habitat indices
(DHIs). B) Results from variation partitioning of primary and modelled
forest structure variables on the DHIs. Both visualized analyses are
across the entire dataset. See supplementary information for results
from each Biogeoclimatic Ecosystem Classification zone and forest type.

Figure 3: Stacked bar plot proportion of the variation in the Dynamic
Habitat Indices (DHIs) variation explained by primary, modelled, and the
overlap between primary and modelled structure variables across the
stratified and overall datasets.

Figure 4: Parallel coordinate plots of average absolute significant
loading strength by Biogeoclimatic Ecosystem Classification (BEC) group,
forest type, and British Columbia. Note the varying y axes.

Figure 5: False colour maps by Biogeoclimatic Ecosystem Classification
(BEC) zone of axis loadings for the first redundancy analysis (RDA) axis
(A and B) and second RDA axis (C and D). Colour values are normalized to
the maximum loading of each variable. A and C show axis loadings for
canopy height (CH; red), canopy cover (CC; green) and structural
complexity (SC; blue). B and D show axis loadings for basal area (BA;
red), total biomass (TB; green) and gross stem volume (GSV; blue). Black
indicates no significant loadings, while greyscale colours indicate
similar loadings across the three variables. Annotated colour wheels for
each set of variables also shown.