### 2024-01-10 scripts/4_vegan_clean.qmd
began revisions round 2, cleaned 4_vegan to be easier to work with (less defunct extra code)

---
title: "vegan_analysis"
format: html
editor: visual
---

```{r packages}
library(vegan)
library(tidyverse)
library(cols4all)
library(terra)
library(sf)
library(patchwork)

source(here::here("scripts", "get_keys.R"))
```

```{r normal-fun}
mylog <- function(data) {
  if (any(data <= 0)) {
    return(log1p(data))
  } else {
    return(log(data))
  }
}

# normalization from grace et al 2016
# first does mylog (does natural logarithm. if 0s present, does log(one + x))
# then scales to z scores
# results it more normal looking data
normal <- function(data) {
  data %>%
    mylog() %>%
    scale() %>%
    as.numeric()
}

```

```{r theme}
my_theme <- theme_bw() +
  theme(panel.grid = element_blank())

theme_set(my_theme)
```

```{r data-prep}
zones <- keys$bec %>% pull(zone)
groups <- keys$bec %>% pull(group) %>% unique()

bec_sample_loc <- here::here("data", "samples", "vlce_extracted.shp")

bec_samples_e <- vect(bec_sample_loc)

bec_df <- bec_samples_e %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI) %>%
  drop_na() %>%
  tibble() %>%
  filter(elev_cv <= 1000) %>% #erroneous pixels
  filter(elev_p95 >= 2000) # non-treed pixels

names_manual <- c(
  "bec",
  "category",
  "basal_area",
  "elev_cv",
  "elev_mean",
  "elev_p95",
  "elev_stddev",
  "gross_stem_volume",
  "loreys_height",
  "percentage_first_returns_above_2m",
  "percentage_first_returns_above_mean",
  "total_biomass",
  "CumDHI",
  "MinDHI",
  "VarDHI"
)

# names(bec_df) <- names(all_rasts)
names(bec_df) <- names_manual

bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 
```

```{r bec-nest}
to_nest <- bec_normal %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean)

bec_nested <- to_nest %>%
  group_by(bec) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = bec)

class_nested <- to_nest %>%
  group_by(category) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = category)

set.seed(69420)

strata_nested <- bind_rows(bec_nested, class_nested) %>%
  mutate(data = map(data, slice_sample, n = 500))
# this slice sample is the ANALYTICAL difference between 4_vegan and 
# 4_vegan_clean, other than the actual cleanup of the .rmd file

global <- strata_nested %>%
  filter(strata %in% zones) %>%
  unnest(data)
```

```{r global-model}
forest_types <- global %>%
  pull(category) %>%
  unique()

dhi_tab <- global %>%
  select(ends_with("DHI"))

primary <- global %>%
  select(elev_p95,
         elev_cv,
         percentage_first_returns_above_2m)

modelled <- global %>%
  select(basal_area,
         total_biomass,
         gross_stem_volume)

rda_out <-
  rda(
    dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/primary
      basal_area + gross_stem_volume + total_biomass,
    # modelled
    data = global
  )

ordiplot(rda_out, choices = c(1, 2))

rda_sum <- summary(rda_out)

var_exp <- rda_sum$constr.chi / rda_sum$tot.chi

rda_eig <- rda_out$CCA$eig %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  tibble() %>%
  pivot_longer(cols = everything()) %>%
  mutate(prop = value / sum(value),
         var_exp = prop * var_exp)

rcorr <- summary(rda_out)$sites %>%
  as.data.frame() %>%
  select(starts_with("RDA")) %>%
  bind_cols(global) %>%
  select(-strata,-category) %>%
  as.matrix() %>%
  Hmisc::rcorr()

rvals <- rcorr$r %>%
  as.data.frame() %>%
  select(starts_with("RDA")) %>%
  mutate(var = row.names(.)) %>%
  pivot_longer(starts_with("RDA"),
               values_to = "axis_loading",
               names_to = "axis") %>%
  filter(!startsWith(var, "RDA"))

pvals <- rcorr$P %>%
  as.data.frame() %>%
  select(starts_with("RDA")) %>%
  mutate(var = row.names(.)) %>%
  pivot_longer(starts_with("RDA"),
               values_to = "axis_p",
               names_to = "axis") %>%
  filter(!startsWith(var, "RDA"))

all_data_loadings <- left_join(rvals, pvals) %>%
  filter(axis != "RDA3") %>%
  mutate(sig = cut(
    axis_p,
    breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
    labels = c("***", "**", "*", ".", "NS"),
    right = F
  ))

var_out <- varpart(dhi_tab, primary, modelled)

var_out %>%
  plot(Xnames = c("Primary", "Modelled"))

path_r2 <- var_out$part$indfract %>%
  mutate(type = c("primary", "modelled", "overlap", "residual")) %>%
  select(type, Adj.R.squared) %>%
  tibble()
```

```{r r2-global}
RsquareAdj(rda_out)$adj.r.squared
```

```{r anova-global}
anova.cca(rda_out, step = 1000)

anova_out <- rda_out  %>%
  anova(by = "axis")

anova_out$`Pr(>F)` %>%
  as.data.frame() %>%
  rename(p_value = 1) %>%
  drop_na() %>%
  mutate(sig = cut(p_value, 
                   breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
                   labels = c("***", "**", "*", ".", "NS"))) %>%
  mutate(axis = paste0("RDA", row_number())) %>%
  relocate(axis)
```

```{r}
# linear relationships * homoschedasticity check

bec_normal %>%
  pivot_longer(cols = ends_with("DHI"),
               names_to = "dhi",
               values_to = "dhi_val") %>%
  pivot_longer(
    cols = basal_area:total_biomass,
    names_to = "struct",
    values_to = "struct_val"
  ) %>%
  ggplot(aes(x = struct_val, y = dhi_val)) +
  geom_hex() +
  facet_grid(rows = vars(dhi),
             cols = vars(struct),
             scales = "free") +
  labs(x = "Structure Value (Normalized)",
       y = "DHI Value (Normalized)")
```

```{r rda-var-funs}
# function to take a tibble and get a list out outputs that can then be unnested and examined as needed. function(tibble) gets rda and varpart, as plottable objects ideally. then get the var_out$part$indfract. i'll also need to look at the significance using an anova at some point

my_rda_fun <- function(tibble) {
  dhi_tab <- tibble %>%
    select(ends_with("DHI"))

  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/primary
      basal_area + gross_stem_volume + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

my_var_fun <- function(tibble) {
  dhi_tab <- tibble %>%
    select(ends_with("DHI"))

  primary <- tibble %>%
    select(elev_p95,
           elev_cv,
           # elev_stddev,
           percentage_first_returns_above_2m)

  modelled <- tibble %>%
    select(basal_area,
           total_biomass,
           gross_stem_volume)

  varpart(dhi_tab, primary, modelled)
}
```

```{r rda-var-bec}
outs <- strata_nested %>%
  mutate(rda = map(data, my_rda_fun, .progress = "RDA"),
         var = map(data, my_var_fun, .progress = "VAR")) 
```

```{r fracts-out}
fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("primary", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(strata, primary, modelled, overlap) %>%
  pivot_longer(c(primary, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()

fracts_out %>% 
  mutate(total = primary + modelled + overlap) %>%
  arrange(strata %in% forest_types,
          desc(total))
```

Fracts Bar Plot (Figure)

```{r}
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "British Columbia") %>% 
  relocate(strata)

fracts_all <- bind_rows(fracts_out,
                        fract_all)

fracts_tot <- fracts_all %>% 
  mutate(total = primary + modelled + overlap) %>%
  select(strata, total)

fracts_bar_tab <- fracts_all %>%
  pivot_longer(-strata) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "British Columbia", "BC", group)) %>%
  left_join(fracts_tot) %>%
  mutate(zone_nm = ifelse(is.na(zone_nm), strata, zone_nm))

fracts_bar_tab

fracts_bar_plot <- fracts_bar_tab %>%
  ggplot(aes(x = value, y = fct_reorder(zone_nm, total), fill = str_to_title(name))) +
  geom_col(position = "stack", col = "white") +
  facet_grid(rows = vars(fct_reorder(group, total, .desc = T)),
             scales = "free_y",
             space = "free") +
  labs(x = "Variance Explained",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom",
        panel.spacing.y = unit(0, "lines")) +
  guides(fill = guide_legend(reverse = T)) +
  scale_x_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = c("#6666ff", "#c22966", "#ff6666"))

fracts_bar_plot
ggsave(here::here("outputs", "fracts_bar_plot.png"), plot = fracts_bar_plot,
       height = 6, width = 8)
```

```{r boxplots-dhi-bec}
bec_df %>%
  select(bec, ends_with("DHI")) %>%
  pivot_longer(cols = ends_with("DHI")) %>%
  group_by(name) %>% 
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  mutate(bec = fct_reorder(bec, value)) %>%
  ggplot(aes(x = value, y = bec, fill = group)) +
  geom_boxplot() +
  facet_wrap(~name, scales = "free") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(x = "DHI Value",
       y = NULL,
       fill = NULL) +
  scale_fill_discrete_c4a_cat(palette = "batlow")
```

```{r eig-anova}
eig_outs <- outs %>%
  mutate(eig = map(
    rda,
    .f = function(rda) {
      rda_sum <- summary(rda)

      var_exp <- rda_sum$constr.chi / rda_sum$tot.chi

      rda_eig <- rda$CCA$eig %>%
        as.data.frame() %>%
        t() %>%
        as.data.frame() %>%
        tibble() %>%
        pivot_longer(cols = everything(), names_to = "axis",
                     values_to = "eigenvalue") %>%
        mutate(prop = eigenvalue / sum(eigenvalue),
               var_exp = prop * var_exp)
    }
  )) %>%
  unnest(eig) %>%
  select(strata, axis:var_exp)

anova_outs <- outs %>%
  mutate(
    anova = map(rda, .f = function(rda) {
      anova_out <- rda  %>%
        anova(by = "axis")

      anova_out$`Pr(>F)` %>%
        as.data.frame() %>%
        rename(anova_p = 1) %>%
        drop_na() %>%
        mutate(anova_sig = cut(
          anova_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS")))  %>%
        mutate(axis = paste0("RDA", row_number())) %>%
        relocate(axis)
  }, .progress = "anovas"))

anova_sig_axes <- anova_outs %>%
  select(strata, anova) %>%
  unnest(cols = anova) # %>%
  # filter(anova_sig != "NS")

axis_varexp <- anova_sig_axes %>%
  inner_join(eig_outs)

axis_varexp %>%
  filter(anova_sig != "NS") %>%
  filter(strata %in% forest_types)

axis_varexp %>%
  group_by(strata) %>%
  # filter(n() == 2) %>%
  select(strata, axis, var_exp) %>% 
  pivot_wider(names_from = axis,
              values_from = var_exp) %>%
  mutate(ratio = RDA1 / RDA2)

# smaller ratios present when the RDA1 doesn't explain much. when variance explained
# is low, the ratio is low. this is because there isn't much variance explained
```

```{r varexp}
rda_varexp <- outs %>%
  mutate(varexp = map(rda, .f = \(rda) {
    rda_sum <- summary(rda)
    var_exp <- rda_sum$constr.chi / rda_sum$tot.chi
  })) %>%
  unnest(varexp) %>%
  arrange(desc(varexp)) %>%
  select(strata, varexp)

rda_varexp %>%
  arrange(strata %in% forest_types,
          desc(varexp))
```

```{r bec-loadings}
out_loadings <- outs %>%
  mutate(loadings = map(rda, .f = function(rda) {
    summary(rda)$sites %>%
      as.data.frame() %>%
      select(starts_with("RDA"))
  }),
  loadings = map2(.x = loadings, .y = data, .f = bind_cols),
  loadings = map(loadings, .f = function(x) {
    rcorr <- x %>% select(-any_of(c("bec", "category"))) %>%
      as.matrix() %>%
      Hmisc::rcorr()

    rvals <- rcorr$r %>%
      as.data.frame() %>%
      select(starts_with("RDA")) %>%
      mutate(var = row.names(.)) %>%
      pivot_longer(starts_with("RDA"),
                   values_to = "axis_loading",
                   names_to = "axis") %>%
      filter(!startsWith(var, "RDA"))

    pvals <- rcorr$P %>%
      as.data.frame() %>%
      select(starts_with("RDA")) %>%
      mutate(var = row.names(.)) %>%
      pivot_longer(starts_with("RDA"),
                   values_to = "axis_p",
                   names_to = "axis") %>%
      filter(!startsWith(var, "RDA"))

    left_join(rvals, pvals) %>%
      mutate(n = rcorr$n[[1, 1]])
  }

  )) %>%
  select(strata, loadings) %>%
  unnest(loadings)
```

```{r}
sig_loads <- out_loadings %>% 
  mutate(loading_sig = cut(
          axis_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS"),
          right = F)) %>%
  inner_join(anova_sig_axes) %>%
  #filter(axis_sig == "NS") %>%
  arrange(desc(abs(axis_loading))) %>%
  filter(!(var %in% c("percentage_first_returns_above_mean", 
                    "elev_mean", 
                    "elev_stddev"))) %>%
  drop_na()# removes ima mindhi which has no variance

sig_loads %>%
  filter(anova_sig != "NS") %>%
  filter(endsWith(var, "DHI"),
         axis == "RDA2") %>%
  arrange(strata) %>%
  select(strata, var, axis_loading) %>%
  pivot_wider(names_from = var,
              values_from = axis_loading)
# when a second axis is present, it has a weak loading on the cumulative DHI, as compared to the other DHIs
# while the majority of the variability is contained within all three DHIs together, the second axis, when present, is more linked to the minmimum and variation DHI.

clean_loads_raw <- sig_loads %>%
  filter(anova_sig != "NS",
         loading_sig != "NS") %>%
  bind_rows(all_data_loadings %>%
              mutate(strata = "British Columbia") %>%
              rename(loading_sig = sig)) %>%
  select(strata:axis_loading)

clean_loads_abs <- clean_loads_raw %>%
  mutate(axis_loading = abs(axis_loading))
```

# run to here for analysis to be done

# drawio

```{r drawio}

# make dir
drawio_dir <- here::here("outputs", "drawio")

dir.create(drawio_dir, showWarnings = F)

# templates
one_rda <-
  read_lines(here::here(drawio_dir, "_rda_path-text_replace-RDA1.drawio"))

two_rda <-
  read_lines(here::here(drawio_dir, "_rda_path-text_replace-RDA2.drawio"))

# list to iterate
stratum <- clean_loads_raw %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads_raw %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

# fractions dfs
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "British Columbia") %>% 
  relocate(strata)

fracts_all <- fracts_out %>%
  bind_rows(fract_all)

# variance explained by axis

all_varexp <- rda_eig %>%
  filter(name != "RDA3") %>%
  mutate(strata = "British Columbia") %>%
  rename(axis = name) %>%
  select(-value)

varexp_all <- axis_varexp %>%
  filter(anova_sig != "NS") %>%
  select(strata, axis, prop, var_exp) %>%
  bind_rows(all_varexp)

for (s in stratum) {
  print(s)
  strata_tp <- clean_loads_raw %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

  if (strata_tp %>% pull(axis) %>% unique() %>% length() == 1) {
    strata_drawio <- one_rda
  } else {
    strata_drawio <- two_rda
  }

  textreplace <- strata_tp %>%
    pull(textreplace)

  replaceval <- strata_tp %>%
    pull(axis_loading)

  for (i in 1:length(textreplace)) {
    text <- textreplace[[i]]
    val <- replaceval[[i]]

    strata_drawio <- str_replace(strata_drawio, text, val)
  }

  strata_drawio <- str_replace(strata_drawio, "title_strata", s)

  #strata_drawio <- strata_drawio[!str_detect(strata_drawio, "RDA")]
  # filter through all possibilities remaining, replace them with NS

  for (p in possibles) {
    strata_drawio <- str_replace(strata_drawio, p, "NS")
  }

  strata_fract <- fracts_all %>%
    filter(strata == s) %>%
    pivot_longer(primary:overlap) %>%
    mutate(value = (value * 100) %>%
             round(1) %>%
             paste0(., "%"))

  fract_names <- strata_fract %>%
    pull(name)

  fract_value <- strata_fract %>%
    pull(value)

  for (i in 1:length(fract_names)) {
    strata_drawio <-
      str_replace(strata_drawio, fract_names[[i]], fract_value[[i]])
  }

  strata_varexp <- varexp_all %>%
    filter(strata == s) %>%
    mutate(var_exp = (var_exp * 100) %>%
             round(1) %>%
             paste0(., "%"))

  axis_names <- strata_varexp %>%
    pull(axis) %>%
    paste("var_exp", ., sep = "-")

  axis_vals <- strata_varexp %>%
    pull(var_exp)

  for (i in 1:length(axis_names)) {
    strata_drawio <-
      str_replace(strata_drawio, axis_names[[i]], axis_vals[[i]])
  }

  write_lines(strata_drawio, here::here(drawio_dir, paste0(s, ".drawio")))
}

file.copy(here::here("outputs", "drawio", "British Columbia.drawio"), here::here("outputs", "British Columbia.drawio"), overwrite = T)
```

Climate Table (Table 2)

```{r bec-clim-tab}
climate_folder <- here::here("data", "climate")
clim_vars <- climate_folder %>%
  list.files(pattern = "csv$") %>%
  tools::file_path_sans_ext()

clim_vars <- clim_vars[!str_detect(clim_vars, "bec")]

bec_clim <- climate_folder %>%
  list.files(pattern = "bec.csv$",
             full.names = T) %>%
  read_csv() %>%
  mutate(var = clim_vars %>% rep(16) %>% sort) %>%
  select(bec = ZONE, mean = MEAN, var) %>%
  pivot_wider(values_from = mean,
              names_from = var)

bec_tab <- keys$bec %>%
  left_join(bec_clim, by = c("zone" = "bec")) %>%
  select(-order) %>%
  arrange(group, zone) %>%
  relocate(group) %>%
  mutate(across(ppt:tmin, \(x) round(x, digits = 1))) %>%
  left_join(strata_nested %>% unnest(data) %>% count(strata),
            by = c("zone" = "strata"))

write_csv(bec_tab, here::here("data", "climate.csv"))

knitr::kable(bec_tab)
```

# rgb map

```{r}
bec_dissolve <- read_sf(here::here("data", "shapefiles", "bec_dissolve.shp")) %>%
  janitor::clean_names() %>%
  select(zone)

# normalized to max
loadings_rgb <- clean_loads_abs %>% 
  group_by(var) %>%
  mutate(axis_loading = axis_loading / max(axis_loading)) %>%
  group_by(axis) %>%
  pivot_wider(names_from = var,
              values_from = axis_loading) %>%
  ungroup() %>%
  replace(is.na(.), 0) %>%
  #mutate(across(elev_cv:elev_p95, \(x) ifelse(x > 0.5, 0.5, x))) %>%
  mutate(dhi_rgb = rgb(CumDHI, VarDHI, MinDHI),
         pri_rgb = rgb(elev_p95, percentage_first_returns_above_2m, elev_cv, maxColorValue = 1),
         mod_rgb = rgb(basal_area, total_biomass, gross_stem_volume, maxColorValue = 1)) %>%
  #select(strata, axis, percentage_first_returns_above_2m, elev_p95, elev_cv, ends_with("rgb")) %>%
  pivot_longer(ends_with("rgb"))

rgb_loads_split <- loadings_rgb %>%
  filter(strata %in% zones) %>%
  group_by(axis, name) %>%
  group_split()
```

```{r}
make_rgb_map <- function(tibble) {

  input <- bec_dissolve %>%
    left_join(tibble, by = c("zone" = "strata")) %>%
    mutate(value = ifelse(is.na(value), "#000000", value))

  rgb_cols = input %>%
    st_drop_geometry() %>%
    pull(value) %>%
    unique() %>%
    sort()

  input %>%
    ggplot(aes(x = zone, y = 2, fill = value)) +
    geom_col() +
    scale_fill_manual(values = rgb_cols)

   input %>%
    ggplot() +
    geom_sf(aes(geometry = geometry, fill = value, col = value), linewidth = 0.1) +
    scale_fill_manual(values = rgb_cols) +
    scale_color_manual(values = rgb_cols) +
    theme_void() +
    theme(legend.position = "none")
}

rgb_maps <- map(rgb_loads_split, make_rgb_map)

cowplot_rgb <- cowplot::plot_grid(rgb_maps[[3]], rgb_maps[[2]],
                                  rgb_maps[[6]], rgb_maps[[5]],
                                  labels = c("A - Primary RDA1", "B - Modelled RDA1",
                                             "C - Primary RDA2", "D - Modelled RDA2"),
                                  vjust = 1.25)

ggsave(here::here("outputs", "rgbplots_cow.png"), cowplot_rgb, height = 7, width = 8, bg = "white")

```

# flip cum to positive

```{r}
flip_cum <- function(data) {
  if(data %>% filter(var == "CumDHI") %>% nrow() == 0) {
    return(data)
  }
  flip <- data %>%
    filter(var == "CumDHI") %>%
    pull(axis_loading) %>%
    sign()

  data %>% mutate(axis_loading = axis_loading * flip)
}

clean_loads_cum_pos <- clean_loads_raw %>%
  group_by(strata, axis) %>%
  group_split() %>%
  map_dfr(., .f = flip_cum)

clean_loads_cum_pos %>%
  filter(strata %in% zones) %>%
  count(var, axis, sign = sign(axis_loading)) %>%
  pivot_wider(names_from = sign, values_from = n) %>%
  arrange(axis)

clean_loads_joined <- clean_loads_cum_pos %>%
  left_join(keys$bec %>%
              select(strata = zone, group)) %>%
  left_join(keys$continuous %>%
              select(var = variable, var_long)) %>%
  mutate(
    group = ifelse(is.na(group), strata, group),
    var_long = ifelse(is.na(var_long), var, var_long),
    pos_neg = ifelse(axis_loading > 0, "Positive", "Negative")
  )

```

Parcoord Noflip (Figure)

```{r}
data_parcoord <- clean_loads_joined %>%
  # mutate(axis_loading = abs(axis_loading)) %>%
  group_by(group, var_long, axis) %>%
  summarize(axis_loading = mean(axis_loading)) %>%
  ungroup() %>%
  complete(group, var_long, axis, fill = list(axis_loading = NA)) %>%
  mutate(inout = ifelse(endsWith(var_long, "DHI"), "Response", "Predictor"), 
         group = fct_relevel(group, "British Columbia", 
                             "Broadleaf", 
                             "Coniferous", 
                             "Mixed Wood", 
                             "Wetland-Treed", 
                             after = 0),
         var_long = ifelse(endsWith(var_long, "DHI"), var_long, str_replace_all(var_long, " ", "\n")),
         var_long = fct_relevel(var_long, 
                                c("Structural\nComplexity", "Canopy\nHeight", "Canopy\nCover",
                                "Basal\nArea", "Aboveground\nBiomass", "Gross\nStem\nVolume",
                                "Cumulative DHI", "Variation DHI", "Minimum DHI")))

# nojitter/lines
parcoord_noflip <- data_parcoord %>%
  ggplot(aes(x = var_long, y = axis_loading)) +
  #geom_boxplot(col = "#00000025", outlier.alpha = 0) +

  geom_line(aes(col = group, group = group), lty = "dashed") +
  geom_point(aes(col = group, group = group, shape = group, fill = group)) +
  geom_hline(yintercept = 0, lty = "dotted", col = "grey", linewidth = 0.5) +
  facet_wrap(~paste(axis, inout),
             scale = "free") +
  #theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL,
       y = "Axis Loading",
       col = NULL,
       shape = NULL,
       fill = NULL) +
  # coord_flip() +
  scale_colour_discrete_c4a_cat(palette = "bold") +
  scale_fill_discrete_c4a_cat(palette = "bold") +
  scale_shape_manual(values = c(21, 22, 22, 22, 22, 25, 25, 25, 25, 25))

parcoord_noflip

ggsave(here::here("outputs", "parcoord_noflip.png"), plot = parcoord_noflip,
       height = 6, width = 8)

# ggsave(here::here("outputs", "parcoord_noflip.svg"), plot = parcoord_noflip,
#        height = 6, width = 8)
```

# meg scatterplots

```{r}
scatter <- bec_normal %>%
  mutate(label = case_when(bec %in% c("CWH", "MS") ~ bec,
                           T ~ "Other"))

ggplot(data = scatter, aes(col = label), alpha = 0.25) +
  geom_point(aes(x = percentage_first_returns_above_2m, y = CumDHI)) +
  geom_density(aes(x = percentage_first_returns_above_2m)) +
  geom_density(aes(y = CumDHI)) +
  scale_fill_manual(values = c("grey", "red", "blue")) +
  scale_colour_manual(values = c("grey", "red", "blue"))

ggplot() +
  geom_point(data = bec_df, aes(x = percentage_first_returns_above_2m, y = CumDHI), alpha = 0.5) +
  geom_point(data = bec_df %>%
               filter(bec == "CWH"), aes(x = percentage_first_returns_above_2m, y = CumDHI), col = "orange") +
  geom_point(data = bec_df %>%
               filter(bec == "MS"), aes(x = percentage_first_returns_above_2m, y = CumDHI), col = "red") +
  geom_histogram(data = bec_df, aes(x = percentage_first_returns_above_2m)) +
  geom_histogram(data = bec_df, aes(y = CumDHI))
```

```{r}
fracts_out

strata_nested %>% 
  unnest(cols = data) %>%
  select(-bec, -category) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Type", group)) %>%
  ggplot() +
  geom_boxplot(aes(x = fct_reorder(.f = strata, .x = MinDHI, .fun = mean), y = MinDHI, fill = group))
```

```{r}
strata_nested %>% 
  unnest(cols = data) %>%
  select(-bec, -category) %>%
  group_by(strata) %>%
  summarize(across(basal_area:VarDHI, .fns = sd)) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Type", group)) %>%
  select(strata, group, basal_area:VarDHI) %>%
  pivot_longer(-c(strata, group)) %>%
  left_join(fracts_out) %>%
  mutate(varexp = primary + modelled + overlap) %>%
  ggplot(aes(x = value, y = varexp, col = group)) +
  geom_point() +
  facet_wrap(~name, strip.position = "bottom") +
  labs(x = NULL,
       y = "Variation Explained",
       title = "x axis is sd of variable") +
  theme(strip.placement = "outside",
        strip.background = element_blank())
```

Supplemental information Hexplot template

```{r}
bec_df %>%
  left_join(fracts_out %>%
              mutate(all = (primary + modelled + overlap) * 100) %>%
              select(strata, all), by = c("bec" = "strata")) %>%
  mutate(facet = paste0(bec, " (", round(all, 2), "%)")) %>%
  ggplot(aes(x = percentage_first_returns_above_2m / 100, y = CumDHI)) +
  geom_hex(aes(fill = after_stat(log(count)))) +
  labs(x = "Canopy Cover (%)") +
  facet_wrap(~fct_reorder(facet, all)) +
  theme(legend.position = "bottom")

ggsave(here::here("outputs", "supplement_hex.png"),
       height = 6, width = 8)
```

IQR sorted plot of canopy cover

```{r iqr-plot}
strata_nested %>% 
  unnest(data) %>% 
  ggplot(aes(x = fct_reorder(strata, percentage_first_returns_above_2m, IQR), y = percentage_first_returns_above_2m)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Study Area Map (Figure 1)

```{r study-map}
library(geodata)

provinces <- geodata::gadm(country = "Canada", level = 1, path = here::here("data", "province"))

prov_sf <- provinces %>%
  st_as_sf() %>%
  st_transform(3005)

bbox <- prov_sf %>%
  filter(NAME_1 == "British Columbia") %>%
  st_bbox()

usa <- geodata::gadm(country = "USA", level = 0, path = here::here("data", "usa")) %>%
  st_as_sf()

countries <- geodata::world(level = 0, path = here::here("data", "world")) %>%
  st_as_sf()

na_bbox <- countries %>%
  filter(GID_0 == "USA" |
         NAME_0 == "Canada") %>%
  st_bbox()

cad_bbox <- prov_sf %>%
  st_transform(3348) %>%
  st_bbox()

countries_map <- countries %>%
  filter(GID_0 == "USA" |
         NAME_0 == "Greenland") %>%
  st_transform(3348)

lakes <- read_sf(here::here("data", "great_lakes", "great_lakes.shp")) %>%
  st_transform(3348)

bec_map <- ggplot() +
  geom_sf(data = bcmaps::bc_bound(),
          fill = "grey70", col = NA) +
  geom_sf(
    data = prov_sf %>% filter(NAME_1 != "British Columbia"),
    fill = "grey70"
  ) +
  geom_sf(data = usa,
          fill = "grey50") +
  geom_sf(data = bec_dissolve %>%
            left_join(keys$bec),
          aes(fill = fct_reorder(zone_nm, zone)),
          colour = NA) +
  ggsci::scale_fill_igv() +
  labs(fill = NULL,
       linetype = NULL) +
  coord_sf(
    xlim = c(bbox$xmin - 300000, bbox$xmax),
    ylim = c(bbox$ymin, bbox$ymax),
    crs = 3005,
    datum = 4326
  ) +
  theme_grey() +
  theme(
    panel.background = element_rect(fill = "#525252"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = c(.15, .4),
    legend.background = element_rect(fill = NA),
    legend.box.background = element_rect(fill = "white")
  )

#ggsave(here::here("outputs", "bec_map_no_inset.png"), bec_map, device = "png")

inset = ggplot() +
  geom_sf(data = countries_map, fill = "grey50") +
  geom_sf(data = prov_sf %>% st_transform(3348) %>% filter(NAME_1 != "British Columbia"), fill = "grey70", size = 0.5) +
  geom_sf(data = prov_sf %>% st_transform(3348) %>% filter(NAME_1 == "British Columbia"), fill = "#e5e5e5", size = 0.5) +
  geom_sf(data = lakes, fill = "#525252") +
  geom_sf(data = bbox %>% st_as_sfc %>% st_transform(3348), fill = NA, colour = "red", size = 2) +
  coord_sf(xlim = c(cad_bbox$xmin - 300000, cad_bbox$xmax),
           ylim = c(cad_bbox$ymin, cad_bbox$ymax)) +
  theme_void() +
  theme(panel.border = element_rect(colour = "black"),
        panel.background = element_rect(fill = "#525252"))

# map_inset <- ggdraw(bec_map) +
#   draw_plot(inset, x = 0.7, y = 0.7, width = .28, height = .28)

map_inset <- bec_map + inset_element(inset, left = 0.71,
                                            bottom = 0.7,
                                            right = 1,
                                            top = 0.98, 
                                     align_to = "plot")

ggsave(here::here("outputs", "bec_map.png"), plot = map_inset, device = "png", height = 7, width = 10)

```